---
title: "Demo"
author: "Mayleen"
date: "2025-05-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
# Load Libraries
```{r, echo=FALSE}
library(readr)
library(tidyverse, warn.conflicts = FALSE)
library(data.table)
library(dplyr, warn.conflicts = FALSE)
source("SCRIPTS/clean_CORE.R")
source("SCRIPTS/clean_helper.R")
source("SCRIPTS/download_data.R")
source("SCRIPTS/clean_BMF.R")
```

Make sure you are in the home directory "rgai_summer2025_nccs"

# Download data
If you already downloaded the data you want to work with, you can skip this part!

```{r, eval=FALSE}
tscope_values <- c("501c3") # Download files from Tax Exempt Types in this list; possible values: "501c3", "501ce" 
fscope_values <- c("-pz") # Download files from IRS 990 Form Scope in this list; possible values: "-pz", "-pc", "-pf"
year_values <- seq(from = 1989, to = 2022, by = 1) # Download files from years in this list

download_CORE(tscope_values, fscope_values, year_values) # this function from download_data.R will download the data for you onto your local computer
download_dicts() # this function from download_data.R will download the data dictionaries and Business Master File (BMF) for you onto your local computer
```

# Clean Business Master File (BMF)
If you already cleaned the BMF file, you can skip this part.
```{r, eval=FALSE}
vars_to_keep <- c("EIN2", "NTEEV2", "F990_TOTAL_ASSETS_RECENT", "CENSUS_STATE_ABBR", "CENSUS_COUNTY_NAME", "ORG_YEAR_FIRST", "ORG_YEAR_LAST", "LATITUDE", "LONGITUDE", "CENSUS_CBSA_FIPS") # variables in BMF we wil keep (the rest will be dropped)
thresh <- 0.5*length(vars_to_keep) # Any rows with more NA values than this threshold will be dropped

dir.create("CLEAN")
clean_BMF(vars_to_keep, thresh) # this function from clean_BMF.R cleans the BMF file for you and saves to your computer in CLEAN folder
```

# Clean CORE Data and merge with BMF
If you already cleaned the CORE files and created the resulting merged dataset, you can skip this part.
```{r, eval=FALSE}
outcome_vars <- c("F9_10_ASSET_TOT_BOY", "F9_10_ASSET_TOT_EOY", "F9_01_REV_TOT_CY", "F9_08_REV_TOT_TOT")
predictor_vars <- c("F9_00_EXEMPT_STAT_501C3_X", "TAX_YEAR")
all_relevant_vars <- c(outcome_vars, predictor_vars, "EIN2", "F9_00_TAX_PERIOD_END_DATE")

prop_NA <- 0.5 # If a column has more than this fraction of missing values, warning message prints
file_name_tag <- "-501C3-CHARITIES-PZ-HRMN.csv" 
file_dir <- "CORE/pz/CORE-"
save_dir <- "CLEAN/pz/"

dir.create(c("CLEAN/pz", "CLEAN/pf", "CLEAN/pc"))
clean_CORE(all_relevant_vars, year_values, file_name_tag, file_dir, save_dir, prop_NA) # this function from clean_CORE.R cleans the CORE files, merges with cleaned BMF file, and saves resulting data frame
```

# Load the (cleaned/merged) data frame into the environment
```{r}
save_dir <- "CLEAN/pz/"
year_values <- seq(from = 1989, to = 2022, by = 1)

filename <- paste(save_dir, "cleanCORE_", year_values[1], "-", year_values[length(year_values)], ".csv", sep="")
df <- as.data.table(read_csv(filename, show_col_types = FALSE)) |> filter(TAX_YEAR != 2022)

head(df |> select(EIN2, TOT_REV, TAX_YEAR, NTEEV2, SIZE))
```

# Visualizations

## Histograms
```{r}
ggplot(data = df, mapping = aes(x = log(TOT_REV + 1))) +
  geom_histogram() +
      labs(x = "LOG(TOTAL REVENUE) in USD", title = "Histogram of log(total revenue)")
```

## Line Plots
Total revenue across all nonprofits over time
```{r}
df |>
  group_by(TAX_YEAR) |>
  summarise(sum_TOT_REV = sum(TOT_REV, na.rm = TRUE)) |>
  ggplot(aes(x = TAX_YEAR, y = sum_TOT_REV)) +
  geom_point() +
  geom_line() +
  labs(x = "Year", y = "Dollars", title = "Total Revenues, 1989-2022")
```

Total revenue across all nonprofits, grouped by size (total assets), over time
```{r}
size_labs <- c("Under $100,000", "$100,000 - $499,999", "$500,000 - $999,999", "$1 Million - $4.99 Million", "$5 Million - $9.99 Million", "Above $10 Million")
size_palette <- c("#1796D2", "#D2D2D2", "#000000", "#FDC010", "#ED028B", "#55B748")

df |>
  mutate(SIZE = factor(SIZE)) |>
  group_by(SIZE, TAX_YEAR) |>
  summarise(sum_TOT_REV = sum(TOT_REV, na.rm = TRUE)) |>
  ggplot(aes(x = TAX_YEAR, y = sum_TOT_REV, color = SIZE)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = size_palette, labels = size_labs, na.translate = FALSE) +
  labs(x = "Year", y = "Dollars", title = "Total Revenues, 1989-2022")
```

Total revenue across all nonprofts, grouped by NTEE code, over time
```{r}
ntee_labs <- c("Arts, Culture, and Humanities", "Education (minus Universities)", "Environment and Animals", "Health (minus Hospitals)", "Human Services", "Hospitals", "International, Foreign Affairs", "Mutual/Membership Benefit", "Public, Societal Benefit", "Religion Related", "Universities", "Other")

ntee_palette <- c("#1796D2", "#8DBCD2", "#989898", "#131313", "#F99C28", "#8A6908", "#EF2174", "#A7546C", "#5A7254", "#8A4846", "#DC2B28", "#55AE49")

df |>
  mutate(NTEEV2 = factor(NTEEV2)) |>
  group_by(NTEEV2, TAX_YEAR) |>
  summarise(sum_TOT_REV = sum(TOT_REV, na.rm = TRUE)) |>
  ggplot(aes(x = TAX_YEAR, y = sum_TOT_REV, color = NTEEV2)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = ntee_palette, labels = ntee_labs, na.translate = FALSE) +
  labs(x = "Year", y = "Dollars", title = "Total Revenues, 1989-2022")
```

Total revenue for each individual organization over time
```{r}
df |>
  ggplot(aes(x = TAX_YEAR, y = TOT_REV, color = EIN2)) +
  geom_line() +
  scale_color_discrete(na.translate = FALSE) +
  labs(x = "Year", y = "Dollars", title = "Total Revenues, 1989-2022") +
  theme(legend.position = "none")
```

Total revenue for individual organizations, but each plot is a different NTEE code group
```{r}
ntee_codes <- c("ART", "EDU", "ENV", "HEL", "HMS", "HOS", "IFA", "MMB", "PSB", "REL", "UNI", "UNU")

for (i in seq(1,length(ntee_codes),1)){
  my_plot <- df |>
  filter(NTEEV2 == ntee_codes[i]) |>
  ggplot(aes(x = TAX_YEAR, y = TOT_REV, color = EIN2)) +
  geom_line() +
  scale_color_discrete(na.translate = FALSE) +
  labs(x = "Year", y = "Dollars", title = paste("Total Revenues (", ntee_labs[i], "), 1989-2022", sep="")) +
  theme(legend.position = "none")
  print(my_plot)
}
```

Total revenue for individual organizations, but each plot is a different size (total assets)
```{r}
for (i in seq(1,length(size_labs),1)){
  my_plot <- df |>
  filter(SIZE == i) |>
  ggplot(aes(x = TAX_YEAR, y = TOT_REV, color = EIN2)) +
  geom_line() +
  scale_color_discrete(na.translate = FALSE) +
  labs(x = "Year", y = "Dollars", title = paste("Total Revenues, 1989-2022, SIZE: ", size_labs[i], sep="")) +
  theme(legend.position = "none")
  print(my_plot)
}
```

