---
title: "Analysis"
author: "Mayleen"
date: "2026-01-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# PanelMatch - No Interference
## Setup
```{r, message = F}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# These are some socioeconomic covariates we may evaluate balance for but need to convert from string to num
covs_to_check <- c("white_perc", "bachelors_perc", "poverty_perc", "unemployment", "med_family_income_adj", "med_household_income_adj")
df <- df |> mutate(across(all_of(covs_to_check), as.numeric)) # make sure these are numeric

# These are the variables we will assess balance for
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "pop_estimate", "med_precip", "med_avg_temp", "TOT_REV", "TOT_ASSET", "NUM_ORGS", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA")

df <- df |> select(all_of(c(c("CountyID", "TAX_YEAR", "treat.FEMA", "lat", "lng"), covs_to_check)))

#TODO: add spatially varying coefficients

# https://stackoverflow.com/questions/74815244/panelmatch-r-errors-please-convert-time-id-to-consecutive-integers-and-pleas
panel_df <- PanelData(panel.data = as.data.frame(df),
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA",
                      outcome = "TOT_REV")
rm(df)
saveRDS(panel_df, "no_int_panelV2.rds")
```

```{r}
sum(is.na(panel_df$treat.FEMA)) # county_FEMA_indicator <- 3.78 #2019

sum(df$TotPerCapADJ >= 2.78 & df$TotPerCapADJ < 3.78) # how many "suss" control units: 3433 or ~3.5% of total records
sum(df$TotPerCapADJ >= 3.78 & df$TotPerCapADJ < 4.78) # how many "suss" treated units: 2698 or ~2.8% of total records
sum(df$TotPerCapADJ <= 1) # 41983
quantile(df$TotPerCapADJ, probs = seq(0,1,0.1))

treatVarPlot <- DisplayTreatment(panel.data = panel_df, legend.position = "none", xlab = "TAX_YEAR", ylab = "County Code", hide.x.tick.label = TRUE, hide.y.tick.label = TRUE, dense.plot = TRUE)

treatVarPlot
#ggsave("treatment_variation_plot.png", plot = treatVarPlot, width = 7.5, height = 10, units = "in")
```

## Initial Imbalance
### Lag = 1
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# Set up the folder path and saved object names
L = 1 # lag, options: 1, 3, 5
folder_path_match <- "matches/"
folder_path_balance <- "balances/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
summary(matches_list$none)
summary(matches_list$ps.match)
# summary(matches_list$cbps.match)
# summary(matches_list$ps.weight)
# summary(matches_list$cbps.weight)
```


Default Scatter Plots
```{r}
plot(balances_list$none, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE)
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE)
# plot(balances_list$cbps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE)
# plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE)
# plot(balances_list$cbps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE)
```


Default Covariate Balance Plots
```{r}
plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$cbps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$cbps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```


My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "pop_estimate", "med_precip", "med_avg_temp", "TOT_REV", "TOT_ASSET", "NUM_ORGS", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
# bal_cbps.match <- tidy_bal_df(bal_data = bal_df$cbps.match, covariates = paste0("m.out.",covs_to_check))
# bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
# bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$cbps.weight, covariates = paste0("m.out.",covs_to_check))

p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = c(-0.15,0.05), 
                              ylim_ntee = c(-0.15,0.05),
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = c(-0.15,0.05), 
                                  ylim_ntee = c(-0.15,0.05), 
                                  ylim_unrefined = c(-0.15,0.05),
                                  include.unrefined.panel = TRUE)
# p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match, 
#                                   match_method = "CBPS Matching",
#                                   ylim_base = c(-0.15,0.05), 
#                                   ylim_ntee = c(-0.15,0.05), 
#                                   include.unrefined.panel = FALSE)
# p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight, 
#                                   match_method = "PS Weighting",
#                                   ylim_base = c(-0.15,0.05), 
#                                   ylim_ntee = c(-0.15,0.05), 
#                                   include.unrefined.panel = FALSE)
# p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight, 
#                                   match_method = "CBPS Weighting",
#                                   ylim_base = c(-0.15,0.05), 
#                                   ylim_ntee = c(-0.15,0.05), 
#                                   include.unrefined.panel = FALSE)

p_none$base
p_none$ntee

p_ps.match$base
p_ps.match$base_unrefined
p_ps.match$ntee
p_ps.match$ntee_unrefined

# p_cbps.match$base
# p_cbps.match$ntee
# 
# p_ps.weight$base
# p_ps.weight$ntee
# 
# p_cbps.weight$base
# p_cbps.weight$ntee
```


### Lag = 3
...


### Outcome 5-years post-disaster
```{r}
init.L1F5 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "none",
                 qoi = "att",
                 match.missing = FALSE,
                 lead = 5,
                 matching = FALSE,
                 placebo.test = TRUE)

summary(init.L1F5)

bal.init.L1F5 <- get_covariate_balance(init.L1F5, panel.data = panel_df, covariates = covs_to_check)

temp <- summary(bal.init.L1F5, include.unrefined = FALSE)
bal_df <- as.data.frame(temp$init.L1F5)
write.csv(bal_df, file = "balance_tables.csv")

bal_df
plot(bal.init.L1F5, type = "panel", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.2, 0.2)) 
plot(bal.init.L1F5, type = "scatter", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.05, 0.2))
```

```{r}
init.L3F5 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "none",
                 qoi = "att",
                 match.missing = FALSE,
                 lead = 5,
                 matching = FALSE,
                 placebo.test = TRUE)

summary(init.L3F5)

bal.init.L3F5 <- get_covariate_balance(init.L3F5, panel.data = panel_df, covariates = covs_to_check)

temp <- summary(bal.init.L3F5, include.unrefined = FALSE)
bal_df <- as.data.frame(temp$init.L3F5)
write.csv(bal_df, file = "balance_tables.csv")

bal_df
plot(bal.init.L3F5, type = "panel", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.2, 0.2))
plot(bal.init.L3F5, type = "scatter", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.05, 0.2))
```


```{r}
init.L5F5 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "none",
                 qoi = "att",
                 match.missing = FALSE,
                 lead = 5,
                 matching = FALSE,
                 placebo.test = TRUE)

summary(init.L5F5)

bal.init.L5F5 <- get_covariate_balance(init.L5F5, panel.data = panel_df, covariates = covs_to_check)

temp <- summary(bal.init.L5F5, include.unrefined = FALSE)
bal_df <- as.data.frame(temp$init.L5F5)
write.csv(bal_df, file = "balance_tables.csv")

bal_df
plot(bal.init.L5F5, type = "panel", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.2, 0.2))
plot(bal.init.L5F5, type = "scatter", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.05, 0.2))
```


## Matching Methods - No SVC
```{r}
formula_L1 <- "~ TAX_YEAR + lat*lng + I(lag(NUM_ORGS, 1:1)) + I(lag(TOT_ASSET, 1:1)) + I(lag(bachelors_perc, 0:1)) + I(lag(med_household_income_adj, 0:1)) + I(lag(white_perc, 0:1)) + I(lag(pop_estimate, 0:1)) + I(lag(TOT_REV, 1:1)"

formula_L3 <- "~ TAX_YEAR + lat*lng + I(lag(NUM_ORGS, 1:3)) + I(lag(TOT_ASSET, 1:3)) + I(lag(bachelors_perc, 0:3)) + I(lag(med_household_income_adj, 0:3)) + I(lag(white_perc, 0:3)) + I(lag(pop_estimate, 0:3)) + I(lag(TOT_REV, 1:3)"

formula_L5 <- "~ TAX_YEAR + lat*lng + I(lag(NUM_ORGS, 1:5)) + I(lag(TOT_ASSET, 1:5)) + I(lag(bachelors_perc, 0:5)) + I(lag(med_household_income_adj, 0:5)) + I(lag(white_perc, 0:5)) + I(lag(pop_estimate, 0:5)) + I(lag(TOT_REV, 1:5)"
```

### Outcome 1-year post-disaster
#### Propensity Score Matching
```{r}
ps.matchL1F1 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "ps.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L1),
                 lead = 1,
                 placebo.test = TRUE)

ps.matchL3F1 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "ps.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L3),
                 lead = 1,
                 placebo.test = TRUE)

ps.matchL5F1 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "ps.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L5),
                 lead = 1,
                 placebo.test = TRUE)

summary(ps.matchL1F1)
summary(ps.matchL3F1)
summary(ps.matchL5F1)
```


```{r}
plot(ps.matchL1F1)
plot(ps.matchL3F1)
plot(ps.matchL5F1)
```


```{r}
bal.ps.matchF1 <- get_covariate_balance(ps.matchL1F1, ps.matchL3F1, ps.matchL5F1, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.ps.matchF1)

temp <- summary(bal.ps.matchF1)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/ps_matchF1.xlsx")
```


```{r}
plot(bal.ps.matchF1, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, ylim = c(-0.05,0.05))
plot(bal.ps.matchF1, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```

#### CBPS Matching
```{r}
cbps.matchL1F1 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "CBPS.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L1),
                 lead = 1,
                 placebo.test = TRUE)

cbps.matchL3F1 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "CBPS.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L3),
                 lead = 1,
                 placebo.test = TRUE)

cbps.matchL5F1 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "CBPS.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L5),
                 lead = 1,
                 placebo.test = TRUE)

summary(cbps.matchL1F1)
summary(cbps.matchL3F1)
summary(cbps.matchL5F1)
```


```{r}
plot(cbps.matchL1F1)
plot(cbps.matchL3F1)
plot(cbps.matchL5F1)
```


```{r}
bal.cbps.matchF1 <- get_covariate_balance(cbps.matchL1F1, cbps.matchL3F1, cbps.matchL5F1, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.ps.matchF1)

temp <- summary(bal.cbps.matchF1)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/cbps_matchF1.xlsx")
```

```{r}
plot(bal.cbps.matchF1, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, ylim = c(-0.05,0.05))
plot(bal.cbps.matchF1, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```

#### Propensity Score Weighting

```{r}
ps.weightL1F1 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "ps.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L1),
                 lead = 1,
                 placebo.test = TRUE)

ps.weightL3F1 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "ps.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L3),
                 lead = 1,
                 placebo.test = TRUE)

ps.weightL5F1 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "ps.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L5),
                 lead = 1,
                 placebo.test = TRUE)

summary(ps.weightL1F1)
summary(ps.weightL3F1)
summary(ps.weightL5F1)
```


```{r}
plot(ps.weightL1F1)
plot(ps.weightL3F1)
plot(ps.weightL5F1)
```


```{r}
bal.ps.weightF1 <- get_covariate_balance(ps.weightL1F1, ps.weightL3F1, ps.weightL5F1, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.ps.matchF1)

temp <- summary(bal.ps.weightF1)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/ps.weightF1.xlsx")
```

```{r}
plot(bal.ps.weightF1, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, ylim = c(-0.2,0.2))
plot(bal.ps.weightF1, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```

#### CBPS Weighting
```{r}
cbps.weightL1F1 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "CBPS.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L1),
                 lead = 1,
                 placebo.test = TRUE)

cbps.weightL3F1 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "CBPS.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L3),
                 lead = 1,
                 placebo.test = TRUE)

cbps.weightL5F1 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "CBPS.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L5),
                 lead = 1,
                 placebo.test = TRUE)

summary(cbps.weightL1F1)
summary(cbps.weightL3F1)
summary(cbps.weightL5F1)
```


```{r}
plot(cbps.weightL1F1)
plot(cbps.weightL3F1)
plot(cbps.weightL5F1)
```


```{r}
bal.cbps.weightF1 <- get_covariate_balance(cbps.weightL1F1, cbps.weightL3F1, cbps.weightL5F1, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.cbps.weightF1)

temp <- summary(bal.cbps.weightF1)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/cbps_weightF1.xlsx")

saveRDS(list("L1F1" = cbps.weightL1F1, "L3F1" = cbps.weightL3F1, "L5F1" = cbps.weightL5F1), 
        file = "matches/cbps_weightF1.rds")

saveRDS(bal.cbps.weightF1, file = "balances/cbps_weightF1.rds")
```

```{r}
plot(bal.cbps.weightF1, type = "panel", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.2,0.2))
plot(bal.cbps.weightF1, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```


### Outcome 3-year post-disaster
#### Propensity Score Matching
```{r}
ps.matchL1F3 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "ps.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L1),
                 lead = 3,
                 placebo.test = TRUE)

ps.matchL3F3 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "ps.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L3),
                 lead = 3,
                 placebo.test = TRUE)

ps.matchL5F3 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "ps.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L5),
                 lead = 3,
                 placebo.test = TRUE)

summary(ps.matchL1F3)
summary(ps.matchL3F3)
summary(ps.matchL5F3)
```


```{r}
plot(ps.matchL1F3)
plot(ps.matchL3F3)
plot(ps.matchL5F3)
```


```{r}
bal.ps.matchF3 <- get_covariate_balance(ps.matchL1F3, ps.matchL3F3, ps.matchL5F3, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.ps.matchF1)

temp <- summary(bal.ps.matchF3)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/ps_matchF3.xlsx")
```


```{r}
plot(bal.ps.matchF3, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, ylim = c(-0.2,0.2))
plot(bal.ps.matchF3, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```

#### CBPS Matching
```{r}
cbps.matchL1F3 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "CBPS.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L1),
                 lead = 3,
                 placebo.test = TRUE)

cbps.matchL3F3 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "CBPS.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L3),
                 lead = 3,
                 placebo.test = TRUE)

cbps.matchL5F3 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "CBPS.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L5),
                 lead = 3,
                 placebo.test = TRUE)

summary(cbps.matchL1F3)
summary(cbps.matchL3F3)
summary(cbps.matchL5F3)
```


```{r}
plot(cbps.matchL1F3)
plot(cbps.matchL3F3)
plot(cbps.matchL5F3)
```


```{r}
bal.cbps.matchF3 <- get_covariate_balance(cbps.matchL1F3, cbps.matchL3F3, cbps.matchL5F3, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.ps.matchF3)

temp <- summary(bal.cbps.matchF3)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/cbps_matchF3.xlsx")
```

```{r}
plot(bal.cbps.matchF3, type = "panel", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.2,0.2))
plot(bal.cbps.matchF3, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```

#### Propensity Score Weighting

```{r}
ps.weightL1F3 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "ps.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L1),
                 lead = 3,
                 placebo.test = TRUE)

ps.weightL3F3 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "ps.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L3),
                 lead = 3,
                 placebo.test = TRUE)

ps.weightL5F3 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "ps.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L5),
                 lead = 3,
                 placebo.test = TRUE)

summary(ps.weightL1F3)
summary(ps.weightL3F3)
summary(ps.weightL5F3)
```


```{r}
plot(ps.weightL1F3)
plot(ps.weightL3F3)
plot(ps.weightL5F3)
```


```{r}
bal.ps.weightF3 <- get_covariate_balance(ps.weightL1F3, ps.weightL3F3, ps.weightL5F3, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.ps.matchF1)

temp <- summary(bal.ps.weightF3)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/ps_weightF3.xlsx")
```

```{r}
plot(bal.ps.weightF3, type = "panel", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.2,0.2))
plot(bal.ps.weightF3, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```

#### CBPS Weighting
```{r}
cbps.weightL1F3 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "CBPS.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L1),
                 lead = 3,
                 placebo.test = TRUE)

cbps.weightL3F3 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "CBPS.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L3),
                 lead = 3,
                 placebo.test = TRUE)

cbps.weightL5F3 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "CBPS.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L5),
                 lead = 3,
                 placebo.test = TRUE)

summary(cbps.weightL1F3)
summary(cbps.weightL3F3)
summary(cbps.weightL5F3)
```


```{r}
plot(cbps.weightL1F3)
plot(cbps.weightL3F3)
plot(cbps.weightL5F3)
```


```{r}
bal.cbps.weightF3 <- get_covariate_balance(cbps.weightL1F3, cbps.weightL3F3, cbps.weightL5F3, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.cbps.weightF3)

temp <- summary(bal.cbps.weightF3)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/cbps_weightF3.xlsx")

saveRDS(list("L1F3" = cbps.weightL1F3, "L3F3" = cbps.weightL3F3, "L5F3" = cbps.weightL5F3), 
        file = "matches/cbps_weightF3.rds")

saveRDS(bal.cbps.weightF3, file = "balances/cbps_weightF3.rds")
```

```{r}
plot(bal.cbps.weightF3, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, ylim = c(-0.2,0.2))
plot(bal.cbps.weightF3, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```


### Outcome 5-year post-disaster
#### Propensity Score Matching
```{r}
ps.matchL1F5 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "ps.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L1),
                 lead = 5,
                 placebo.test = TRUE)

ps.matchL3F5 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "ps.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L3),
                 lead = 5,
                 placebo.test = TRUE)

ps.matchL5F5 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "ps.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L5),
                 lead = 5,
                 placebo.test = TRUE)

summary(ps.matchL1F5)
summary(ps.matchL3F5)
summary(ps.matchL5F5)
```


```{r}
plot(ps.matchL1F5)
plot(ps.matchL3F5)
plot(ps.matchL5F5)
```


```{r}
bal.ps.matchF5 <- get_covariate_balance(ps.matchL1F5, ps.matchL3F5, ps.matchL5F5, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.ps.matchF5)

temp <- summary(bal.ps.matchF5)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/ps_matchF5.xlsx")
```


```{r}
plot(bal.ps.matchF5, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, ylim = c(-0.2,0.2))
plot(bal.ps.matchF5, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```

#### CBPS Matching
```{r}
cbps.matchL1F5 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "CBPS.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L1),
                 lead = 5,
                 placebo.test = TRUE)

cbps.matchL3F5 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "CBPS.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L3),
                 lead = 5,
                 placebo.test = TRUE)

cbps.matchL5F5 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "CBPS.match",
                 qoi = "att",
                 size.match = 10,
                 covs.formula = as.formula(formula_L5),
                 lead = 5,
                 placebo.test = TRUE)

summary(cbps.matchL1F5)
summary(cbps.matchL3F5)
summary(cbps.matchL5F5)
```


```{r}
plot(cbps.matchL1F5)
plot(cbps.matchL3F5)
plot(cbps.matchL5F5)
```


```{r}
bal.cbps.matchF5 <- get_covariate_balance(cbps.matchL1F5, cbps.matchL3F5, cbps.matchL5F5, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.ps.matchF3)

temp <- summary(bal.cbps.matchF5)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/cbps_matchF5.xlsx")
```

```{r}
plot(bal.cbps.matchF5, type = "panel", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.2,0.2))
plot(bal.cbps.matchF5, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```

#### Propensity Score Weighting

```{r}
ps.weightL1F5 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "ps.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L1),
                 lead = 5,
                 placebo.test = TRUE)

ps.weightL3F5 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "ps.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L3),
                 lead = 5,
                 placebo.test = TRUE)

ps.weightL5F5 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "ps.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L5),
                 lead = 5,
                 placebo.test = TRUE)

summary(ps.weightL1F5)
summary(ps.weightL3F5)
summary(ps.weightL5F5)
```


```{r}
plot(ps.weightL1F5)
plot(ps.weightL3F5)
plot(ps.weightL5F5)
```


```{r}
bal.ps.weightF5 <- get_covariate_balance(ps.weightL1F5, ps.weightL3F5, ps.weightL5F5, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.ps.matchF1)

temp <- summary(bal.ps.weightF5)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/ps_weightF5.xlsx")
```

```{r}
plot(bal.ps.weightF5, type = "panel", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.2,0.2))
plot(bal.ps.weightF5, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```

#### CBPS Weighting
```{r}
cbps.weightL1F5 <- PanelMatch(panel.data = panel_df,
                 lag = 1,
                 refinement.method = "CBPS.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L1),
                 lead = 5,
                 placebo.test = TRUE)

cbps.weightL3F5 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "CBPS.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L3),
                 lead = 5,
                 placebo.test = TRUE)

cbps.weightL5F5 <- PanelMatch(panel.data = panel_df,
                 lag = 5,
                 refinement.method = "CBPS.weight",
                 qoi = "att",
                 covs.formula = as.formula(formula_L5),
                 lead = 5,
                 placebo.test = TRUE)

summary(cbps.weightL1F5)
summary(cbps.weightL3F5)
summary(cbps.weightL5F5)
```


```{r}
plot(cbps.weightL1F5)
plot(cbps.weightL3F5)
plot(cbps.weightL5F5)
```


```{r}
bal.cbps.weightF5 <- get_covariate_balance(cbps.weightL1F5, cbps.weightL3F5, cbps.weightL5F5, 
                                     panel.data = panel_df,
                                     covariates = covs_to_check)
#summary(bal.cbps.weightF5)

temp <- summary(bal.cbps.weightF5)
temp <- lapply(temp, as.data.frame)
temp <- lapply(temp, tibble::rownames_to_column, var = "time")

write_xlsx(temp, "balances/cbps_weightF5.xlsx")

saveRDS(list("L1F5" = cbps.weightL1F5, "L3F5" = cbps.weightL3F5, "L5F5" = cbps.weightL5F5), 
        file = "matches/cbps_weightF5.rds")

saveRDS(bal.cbps.weightF5, file = "balances/cbps_weightF5.rds")
```

```{r}
plot(bal.cbps.weightF5, type = "panel", include.unrefined.panel = FALSE, legend = FALSE, ylim = c(-0.2,0.2))
plot(bal.cbps.weightF5, type = "scatter", legend = TRUE, ylim = c(-0.05, 0.2))
```


## Matching Methods - SVC
```{r}
# File: perform_refinement.R
# glm(reformulate(response = treatment, termlabels = colnames(pooled)[-c(1:3)]), family = binomial(link = "logit"), data = pooled)
# CBPS::CBPS(reformulate(response = treatment, termlabels = colnames(pooled)[-c(1:3)]), family = binomial(link = "logit"), data = pooled)

library(fields)
library(MASS)

set.seed(2727)
county_subset <- sample.int(n = max(df$CountyID), size = 100, replace = FALSE)
df_sub <- df |> filter(CountyID %in% county_subset)

x <- cbind(df_sub$lat, df_sub$lng) # inputs: latitude and longitude

# SVC
d <- rdist(x)
K <- Matern(d, smoothness=.5) # for smoothness paramter, should we do some sort of optimization?
E <- eigen(K)

y <- as.matrix(df_sub$treat.FEMA)
x_s <- cbind(x[,1] * E$vectors[,1:45], x[,2] * E$vectors[,1:45]) # would have to play around with this number

# add SVC back into panel data
xs_vars <- paste0("x_s", seq_len(ncol(x_s)))
colnames(x_s) <- xs_vars
df_sub <- cbind(df_sub, x_s)

formula_tst <- paste(
  "y ~ bachelors_perc + med_household_income_adj + white_perc + pop_estimate + med_precip + med_avg_temp +",
  paste(xs_vars, collapse = " + "))

# regression
ls <- glm(as.formula(formula_tst), data = df_sub, family = binomial(link = "logit"))
summary(ls)
```


```{r}
panel_df <- PanelData(panel.data = df_sub,
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA",
                      outcome = "TOT_REV")


formula_L1 <- paste(
  "~ I(lag(bachelors_perc, 0:1)) + I(lag(med_household_income_adj, 0:1)) + I(lag(white_perc, 0:1)) + I(lag(pop_estimate, 0:1)) + med_precip + med_avg_temp +",
  paste(xs_vars, collapse = " + "))

formula_L3 <- paste(
  "~ I(lag(bachelors_perc, 0:3)) + I(lag(med_household_income_adj, 0:3)) + I(lag(white_perc, 0:3)) + I(lag(pop_estimate, 0:3)) + med_precip + med_avg_temp +",
  paste(xs_vars, collapse = " + "))

formula_L5 <- paste(
  "~ I(lag(bachelors_perc, 0:5)) + I(lag(med_household_income_adj, 0:5)) + I(lag(white_perc, 0:5)) + I(lag(pop_estimate, 0:5)) + med_precip + med_avg_temp +",
  paste(xs_vars, collapse = " + "))
```

# PanelMatch - Interference

```{r}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)
library(geosphere)
source("matches_helper.R")

df <- readRDS("county_long.rds") |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# These are some socioeconomic covariates we need to convert from char to numeric
covs_to_check <- c("white_perc", "bachelors_perc", "poverty_perc", "unemployment", "med_family_income_adj", "med_household_income_adj")
df <- df |> mutate(across(all_of(covs_to_check), as.numeric))

# These are the variables we will assess balance for
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "pop_estimate", "med_precip", "med_avg_temp", "TOT_REV", "TOT_ASSET", "NUM_ORGS", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA")

df <- df |> select(all_of(c(c("CountyID", "geoid_2010", "TAX_YEAR", "treat.FEMA", "lat", "lng"), covs_to_check)))

# Create new treatment variable that takes a distance-based cutoff adjustment
adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 50)
res <- add_exposed(panel_data = df, adj_list = adj_lists$union)

#TODO: add spatially varying coefficients

# https://stackoverflow.com/questions/74815244/panelmatch-r-errors-please-convert-time-id-to-consecutive-integers-and-pleas
panel_df <- PanelData(panel.data = as.data.frame(res$df),
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA.adj",
                      outcome = "TOT_REV")

saveRDS(panel_df, "int_panelV1.rds")

DisplayTreatment(panel.data = panel_df, legend.position = "none", xlab = "TAX_YEAR", ylab = "County Code", hide.x.tick.label = TRUE, hide.y.tick.label = TRUE, dense.plot = TRUE)
```

```{r}
none.L3 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "none",
                 qoi = "att",
                 match.missing = TRUE,
                 lead = 1:5,
                 matching = FALSE)

bal.none.L3 <- get_covariate_balance(none.L3, panel.data = panel_df, covariates = covs_to_check)

temp <- summary(bal.none.L3, include.unrefined = FALSE)
bal_df <- as.data.frame(temp$none.L3)
#write.csv(bal_df, file = "balance_tables.csv")

summary(none.L3)
plot(none.L3)
bal_df
plot(bal.none.L3, type = "panel", include.unrefined.panel = FALSE, legend = TRUE) # ylim = c(-0.5, 0.5)
plot(bal.none.L3, type = "scatter", include.unrefined.panel = FALSE)
```

```{r}
formulas <- list("L1" = "~ TAX_YEAR + lat*lng + I(lag(TOT_ASSET, 1:1)) + I(lag(NTEEV2_ART, 1:1)) + I(lag(NTEEV2_EDU, 1:1)) + I(lag(NTEEV2_ENV, 1:1)) + I(lag(NTEEV2_HEL, 1:1)) + I(lag(NTEEV2_HMS, 1:1)) + I(lag(NTEEV2_HOS, 1:1)) + I(lag(NTEEV2_IFA, 1:1)) + I(lag(NTEEV2_MMB, 1:1)) + I(lag(NTEEV2_PSB, 1:1)) + I(lag(NTEEV2_REL, 1:1)) + I(lag(NTEEV2_UNI, 1:1)) + I(lag(NTEEV2_UNU, 1:1)) + I(lag(NTEEV2_NA, 1:1)) + I(lag(bachelors_perc, 0:1)) + I(lag(med_household_income_adj, 0:1)) + I(lag(white_perc, 0:1)) + I(lag(pop_estimate, 0:1)) + I(lag(TOT_REV, 1:1))", 
                 "L3" = "~ TAX_YEAR + lat*lng + I(lag(TOT_ASSET, 1:3)) + I(lag(NTEEV2_ART, 1:3)) + I(lag(NTEEV2_EDU, 1:3)) + I(lag(NTEEV2_ENV, 1:3)) + I(lag(NTEEV2_HEL, 1:3)) + I(lag(NTEEV2_HMS, 1:3)) + I(lag(NTEEV2_HOS, 1:3)) + I(lag(NTEEV2_IFA, 1:3)) + I(lag(NTEEV2_MMB, 1:3)) + I(lag(NTEEV2_PSB, 1:3)) + I(lag(NTEEV2_REL, 1:3)) + I(lag(NTEEV2_UNI, 1:3)) + I(lag(NTEEV2_UNU, 1:3)) + I(lag(NTEEV2_NA, 1:3)) + I(lag(bachelors_perc, 0:3)) + I(lag(med_household_income_adj, 0:3)) + I(lag(white_perc, 0:3)) + I(lag(pop_estimate, 0:3)) + I(lag(TOT_REV, 1:3))", 
                 "L5" = "~ TAX_YEAR + lat*lng + I(lag(TOT_ASSET, 1:5)) + I(lag(NTEEV2_ART, 1:5)) + I(lag(NTEEV2_EDU, 1:5)) + I(lag(NTEEV2_ENV, 1:5)) + I(lag(NTEEV2_HEL, 1:5)) + I(lag(NTEEV2_HMS, 1:5)) + I(lag(NTEEV2_HOS, 1:5)) + I(lag(NTEEV2_IFA, 1:5)) + I(lag(NTEEV2_MMB, 1:5)) + I(lag(NTEEV2_PSB, 1:5)) + I(lag(NTEEV2_REL, 1:5)) + I(lag(NTEEV2_UNI, 1:5)) + I(lag(NTEEV2_UNU, 1:5)) + I(lag(NTEEV2_NA, 1:5)) + I(lag(bachelors_perc, 0:5)) + I(lag(med_household_income_adj, 0:5)) + I(lag(white_perc, 0:5)) + I(lag(pop_estimate, 0:5)) + I(lag(TOT_REV, 1:5))"
)

ps.L3 <- PanelMatch(panel.data = panel_df,
                 lag = 3,
                 refinement.method = "ps.match",
                 qoi = "att",
                 match.missing = TRUE,
                 lead = 1:5,
                 covs.formula = as.formula(formulas$L3),
                 matching = TRUE)

bal.ps.L3 <- get_covariate_balance(ps.L3, panel.data = panel_df, covariates = covs_to_check)

temp <- summary(bal.ps.L3, include.unrefined = FALSE)
bal_df <- as.data.frame(temp$ps.L3)
#write.csv(bal_df, file = "balance_tables.csv")

summary(ps.L3)
plot(ps.L3)
bal_df
plot(bal.ps.L3, type = "panel", include.unrefined.panel = FALSE, legend = TRUE) # ylim = c(-0.5, 0.5)
plot(bal.ps.L3, type = "scatter", include.unrefined.panel = TRUE)
```


# MATCHIT: Total Effect of a County Experiencing 1 or More Disasters on the County's Nonprofit Revenue
Let the unit of analysis $i$ represent a US county.
Let $A_i^X$ represent the treatment of county $i$ in year $X$. If county $i$ had nonzero disaster damage in the year $X$, then $A_i^X=1$. Otherwise, $A_i^X=0$.
Let $Y_{tot,i}^X$ be the total nonprofit revenue of county $i$ reported in the year $X$. Let $Y_{med,i}^X$ be the median nonprofit revenue of county $i$ reported in the year $X$. 

First, let's create a county-level dataset where each row corresponds to a county and we have a column `treatment` representing $A_i^{2000}$. We will also include the following:
- geoid of the county (FIPS code)
- state the county is in 
- latitude and longitude of the centroid of the county
- the median temperature for the state the county belongs to for the years 1997, 1998, 1999
- socioeconomic factors for the year 1990
- the total number of disasters in that county in the period 1997-1999
- US Census Region or Division
## Data Pre-processing
```{r, message = F}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)

YEAR <- 2000

# Basic county information (lat/long of centroid, county FIPS, county name, state)
counties <- as.data.table(read_csv("counties_data/uscounties.csv", show_col_types = FALSE)) # https://simplemaps.com/data/us-counties

# socioeconomic factors
source("https://raw.githubusercontent.com/UI-Research/nccs-geo/main/get_census_data.R")
census_df  <- get_census_data( geo="county", years=c(1990))   #3,142 counties

df <- merge(census_df, counties, by.x = "geoid_2010", by.y = "county_fips", all.x = TRUE)

# removing unnecessary columns
df <- df |> select(-population, -county_ascii)

# disaster data
disasters <- readRDS("disasters.rds")
disasters <- disasters |> 
      filter(TAX_YEAR >= YEAR-3 & TAX_YEAR <= YEAR) |> 
      group_by(county.census.geoid, TAX_YEAR) |>
      summarize(
            n_disasters = n(),
            across(where(is.numeric), sum, .names = "{.col}_TOT")
      )

df <- merge(df, disasters |> 
                  filter(TAX_YEAR == YEAR) |> 
                  mutate(LOG_TotPropDmg2015 = log(`PropertyDmg(ADJ 2015)_TOT` + 1)) |> 
                  select(county.census.geoid, n_disasters, LOG_TotPropDmg2015), 
            by.x = 'geoid_2010', by.y = 'county.census.geoid', all.x = TRUE)

df <- df |> mutate(n_disasters = replace_na(n_disasters, 0),
                   LOG_TotPropDmg2015 = replace_na(LOG_TotPropDmg2015, 0))

df <- merge(df, disasters |> filter(TAX_YEAR < YEAR) |> group_by(county.census.geoid) |> summarize(n_disasters.PRE = sum(n_disasters)),
            by.x = 'geoid_2010', by.y = 'county.census.geoid', all.x = TRUE )

# If missing, that means there was no corresponding record in disasters df, which means county did not have any disasters in the period 1997-1999
df <- df |> mutate(n_disasters.PRE = replace_na(n_disasters.PRE, 0))

# location (categorical)
source("../SCRIPTS/clean_helper.R")
df <- add_regions_and_divisions(df |> rename(CENSUS_STATE_ABBR = state_id))
```

There are some missing values we need to handle.
```{r}
missing_dat <- na_counts_df(df)

missing_counties <- df |> filter(is.na(county)) |> select(geoid_2010)
missing_counties <- missing_counties$geoid_2010
```
From [County Geography Changes 1990 to Present](https://www.cdc.gov/nchs/nvss/bridged_race/county_geography-_changes1990-present.pdf#:~:text=Valdez%2DCordova%20Census%20Area%2C%20Alaska%20(FIPS,Census%20Area%20(FIPS%20Code%3D%2002066).)
- "02261" : Valdez-Cordova Census Area (Alaska). In 2019, was split into two areas: 02063 and 02066
- "02270" : Wade Hampton Census Area (Alaska). In 2015, was renamed and given new FIPS: 02158
- "46113" : Shannon County (South Dakota). In 2015, was renamed and given new FIPS: 46102

```{r}
cols_to_fill <- c("county", "county_full", "CENSUS_STATE_ABBR", "state_name", "lat", "lng", "REGION", "DIVISION")

df[df$geoid_2010 == "02261", cols_to_fill] <- list("Valdez-Cordova",
                                                   "Valdez-Cordova Census Area",
                                                   "AK",
                                                   "Alaska",
                                                   61.34984,
                                                   -145.023141,
                                                   "WEST",
                                                   "PACIFIC")

df[df$geoid_2010 == "02270", cols_to_fill] <- list("Kusilvak",
                                                   "Kusilvak Census Area",
                                                   "AK",
                                                   "Alaska",
                                                   62.1547,
                                                   -163.3812,
                                                   "WEST",
                                                   "PACIFIC")

df[df$geoid_2010 == "46113", cols_to_fill] <- list("Oglala Lakota",
                                                   "Oglala Lakota County",
                                                   "SD",
                                                   "South Dakota",
                                                   43.3355,
                                                   -102.5516,
                                                   "MIDWEST",
                                                   "CENTRAL-NW")
```

From [Bureau of Labor Statistics - New 2024 Connecticut Counties](https://www.bls.gov/cew/classifications/areas/new-2024-connecticut-counties.htm#:~:text=Table_title:%20New%20County%20Codes%20and%20Titles%20Table_content:,Title:%20Northeastern%20Connecticut%20Planning%20Region%2C%20Connecticut%20%7C):
"In 2022, the State of Connecticut requested that The U.S. Census Bureau adopt its nine planning regions as county-equivalents. These nine planning regions replace Connecticut's eight counties used historically... There is no direct mapping between the old and new counties."
- "09001" : Fairfield County
- "09003" : Hartford County
- "09005" : Litchfield County
- "09007" : Middlesex County
- "09009" : New Haven County
- "09011" : New London County
- "09013" : Tolland County
- "09015" : Windham County

```{r}
df[df$geoid_2010 == "09001", cols_to_fill] <- list("Fairfield", "Fairfield County",
                                                   "CT", "Connecticut",
                                                   41.227413, -73.367061,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09003", cols_to_fill] <- list("Hartford", "Hartford County",
                                                   "CT", "Connecticut",
                                                   41.806053, -72.732916,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09005", cols_to_fill] <- list("Litchfield", "Litchfield County",
                                                   "CT", "Connecticut",
                                                   41.791883, -73.235404,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09007", cols_to_fill] <- list("Middlesex", "Middlesex County",
                                                   "CT", "Connecticut",
                                                   41.433003, -72.52278,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09009", cols_to_fill] <- list("New Haven", "New Haven County",
                                                   "CT", "Connecticut",
                                                   41.349717, -72.900204,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09011", cols_to_fill] <- list("New London", "New London County",
                                                   "CT", "Connecticut",
                                                   41.472652, -72.108634,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09013", cols_to_fill] <- list("Tolland", "Tolland County",
                                                   "CT", "Connecticut",
                                                   41.858081, -72.340978,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09015", cols_to_fill] <- list("Windham", "Windham County",
                                                   "CT", "Connecticut",
                                                   41.824999, -71.990702,
                                                   "NORTHEAST", "NEW-ENGLAND")
```

Now, add climate data and handle missing values.
```{r}
# The following counties had a data download failure: 02261, 02270, 11001, 46113, 51678
# I will fix these by hand below
# Starting with precipitation
precipitation <- readRDS("climate_data/all_county_precip_data.rds")

url_pars <- list(scope = "county",
                 parameter = "pcp",
                 timescale = 1,
                 month = 0,
                 begYear = 1990,
                 endYear = 2020,
                 format = "csv")

all_counties <- data.frame(state = character(), geoid = character(), year = integer(), month = integer(), value = numeric())

# 02261 : Valdez-Cordova Census Area (Alaska). In 2019, was split into two areas: 02063 (Chugach Census Area) and 02066 (Copper River Census Area). 
# Will combine the data via a weighted average: https://data.census.gov/profile gives area (in square miles) of US counties
# Chugach (9,575 qmi) + Copper River (24,647 sqmi) = Valdez-Cordova (34,222 sqmi)

url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-063", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

chugach <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dates <- chugach[[1]] # Note these dates, years, and months vectors will be the same for all records!
dat.year <- as.integer(substr(dates, 1, 4))
dat.month <- as.integer(substr(dates, 5, 6))
chugach.values <- as.double(chugach[[2]])

url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-066", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

copper <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
copper.values <- as.double(copper[[2]])

weighted_avg.values <- (0.28 * chugach.values) + (0.72 * copper.values)

all_counties <- rbind(all_counties, 
                      data.frame(state = "Alaska", geoid = "02261", year = dat.year, month = dat.month, val = weighted_avg.values))

# 02270 : Wade Hampton Census Area (Alaska). In 2015, was renamed and given new FIPS: 02158 (Kusilvak Census Area)
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-158", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "Alaska", geoid = "02270", year = dat.year, month = dat.month, val = dat.values))

# 11001 : Washington D.C. is listed as a county in Maryland (MD-511) in the NOAA API
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "MD-511", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "District of Columbia", geoid = "11001", year = dat.year, month = dat.month, val = dat.values))

# 46113 : Shannon County (South Dakota). In 2015, was renamed and given new FIPS: 46102 (Oglala Lakota County)
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "SD-102", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.year <- as.integer(substr(dat[[1]], 1, 4))
dat.month <- as.integer(substr(dat[[1]], 5, 6))
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "South Dakota", geoid = "46113", year = dat.year, month = dat.month, val = dat.values))

# 51678 : Lexington City (VA) - an independent city meaning not part of a county; either drop or impute as average/median values of Virginia. From Wikipedia: "It is the county seat of Rockbridge County, although the two are separate jurisdictions, and is combined with it for statistical purposes by the Bureau of Economic Analysis." - will impute with Rockbridge County values
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "VA-163", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.year <- as.integer(substr(dat[[1]], 1, 4))
dat.month <- as.integer(substr(dat[[1]], 5, 6))
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "Virginia", geoid = "51678", year = dat.year, month = dat.month, val = dat.values))

precipitation <- rbind(precipitation, all_counties)
saveRDS(precipitation, "climate_data/precipitation.rds")
```

```{r}
# Moving on to average temperature
rm(list=ls())
average_temp <- readRDS("climate_data/all_county_avg_temp_data.rds")

url_pars <- list(scope = "county",
                 parameter = "tavg",
                 timescale = 1,
                 month = 0,
                 begYear = 1990,
                 endYear = 2020,
                 format = "csv")

all_counties <- data.frame(state = character(), geoid = character(), year = integer(), month = integer(), value = numeric())

# 02261 : Valdez-Cordova Census Area (Alaska). In 2019, was split into two areas: 02063 (Chugach Census Area) and 02066 (Copper River Census Area). 
# Will combine the data via a weighted average: https://data.census.gov/profile gives area (in square miles) of US counties
# Chugach (9,575 qmi) + Copper River (24,647 sqmi) = Valdez-Cordova (34,222 sqmi)

url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-063", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

chugach <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dates <- chugach[[1]] # Note these dates, years, and months vectors will be the same for all records!
dat.year <- as.integer(substr(dates, 1, 4))
dat.month <- as.integer(substr(dates, 5, 6))
chugach.values <- as.double(chugach[[2]])

url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-066", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

copper <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
copper.values <- as.double(copper[[2]])

weighted_avg.values <- (0.28 * chugach.values) + (0.72 * copper.values)

all_counties <- rbind(all_counties, 
                      data.frame(state = "Alaska", geoid = "02261", year = dat.year, month = dat.month, val = weighted_avg.values))

# 02270 : Wade Hampton Census Area (Alaska). In 2015, was renamed and given new FIPS: 02158 (Kusilvak Census Area)
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-158", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "Alaska", geoid = "02270", year = dat.year, month = dat.month, val = dat.values))

# 11001 : Washington D.C. is listed as a county in Maryland (MD-511) in the NOAA API
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "MD-511", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "District of Columbia", geoid = "11001", year = dat.year, month = dat.month, val = dat.values))

# 46113 : Shannon County (South Dakota). In 2015, was renamed and given new FIPS: 46102 (Oglala Lakota County)
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "SD-102", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.year <- as.integer(substr(dat[[1]], 1, 4))
dat.month <- as.integer(substr(dat[[1]], 5, 6))
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "South Dakota", geoid = "46113", year = dat.year, month = dat.month, val = dat.values))

# 51678 : Lexington City (VA) - an independent city meaning not part of a county; either drop or impute as average/median values of Virginia. From Wikipedia: "It is the county seat of Rockbridge County, although the two are separate jurisdictions, and is combined with it for statistical purposes by the Bureau of Economic Analysis." - will impute with Rockbridge County values
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "VA-163", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.year <- as.integer(substr(dat[[1]], 1, 4))
dat.month <- as.integer(substr(dat[[1]], 5, 6))
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "Virginia", geoid = "51678", year = dat.year, month = dat.month, val = dat.values))

average_temp <- rbind(average_temp, all_counties)
saveRDS(average_temp, "climate_data/average_temperature.rds")
```

```{r}
# climate data
precipitation <- readRDS("climate_data/precipitation.rds") |> filter(year >= YEAR-3 & year < YEAR) 

df <- merge(df, precipitation |> group_by(geoid) |> summarize(med_precip = median(val, na.rm = TRUE), mean_precip = mean(val, na.rm = TRUE)),
      by.x = "geoid_2010", by.y = "geoid", all.x = TRUE)

avg_temp <- readRDS("climate_data/average_temperature.rds") |> filter(year >= YEAR-3 & year < YEAR)

df <- merge(df, avg_temp |> group_by(geoid) |> summarize(med_avg_temp = median(val, na.rm = TRUE), mean_avg_temp = mean(val, na.rm = TRUE)),
      by.x = "geoid_2010", by.y = "geoid", all.x = TRUE)

missing_dat <- na_counts_df(df)
```

Let's add in a few treatment options.
```{r}
df <- df |> mutate(treat.n_disasters = as.numeric(n_disasters > 0),
                   treat.median_log_damage = as.numeric(LOG_TotPropDmg2015 > 11), # about 60K of property damage
                   treat.cat = case_when(
                         LOG_TotPropDmg2015 <= 0 ~ "none",
                         LOG_TotPropDmg2015 <= 11 ~ "small", # up to ~60K
                         LOG_TotPropDmg2015 <= 13 ~ "med", # up to ~442K
                         LOG_TotPropDmg2015 <= 22 ~ "large", # up to 3.5 trillion
                   ))

df |> select(geoid_2010, LOG_TotPropDmg2015, n_disasters, treat.n_disasters, treat.median_log_damage, treat.cat)
```


## Pre-matching Balance
Now we can assess covariate balance prior to matching.
```{r}
library(MatchIt)

m.initial <- matchit(treat.median_log_damage ~ lat*lng + med_precip + med_avg_temp,
                     data = df,
                     method = NULL,
                     distance = "glm")

summary(m.initial)
```

# Scratchwork
```{r}
create_adj_list <- function(panel_data, dist_cutoff){
      coords <- as.data.frame(panel_data |> select(geoid_2010, lng, lat) |> unique())
      county_IDs <- coords$geoid_2010
      
      # Border-based adj list
      census_df <- read.delim("counties_data/county_adjacency2010.txt",
                       header = FALSE,
                       sep = "\t",
                       col.names = c("county_name", "county_geoid",
                                     "neighbor_name", "neighbor_geoid"),
                       colClasses = c("character", "character", "character", "character"),
                       stringsAsFactors = FALSE)
      
      census_df <- census_df |>
            mutate(across(everything(), \(x) na_if(x, y = "")))
      
      census_df <- census_df |>
            fill(county_name, county_geoid, .direction = "down")
      
      census_df <- census_df[census_df$county_geoid != census_df$neighbor_geoid, ]
      
      census_df <- census_df |> filter((county_geoid %in% county_IDs) & (neighbor_geoid %in% county_IDs))
      
      census_adj_list <- split(census_df$neighbor_geoid, census_df$county_geoid)
      
      missing <- setdiff(county_IDs, names(census_adj_list))
      for (m in missing) {
            census_adj_list[[m]] <- character(0)
      }
      
      # Distance based adj list
      D <- distm(coords |> select(-geoid_2010), fun = distHaversine) / 1609.344 #(meters to miles conversion)
      dimnames(D) <- list(county_IDs, county_IDs)
      
      dist_cutoff <- 25
      idx <- which(D <= dist_cutoff & D > 0, arr.ind = TRUE) # edge list (directed: (i,j) and (j,i) both appear)
      
      edges <- data.frame(county1 = county_IDs[idx[, 1]], 
                          county2 = county_IDs[idx[, 2]], 
                          stringsAsFactors = FALSE)
      
      dist_adj_list <- split(edges$county2, edges$county1)
      
      # If there are missing counties, that means that they had no neighboring counties within the cutoff distance -- so can list an empty vector for them
      missing <- setdiff(unique(df$geoid_2010), names(dist_adj_list))
      for (m in missing) {
            dist_adj_list[[m]] <- character(0)
      }
      
      # for every county, new adj_list is union of the two
      adj_union <- setNames(lapply(names(dist_adj_list), function(n) union(dist_adj_list[[n]], census_adj_list[[n]])),
                            names(dist_adj_list))
      
      return(list("union" = adj_union, "border" = census_adj_list, "dist" = dist_adj_list))
}

add_exposed <- function(panel_data, adj_list){
      panel_data$treat.FEMA.adj <- panel_data$treat.FEMA # initialize adjusted treatment with original
      years <- unique(panel_data$TAX_YEAR)
      
      for (yr in years) {
            idx <- which(panel_data$TAX_YEAR == yr) # Subset indices for this year
            treated <- panel_data$geoid_2010[idx][panel_data$treat.FEMA[idx] == 1]
            
            if (length(treated) == 0) next # If no treated counties, skip
            
            neighbors <- unique(unlist(adj_list[treated])) # Get neighbors of treated counties
            
            untreated_idx <- idx[panel_data$treat.FEMA[idx] == 0] # Identify untreated counties in this year
            
            exposed <- panel_data$geoid_2010[untreated_idx] %in% neighbors # Among untreated, find those exposed
            
            panel_data$treat.FEMA.adj[untreated_idx][exposed] <- NA # Set exposed controls to NA
      }
      
      # per year I want to see the counts of treated, control and missing for treat.FEMA.adj
      per_year_summary <- panel_data |> group_by(TAX_YEAR) |>
            summarise(
                  n_1 = sum(treat.FEMA.adj == 1, na.rm = TRUE),
                  n_0 = sum(treat.FEMA.adj == 0, na.rm = TRUE),
                  n_NA = sum(is.na(treat.FEMA.adj)),
                  n = n())
      
      return(list("df" = panel_data, "summary" = per_year_summary))
}

library(tidyr)
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)
library(geosphere)

df <- readRDS("county_long.rds") |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 50)
resUNION <- add_exposed(panel_data = df, adj_list = adj_lists$union)
resBORDER <- add_exposed(panel_data = df, adj_list = adj_lists$border)
resDIST <- add_exposed(panel_data = df, adj_list = adj_lists$dist)

resUNION$summary
resBORDER$summary
resDIST$summary
```

```{r}
ntee <- colnames(df)[10:22]
paste0("I(lag(", ntee, ", 1:1))", collapse = " + ")
```

```{r}
add_exposedOLD <- function(dist_data, dist_cutoff, panel_data){
      # Filter by distance cutoff
      edges <- dist_data[dist_data$mi_to_county <= dist_cutoff, c("county1", "county2")]
      
      # Build adjacency list
      adj_list <- split(edges$county2, edges$county1) 
      
      # If there are missing counties, that means that they had no neighboring counties within the cutoff distance -- so can list an empty vector for them
      missing <- setdiff(unique(panel_data$geoid_2010), names(adj_list))
      for (m in missing) {
            adj_list[[m]] <- character(0)
      }
      
      panel_data$treat.FEMA.adj <- panel_data$treat.FEMA # initialize adjusted treatment with original
      years <- unique(panel_data$TAX_YEAR)
      
      for (yr in years) {
            idx <- which(panel_data$TAX_YEAR == yr) # Subset indices for this year
            treated <- panel_data$geoid_2010[idx][panel_data$treat.FEMA[idx] == 1]
            
            if (length(treated) == 0) next # If no treated counties, skip
            
            neighbors <- unique(unlist(adj_list[treated])) # Get neighbors of treated counties
            
            untreated_idx <- idx[panel_data$treat.FEMA[idx] == 0] # Identify untreated counties in this year
            
            exposed <- panel_data$geoid_2010[untreated_idx] %in% neighbors # Among untreated, find those exposed
            
            panel_data$treat.FEMA.adj[untreated_idx][exposed] <- NA # Set exposed controls to NA
      }
      
      # per year I want to see the counts of treated, control and missing for treat.FEMA.adj
      per_year_summary <- panel_data |> group_by(TAX_YEAR) |>
            summarise(
                  n_1 = sum(treat.FEMA.adj == 1, na.rm = TRUE),
                  n_0 = sum(treat.FEMA.adj == 0, na.rm = TRUE),
                  n_NA = sum(is.na(treat.FEMA.adj)),
                  n = n())
      
      return(list("df" = panel_data, "summary" = per_year_summary))
}
```

```{r}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)
library(fields)

df <- readRDS("county_long.rds") |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

nber_data <- read_dta("counties_data/sf12010countydistancemiles.dta")

res25 <- add_exposedOLD(dist_data = nber_data, dist_cutoff = 25, panel_data = df)
res50 <- add_exposedOLD(dist_data = nber_data, dist_cutoff = 50, panel_data = df)
res100 <- add_exposedOLD(dist_data = nber_data, dist_cutoff = 100, panel_data = df)

res25$summary
res50$summary
res100$summary
```

```{r}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)

df <- readRDS("county_long.rds") |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

res25e <- add_exposed(dist_cutoff = 25, panel_data = df)
res50e <- add_exposed(dist_cutoff = 50, panel_data = df)
res100e <- add_exposed(dist_cutoff = 100, panel_data = df)

res25e$summary
res50e$summary
res100e$summary
```


```{r}
# panel_data = df
# dist_cutoff <- 25
library(geosphere)
add_exposed <- function(dist_cutoff, panel_data){
      coords <- as.data.frame(panel_data |> select(geoid_2010, lng, lat) |> unique())
      county_IDs <- coords$geoid_2010
      
      D <- distm(coords |> select(-geoid_2010), fun = distHaversine) / 1609.344 #(meters to miles conversion)
      #D <- rdist.earth(coords |> select(-geoid_2010), miles = TRUE)
      dimnames(D) <- list(county_IDs, county_IDs)
      #diag(D) <- 0
      
      idx <- which(D <= dist_cutoff & D > 0, arr.ind = TRUE) # edge list (directed: (i,j) and (j,i) both appear)
      
      edges <- data.frame(county1 = county_IDs[idx[, 1]], 
                          county2 = county_IDs[idx[, 2]], 
                          stringsAsFactors = FALSE)
      
      adj_list <- split(edges$county2, edges$county1)
      
      # If there are missing counties, that means that they had no neighboring counties within the cutoff distance -- so can list an empty vector for them
      missing <- setdiff(unique(panel_data$geoid_2010), names(adj_list))
      for (m in missing) {
            adj_list[[m]] <- character(0)
      }
      
      panel_data$treat.FEMA.adj <- panel_data$treat.FEMA # initialize adjusted treatment with original
      years <- unique(panel_data$TAX_YEAR)
      
      for (yr in years) {
            idx <- which(panel_data$TAX_YEAR == yr) # Subset indices for this year
            treated <- panel_data$geoid_2010[idx][panel_data$treat.FEMA[idx] == 1]
            
            if (length(treated) == 0) next # If no treated counties, skip
            
            neighbors <- unique(unlist(adj_list[treated])) # Get neighbors of treated counties
            
            untreated_idx <- idx[panel_data$treat.FEMA[idx] == 0] # Identify untreated counties in this year
            
            exposed <- panel_data$geoid_2010[untreated_idx] %in% neighbors # Among untreated, find those exposed
            
            panel_data$treat.FEMA.adj[untreated_idx][exposed] <- NA # Set exposed controls to NA
      }
      
      # per year I want to see the counts of treated, control and missing for treat.FEMA.adj
      per_year_summary <- panel_data |> group_by(TAX_YEAR) |>
            summarise(
                  n_1 = sum(treat.FEMA.adj == 1, na.rm = TRUE),
                  n_0 = sum(treat.FEMA.adj == 0, na.rm = TRUE),
                  n_NA = sum(is.na(treat.FEMA.adj)),
                  n = n())
      
      return(list("df" = panel_data, "summary" = per_year_summary))
}
```




```{r}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(fastDummies)

df <- readRDS("mega.rds") # 1773 orgs have "00000" as geoid; 86,752 are missing NTEE code (~6.74%) -- 1,285,846 total orgs

# Split up NTEEV2 into 14 dummy variables (12 categories, 1 unknown category, 1 missing category)
ntee <- df |> 
      select(EIN2, NTEEV2, TAX_YEAR, county.census.geoid) |>
      filter(county.census.geoid != "00000") |>
      dummy_cols(select_columns = c("NTEEV2"), ignore_na = FALSE, remove_selected_columns = TRUE)

# Add aggregate the totals of each NTEEV2 category by county and year
ntee <- ntee |> select(-EIN2) |>
      group_by(county.census.geoid, TAX_YEAR) |>
      summarise(across(everything(), ~sum(.x, na.rm = TRUE))) |>
      ungroup()

# Add in total number of organizations and proportion missing column
ntee <- ntee |> mutate(NUM_ORGS = rowSums(ntee |> select(-county.census.geoid, -TAX_YEAR)),
                       NTEE_NA_PROP = NTEEV2_NA/NUM_ORGS)

# Within each county,year pair, I want to keep track of the proportion of records with missing financial data
# I also want to compute the total revenue, expenses, and assets across that county in that year
# For now, we are dropping records that listed the invalid county code "00000" (1773 orgs)
df <- df |> select(TAX_YEAR, TOT_REV, TOT_EXP, TOT_ASSET, county.census.geoid) |>
      rename(REV = TOT_REV, EXP = TOT_EXP, ASSET = TOT_ASSET) |>
      group_by(county.census.geoid, TAX_YEAR) |> 
      summarize(across(where(is.numeric), ~sum(is.na(.x))/n(), .names = "{.col}_prop_miss"),
                across(any_of(c("REV", "EXP", "ASSET")), ~sum(.x, na.rm = TRUE), .names = "TOT_{.col}"),
                .groups = "drop") |>
      ungroup() |>
      filter(county.census.geoid != "00000")

df <- merge(df, ntee, by = c("county.census.geoid", "TAX_YEAR")) 

# IDs for two counties were changed after 2010. Some datasets use latest ID, others use 2010 ID
df <- df |> mutate(geoid_2010 = county.census.geoid)
df$geoid_2010[df$county.census.geoid == "02158"] <- "02270"
df$geoid_2010[df$county.census.geoid == "46102"] <- "46113"
```


```{r}
# File: perform_refinement.R
# glm(reformulate(response = treatment, termlabels = colnames(pooled)[-c(1:3)]), family = binomial(link = "logit"), data = pooled)
# CBPS::CBPS(reformulate(response = treatment, termlabels = colnames(pooled)[-c(1:3)]), family = binomial(link = "logit"), data = pooled)

library(fields)
library(MASS)

## Nonprofit data
set.seed(2727)
county_subset <- sample.int(n = max(df$CountyID), size = 500, replace = FALSE)
df_sub <- df |> filter(CountyID %in% county_subset)

x <- cbind(df_sub$lat, df_sub$lng) # inputs: latitude and longitude

d <- rdist(x)
K <- Matern(d, smoothness=.5) # for smoothness paramter, should we do some sort of optimization?
E <- eigen(K)

y <- as.matrix(df_sub$treat.FEMA)
x_s <- cbind(x[,1] * E$vectors[,1:45], x[,2] * E$vectors[,1:45]) # would have to play around with this number

xs_vars <- paste0("x_s", seq_len(ncol(x_s)))
colnames(x_s) <- xs_vars
df_sub <- cbind(df_sub, x_s)

# c("bachelors_perc", "med_household_income_adj", "white_perc",  "pop_estimate", "med_precip", "med_avg_temp", "TOT_REV")
formula_string <- paste(
  "y ~ bachelors_perc + med_household_income_adj + white_perc + pop_estimate + med_precip + med_avg_temp +",
  paste(xs_vars, collapse = " + "))

ls <- glm(as.formula(formula_string), data = df_sub, family = binomial(link = "logit"))
summary(ls)

panel_df <- PanelData(panel.data = df_sub,
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA",
                      outcome = "TOT_REV")


formula_L1 <- paste(
  "~ I(lag(bachelors_perc, 0:1)) + I(lag(med_household_income_adj, 0:1)) + I(lag(white_perc, 0:1)) + I(lag(pop_estimate, 0:1)) + med_precip + med_avg_temp +",
  paste(xs_vars, collapse = " + "))

formula_L3 <- paste(
  "~ I(lag(bachelors_perc, 0:3)) + I(lag(med_household_income_adj, 0:3)) + I(lag(white_perc, 0:3)) + I(lag(pop_estimate, 0:3)) + med_precip + med_avg_temp +",
  paste(xs_vars, collapse = " + "))

formula_L5 <- paste(
  "~ I(lag(bachelors_perc, 0:5)) + I(lag(med_household_income_adj, 0:5)) + I(lag(white_perc, 0:5)) + I(lag(pop_estimate, 0:5)) + med_precip + med_avg_temp +",
  paste(xs_vars, collapse = " + "))

formula_string <- paste(
  "y ~ bachelors_perc + med_household_income_adj + white_perc + pop_estimate + med_precip + med_avg_temp +",
  paste(colnames(df)[10:24], collapse = " + "))
```

