---
title: "Analysis"
author: "Mayleen"
date: "2026-01-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# PanelMatch - No Interference
## Setup
For matching (just tot rev)
```{r, message = F}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(fastDummies)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# These are some socioeconomic covariates we may evaluate balance for but need to convert from string to num
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population")
df <- df |> mutate(across(all_of(covs_to_check), as.numeric)) # make sure these are numeric

# Add REGION dummy columns
df <- df |> dummy_cols(select_columns = c("REGION"), ignore_na = FALSE, remove_selected_columns = TRUE, remove_first_dummy = FALSE)

# These are the variables we will adjust for
adj_set <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "lat", "lng")

# Other variables
ntee_vars <- grep("^NTEEV2_", names(df), value = TRUE)
region_vars <- grep("^REGION_", names(df), value = TRUE)

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA_nDis"), adj_set, ntee_vars, region_vars)))

# 1859 is New York County
df$treat.FEMA_nDis[((df$CountyID == 1859) & (df$TAX_YEAR == 2001))] <- NA

# log transform
#saveRDS(df |> filter(TOT_REV < 0 | TOT_ASSET < 0), "misc_analysis/no_intV1_negative.rds")
df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
                                       .default = TOT_REV),
                   TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
                                       .default = TOT_ASSET))
df <- df |> mutate(across(c("TOT_REV", "TOT_ASSET"), log1p)) #log1p adds 1 first to avoid 0s

# Adding spatially varying coefficients
source("matches_helper.R")
df <- add_SVC(df, num_E_vec = 100)
df$lat <- NULL
df$lng <- NULL

df$geoid_2010 <- NULL

# https://stackoverflow.com/questions/74815244/panelmatch-r-errors-please-convert-time-id-to-consecutive-integers-and-pleas
panel_df <- PanelData(panel.data = as.data.frame(df),
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA_nDis",
                      outcome = "TOT_REV")
# rm(df)

# saveRDS(panel_df, "panel_data/no_int_panelV1.rds") # No SVC
# saveRDS(panel_df, "panel_data/no_int_panelV2.rds") # Yes SVC

#write.csv(df |> select(TAX_YEAR, FEMA_threshADJ) |> unique(), "updated_threshold.csv", row.names = FALSE)
```

```{r}
sum(is.na(panel_df$treat.FEMA)) # county_FEMA_indicator <- 3.78 #2019

treatVarPlot <- DisplayTreatment(panel.data = panel_df, legend.position = "none", xlab = "TAX_YEAR", ylab = "County Code", hide.x.tick.label = TRUE, hide.y.tick.label = TRUE, dense.plot = TRUE)

treatVarPlot
#ggsave("treatment_variation_plot.png", plot = treatVarPlot, width = 7.5, height = 10, units = "in")
```

## Matching Methods - Config 1, No SVC
### Lag = 3
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/no_int_panelV1.rds")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/no_svc/no_int_panelV1/config1/"
folder_path_balance <- "balances/no_svc/no_int_panelV1/config1/NSEW_"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
summary(matches_list$none)
summary(matches_list$ps.match)
summary(matches_list$CBPS.match)
summary(matches_list$ps.weight)
summary(matches_list$CBPS.weight)

plot(matches_list$ps.match)
#print(matches_list$none)
```


Default Matching Weights Plots
```{r}
# # panel_df <- readRDS("panel_data/no_int_panelV1.rds")
# extracted_matches <- readRDS(paste0(folder_path_match, "extracted_", save_name, ".rds"))
# # plot(extracted_matches$CBPS.weight, panel.data = panel_df, include.missing = FALSE)
# names(extracted_matches$CBPS.weight)
# extracted_matches$CBPS.weight$`1.1995`
# extracted_matches$CBPS.match$`1.1995`
# weights(extracted_matches$CBPS.weight)
# weights(extracted_matches$ps.match)
```


Default Scatter Plots
```{r}
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```

My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.15,0.15), "unrefined" = c(-0.15, 0.05), "all" = c(-0.15, 0.15))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

# p_none$base
# # p_ps.match$base_unrefined
# p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
# p_cbps.weight$base
# 
# p_none$ntee
# # p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
# p_cbps.weight$ntee
# 
# p_none$loc
# # p_ps.match$loc_unrefined
# p_ps.match$loc
# p_cbps.match$loc
# p_ps.weight$loc
# p_cbps.weight$loc
# 
# p_none$all
# # p_ps.match$all_unrefined
# p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
# p_cbps.weight$all
```

```{r}
quantile(signif(abs(bal_none$value[bal_cbps.match$unrefined == FALSE]), 3))
quantile(signif(abs(bal_cbps.match$value[bal_cbps.match$unrefined == FALSE]), 3))
```



### Lag = 5
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/no_int_panelV1.rds")

# Set up the folder path and saved object names
L = 5 # lag, options: 1, 3, 5
folder_path_match <- "matches/no_svc/no_int_panelV1/config1/"
folder_path_balance <- "balances/no_svc/no_int_panelV1/config1/NSEW_"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
# summary(matches_list$none)
# summary(matches_list$ps.match)
# summary(matches_list$CBPS.match)
# summary(matches_list$ps.weight)
# summary(matches_list$CBPS.weight)
# 
# plot(matches_list$none)
```


Default Matching Weights Plots
```{r}
# plot(extract(matches_list$CBPS.weight), panel.data = panel_df)
```


Default Scatter Plots
```{r}
# plot(balances_list$none, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE, 
#      main = "Standardized Mean Difference of Covariates (No Matching)")
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```


My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.15,0.2), "unrefined" = c(-0.15, 0.05), "all" = c(-0.2, 0.2))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

# p_none$base
# # p_ps.match$base_unrefined
# p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
# p_cbps.weight$base
# 
# p_none$ntee
# # p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
# p_cbps.weight$ntee
# 
p_none$loc
# p_ps.match$loc_unrefined
p_ps.match$loc
p_cbps.match$loc
p_ps.weight$loc
p_cbps.weight$loc
# 
# p_none$all
# # p_ps.match$all_unrefined
# p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
# p_cbps.weight$all
```

```{r}
# print("No Matching")
# quantile(signif(abs(bal_none$value[bal_none$unrefined == FALSE]), 3))
# 
# print("PS Matching")
# quantile(signif(abs(bal_ps.match$value[bal_ps.match$unrefined == FALSE]), 3))
# 
# print("CBPS Matching")
# quantile(signif(abs(bal_cbps.match$value[bal_cbps.match$unrefined == FALSE]), 3))
# 
# print("PS Weighting")
# quantile(signif(abs(bal_ps.weight$value[bal_ps.weight$unrefined == FALSE]), 4), probs = seq(0,1,0.1))
# 
# print("CBPS Weighting")
# quantile(signif(abs(bal_cbps.weight$value[bal_cbps.weight$unrefined == FALSE]), 4), probs = seq(0,1,0.1))

quantile(signif(abs(bal_none$value[bal_none$unrefined == FALSE]), 4))
quantile(signif(abs(bal_cbps.weight$value[bal_cbps.weight$unrefined == FALSE]), 4))
```



## Matching Methods - Config 1, With SVC
### Lag = 3
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/no_int_panelV2.rds")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/no_int_panelV2/config1/"
folder_path_balance <- "balances/with_svc/no_int_panelV2/config1/NSEW_"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
summary(matches_list$none)
summary(matches_list$ps.match)
summary(matches_list$CBPS.match)
summary(matches_list$ps.weight)
summary(matches_list$CBPS.weight)

plot(matches_list$none)
```


Default Matching Weights Plots
```{r}
# plot(extract(matches_list$CBPS.weight), panel.data = panel_df)
```


Default Scatter Plots
```{r}
# plot(balances_list$none, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE, 
#      main = "Standardized Mean Difference of Covariates (No Matching)")
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```

My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.15,0.15), "unrefined" = c(-0.15, 0.05), "all" = c(-0.15, 0.15))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

# p_none$base
# # p_ps.match$base_unrefined
# p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
# p_cbps.weight$base

# p_none$ntee
# # p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
# p_cbps.weight$ntee

p_none$loc
# p_ps.match$loc_unrefined
p_ps.match$loc
p_cbps.match$loc
p_ps.weight$loc
p_cbps.weight$loc
# 
# p_none$all
# # p_ps.match$all_unrefined
# p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
# p_cbps.weight$all
```

```{r}
quantile(signif(abs(bal_cbps.match$value[bal_cbps.match$unrefined == FALSE]), 3))
```

### Lag = 5
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/no_int_panelV2.rds")


# Set up the folder path and saved object names
L = 5 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/no_int_panelV2/config1/"
folder_path_balance <- "balances/with_svc/no_int_panelV2/config1/NSEW_"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
# summary(matches_list$none)
# summary(matches_list$ps.match)
# summary(matches_list$CBPS.match)
# summary(matches_list$ps.weight)
# summary(matches_list$CBPS.weight)
# 
# plot(matches_list$none)
# plot(matches_list$ps.match)
```


Default Matching Weights Plots
```{r}
# plot(extract(matches_list$CBPS.weight), panel.data = panel_df)
```


Default Scatter Plots
```{r}
# plot(balances_list$none, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE, 
#      main = "Standardized Mean Difference of Covariates (No Matching)")
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```


My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.15,0.2), "unrefined" = c(-0.15, 0.05), "all" = c(-0.2, 0.2))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

# p_none$base
# # p_ps.match$base_unrefined
# p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
# p_cbps.weight$base
# 
# p_none$ntee
# # p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
# p_cbps.weight$ntee
# 
p_none$loc
# p_ps.match$loc_unrefined
p_ps.match$loc
p_cbps.match$loc
p_ps.weight$loc
p_cbps.weight$loc
# 
# p_none$all
# # p_ps.match$all_unrefined
# p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
# p_cbps.weight$all
```

```{r}
# print("No Matching")
# quantile(signif(abs(bal_none$value[bal_none$unrefined == FALSE]), 3))
# 
# print("PS Matching")
# quantile(signif(abs(bal_ps.match$value[bal_ps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
# 
# print("CBPS Matching")
# quantile(signif(abs(bal_cbps.match$value[bal_cbps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))

# print("PS Weighting")
# quantile(signif(abs(bal_ps.weight$value[bal_ps.weight$unrefined == FALSE]), 4), probs = seq(0,1,0.1))

# print("CBPS Weighting")
# quantile(signif(abs(bal_cbps.weight$value[bal_cbps.weight$unrefined == FALSE]), 4), probs = seq(0,1,0.1))

quantile(signif(abs(bal_none$value[bal_none$unrefined == FALSE]), 3))
quantile(signif(abs(bal_ps.weight$value[bal_ps.weight$unrefined == FALSE]), 3))
```


## Estimates
Full (all outcomes)
```{r, message = F}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(fastDummies)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

outcome_vars <- grep("^REV_", names(df), value = TRUE)

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "treat.FEMA_nDis", "TOT_REV", "TOT_ASSET"), outcome_vars)))

# log transform
# saveRDS(df |> filter(if_any(all_of(c("TOT_REV", "TOT_ASSET", outcome_vars)), ~ .x < 0)), "misc_analysis/no_intV1_negative_stratified.rds")
df <- df |> mutate(across(c("TOT_REV", outcome_vars), ~ if_else(.x < 0, 0, .x)))
df <- df |> mutate(across(c("TOT_REV", outcome_vars), log1p)) #log1p adds 1 first to avoid 0s

df$treat.FEMA_nDis[((df$CountyID == 1859) & (df$TAX_YEAR == 2001))] <- NA
# https://stackoverflow.com/questions/74815244/panelmatch-r-errors-please-convert-time-id-to-consecutive-integers-and-pleas

panel_df <- PanelData(panel.data = as.data.frame(df),
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA_nDis",
                      outcome = "TOT_REV")
# 
# panel_df2 <- as.data.frame(readRDS("panel_data/no_int_panelV2.rds"))

# all.equal(as.data.frame(panel_df) |> select("CountyID", "TAX_YEAR", "treat.FEMA_nDis", "TOT_REV"), 
#           panel_df2 |> select("CountyID", "TAX_YEAR", "treat.FEMA_nDis", "TOT_REV"))

rm(df)
```
### Lag = 3
Stratified by NTEE category
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/no_int_panelV2/config1/"
save_name <- paste0("Lag", L)
estimates_save_name <- "estimates/with_svc_no_intV2_con1_"
match_method <- "CBPS.match"

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))
m.out <- matches_list[[match_method]]

# panel_df <- readRDS("panel_data/no_int_panelV2.rds")

# Need to compute the estimates for each NTEE category
outcome_vars <- c("TOT_REV", "REV_ART", "REV_EDU", "REV_HMS", "REV_PSB", "REV_ENV", "REV_HEL", "REV_HOS", "REV_UNI") # REV_MMB, REV_REL, REV_ENV, REV_IFA, REV_UNU, REV_NA
```

```{r}
res <- list()
for (x in outcome_vars){
      print(x)
      attr(panel_df, "outcome") <- x
      attr(m.out, "outcome") <- x
      attr(m.out, "outcome.var") <- x
      
      res[[x]] <- PanelEstimate(sets = m.out,
                     panel.data = panel_df,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6)
}

for (x in outcome_vars){
      file_name <- paste0("estimates/", x, ".png")
      print(file_name)
      png(file_name)
      plot(res[[x]])
      dev.off()
}

file_name <- paste0(estimates_save_name, save_name, ".xlsx")
print(file_name)
estimate_summaries <- lapply(res, summary)
estimate_summaries <- lapply(estimate_summaries, as.data.frame)
write_xlsx(estimate_summaries, file_name)
```


Just total revenue
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/no_int_panelV2/config1/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

panel_df <- readRDS("panel_data/no_int_panelV2.rds")

# Winner: CBPS Matching with SVC Coefficients
res <- PanelEstimate(sets = matches_list$CBPS.match,
                     panel.data = panel_df,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6)
plot(res)
summary(res)
```
Misc Analysis - Distribution of Matched Set Treatment Effects
```{r}
ind.res <- get_set_treatment_effects(pm.obj = matches_list$CBPS.match,
                                     panel.data = panel_df,
                                     lead = 1:5)

hist(ind.res[[1]], main = "Distribution of Matched Set Treatment Effects\n F = 1",
     xlab = "Treatment Effect",
     ylim = c(0,100))

hist(ind.res[[3]], main = "Distribution of Matched Set Treatment Effects\n F = 3",
     xlab = "Treatment Effect",
     ylim = c(0,100))

hist(ind.res[[5]], main = "Distribution of Matched Set Treatment Effects\n F = 5",
     xlab = "Treatment Effect",
     ylim = c(0,100))
```

Pooled estimates
```{r}
# Winner: CBPS Matching with SVC Coefficients
res.pooled <- PanelEstimate(sets = matches_list$CBPS.match,
                     panel.data = panel_df,
                     se.method = "bootstrap",
                     pooled = TRUE,
                     parallel = TRUE,
                     num.cores = 6)
plot(res.pooled)
```

Different SE computation
```{r}
# Winner: CBPS Matching with SVC Coefficients
# res.indep <- PanelEstimate(sets = matches_list$CBPS.match,
#                      panel.data = panel_df,
#                      se.method = "conditional")
# plot(res.indep)
```

## Choosing the number of SVC
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(fields)
library(MASS)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# These are some socioeconomic covariates we may evaluate balance for but need to convert from string to num
covs_to_check <- c("white_perc", "bachelors_perc", "poverty_perc", "unemployment", "med_family_income_adj", "med_household_income_adj", "total_population")
df <- df |> mutate(across(all_of(covs_to_check), as.numeric)) # make sure these are numeric

# These are the variables we will assess balance for
se_vars <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population")
other_vars <- c("TOT_REV", "TOT_ASSET")
ntee_vars <- c("NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA")

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA_nDis", "lat", "lng"), se_vars, other_vars, ntee_vars)))

df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
                                       .default = TOT_REV),
                   TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
                                       .default = TOT_ASSET))
df <- df |> mutate(across(c("TOT_REV", "TOT_ASSET"), log1p)) #log1p adds 1 first to avoid 0s

library(data.table)
df <- arrange(df, TAX_YEAR)
df.lag <- shift(df[,7:ncol(df)], n = 1:3, give.names = TRUE)
df <- bind_cols(df, df.lag)
rm(df.lag)

df <- df |> filter(TAX_YEAR >= 1994)

df_sub <- df |> dplyr::select(CountyID, lat, lng) |> unique()

x <- cbind(df_sub$lat, df_sub$lng) # inputs: latitude and longitude

# SVC
d <- rdist(x)
K <- Matern(d, smoothness=.5) # for smoothness paramter, should we do some sort of optimization?
E <- eigen(K)

options(repr.plot.width = 6, repr.plot.height = 4)
lls = list()
ls_list = list()
for (num_svc in seq(10,100,10)){
      curr_name <- as.character(num_svc*2)
      x_s <- cbind(x[,1] * E$vectors[,1:num_svc], x[,2] * E$vectors[,1:num_svc]) # would have to play around with this number
      
      # add SVC back into panel data
      xs_vars <- paste0("x_s", seq_len(ncol(x_s)))
      colnames(x_s) <- xs_vars
      df_sub.temp <- cbind(df_sub, x_s)
      
      df.temp <- merge(df, df_sub.temp |> dplyr::select(-lat, -lng), by = "CountyID", all = TRUE)

      upper_formula <- paste(
        "treat.FEMA_nDis ~ TAX_YEAR + bachelors_perc + med_household_income_adj + white_perc + total_population +",
        paste0(do.call(paste0, expand.grid(paste0(se_vars, "_lag_"), 1:3)), collapse = " + "), "+",
        paste0(do.call(paste0, expand.grid(paste0(ntee_vars, "_lag_"), 1:3)), collapse = " + "), "+",
        paste0(do.call(paste0, expand.grid(paste0(other_vars, "_lag_"), 1:3)), collapse = " + "), "+",
        paste(xs_vars, collapse = " + "))
      
      # regression
      ls_list[[curr_name]] <- glm(as.formula(upper_formula), data = df.temp, family = binomial(link = "logit"))
      #summary(ls)

      yweight <- predict(ls_list[[curr_name]], type="response")
      png(paste0("svc_plots/no_interference/svc", num_svc*2, "_Mar19.png"))
      plot(sort(yweight), main = paste("# SVC =", num_svc*2), cex = 0.10)
      dev.off()
} # winner was num_svc = 30 on March 10, but on March 19 200 is the winner by a landslide

library(lmtest)
my_test <- lrtest(ls_list[["100"]], ls_list[["200"]])
my_test

AIC(ls_list[["100"]], ls_list[["200"]])
BIC(ls_list[["100"]], ls_list[["200"]])

y_pred <- predict(ls_list[["200"]], df.temp, type = "response")
quantile(y_pred, na.rm = TRUE, probs = seq(0,1,0.1)) # all probs between 0.02 and 0.95
# sum(near(y_pred, 1), na.rm = TRUE)
# sort(y_pred, decreasing = FALSE) 
```

# PanelMatch - Interference
## Setup
```{r}
library(data.table)
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)
library(geosphere)
library(fastDummies)
source("matches_helper.R")

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# These are some socioeconomic covariates we may evaluate balance for but need to convert from string to num
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population")
df <- df |> mutate(across(all_of(covs_to_check), as.numeric)) # make sure these are numeric

# Add REGION dummy columns
df <- df |> dummy_cols(select_columns = c("REGION"), ignore_na = FALSE, remove_selected_columns = TRUE, remove_first_dummy = FALSE)

# These are the variables we will adjust for
adj_set <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "lat", "lng")

# Other variables
ntee_vars <- grep("^NTEEV2_", names(df), value = TRUE)
region_vars <- grep("^REGION_", names(df), value = TRUE)

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA_nDis"), adj_set, ntee_vars, region_vars)))

# log transform
df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
                                       .default = TOT_REV),
                   TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
                                       .default = TOT_ASSET))
df <- df |> mutate(across(c("TOT_REV", "TOT_ASSET"), log1p)) #log1p adds 1 first to avoid 0s

# Create new treatment variable that takes a distance-based cutoff adjustment
adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 25)

res <- add_exposed(panel_data = df, adj_list = adj_lists$union, treatment = "treat.FEMA_nDis")

df <- res$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))

per_year_summary <- df |> group_by(TAX_YEAR) |>
      summarise(
            n_1 = sum(treat.FEMA.adj == 1, na.rm = TRUE),
            n_0 = sum(treat.FEMA.adj == 0, na.rm = TRUE),
            n = n())

paste("Treated")
quantile(per_year_summary$n_1, probs = seq(0,1,0.1))

paste("Control")
quantile(per_year_summary$n_0, probs = seq(0,1,0.1))

paste("Total")
quantile(per_year_summary$n, probs = seq(0,1,0.1))

# 1859 is New York County
df$treat.FEMA.adj[((df$CountyID == 1859) & (df$TAX_YEAR == 2001))] <- NA # note they were untreated before so don't have to correct above
# df[((df$CountyID == 1859) & (df$TAX_YEAR == 2001)),]

# Adding spatially varying coefficients
df <- add_SVC(df, num_E_vec = 100)
df$lat <- NULL
df$lng <- NULL

df$geoid_2010 <- NULL
df$treat.FEMA_nDis <- NULL
# https://stackoverflow.com/questions/74815244/panelmatch-r-errors-please-convert-time-id-to-consecutive-integers-and-pleas
panel_df <- PanelData(panel.data = as.data.frame(df),
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA.adj",
                      outcome = "TOT_REV")

# rm(df)
# saveRDS(panel_df, "panel_data/int_panelV1.rds") # No SVC, exposed treatment = 1
# saveRDS(panel_df, "panel_data/int_panelV2.rds") # With SVC, exposed treatment = 1

# DisplayTreatment(panel.data = panel_df, legend.position = "none", xlab = "TAX_YEAR", ylab = "County Code", hide.x.tick.label = TRUE, hide.y.tick.label = TRUE, dense.plot = TRUE)
```

## Matching Methods - Config 1, No SVC
### Lag = 3
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/int_panelV1.rds")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/no_svc/int_panelV1/config1/"
folder_path_balance <- "balances/no_svc/int_panelV1/config1/NSEW_"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
# summary(matches_list$none)
# summary(matches_list$ps.match)
# summary(matches_list$CBPS.match)
# summary(matches_list$ps.weight)
# summary(matches_list$CBPS.weight)
# 
# plot(matches_list$ps.match)
#print(matches_list$none)
```


Default Matching Weights Plots
```{r}
# # panel_df <- readRDS("panel_data/int_panelV1.rds")
# extracted_matches <- readRDS(paste0(folder_path_match, "extracted_", save_name, ".rds"))
# # plot(extracted_matches$CBPS.weight, panel.data = panel_df, include.missing = FALSE)
# names(extracted_matches$CBPS.weight)
# extracted_matches$CBPS.weight$`1.1995`
# extracted_matches$CBPS.match$`1.1995`
# weights(extracted_matches$CBPS.weight)
# weights(extracted_matches$ps.match)
```


Default Scatter Plots
```{r}
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```


My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.15,0.2), "unrefined" = c(-0.15, 0.05), "all" = c(-0.2, 0.2))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

# p_none$base
# # p_ps.match$base_unrefined
# p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
# p_cbps.weight$base
# 
# p_none$ntee
# # p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
# p_cbps.weight$ntee
# 
p_none$loc
# p_ps.match$loc_unrefined
p_ps.match$loc
p_cbps.match$loc
p_ps.weight$loc
p_cbps.weight$loc
# 
# p_none$all
# # p_ps.match$all_unrefined
# p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
# p_cbps.weight$all
```

```{r}
quantile(signif(abs(bal_none$value[bal_cbps.weight$unrefined == FALSE]), 3))
quantile(signif(abs(bal_cbps.weight$value[bal_cbps.weight$unrefined == FALSE]), 3))
```



### Lag = 5
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/int_panelV1.rds")

# Set up the folder path and saved object names
L = 5 # lag, options: 1, 3, 5
folder_path_match <- "matches/no_svc/int_panelV1/config1/"
folder_path_balance <- "balances/no_svc/int_panelV1/config1/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
summary(matches_list$none)
summary(matches_list$ps.match)
summary(matches_list$CBPS.match)
summary(matches_list$ps.weight)
summary(matches_list$CBPS.weight)

plot(matches_list$none)
```


Default Matching Weights Plots
```{r}
# plot(extract(matches_list$CBPS.weight), panel.data = panel_df)
```


Default Scatter Plots
```{r}
# plot(balances_list$none, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE, 
#      main = "Standardized Mean Difference of Covariates (No Matching)")
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```


My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))

ylims <- c(-0.2,0.06)
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims, 
                              ylim_ntee = ylims,
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims, 
                                  ylim_ntee = ylims, 
                                  ylim_unrefined = ylims,
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims,
                                  ylim_ntee = ylims,
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims,
                                  ylim_ntee = ylims,
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims,
                                  ylim_ntee = ylims,
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

p_none$base
# p_ps.match$base_unrefined
p_ps.match$base
p_cbps.match$base
p_ps.weight$base
p_cbps.weight$base

# p_none$ntee
# # p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
# p_cbps.weight$ntee
```


### Matching Methods - Config 1, with SVC
### Lag = 3
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/int_panelV1.rds")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/int_panelV2/config1/"
folder_path_balance <- "balances/with_svc/int_panelV2/config1/NSEW_"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# # Print information about matching and balances
# summary(matches_list$none)
# summary(matches_list$ps.match)
# summary(matches_list$CBPS.match)
# summary(matches_list$ps.weight)
# summary(matches_list$CBPS.weight)
# 
# plot(matches_list$ps.match)
# #print(matches_list$none)
```


Default Matching Weights Plots
```{r}
# # panel_df <- readRDS("panel_data/int_panelV1.rds")
# extracted_matches <- readRDS(paste0(folder_path_match, "extracted_", save_name, ".rds"))
# # plot(extracted_matches$CBPS.weight, panel.data = panel_df, include.missing = FALSE)
# names(extracted_matches$CBPS.weight)
# extracted_matches$CBPS.weight$`1.1995`
# extracted_matches$CBPS.match$`1.1995`
# weights(extracted_matches$CBPS.weight)
# weights(extracted_matches$ps.match)
```


Default Scatter Plots
```{r}
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```


My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.15,2), "unrefined" = c(-0.15, 0.05), "all" = c(-0.2, 0.2))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

# p_none$base
# # p_ps.match$base_unrefined
# p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
# p_cbps.weight$base
# 
# p_none$ntee
# # p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
# p_cbps.weight$ntee
# 
p_none$loc
# p_ps.match$loc_unrefined
p_ps.match$loc
p_cbps.match$loc
p_ps.weight$loc
p_cbps.weight$loc
# 
# p_none$all
# # p_ps.match$all_unrefined
# p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
# p_cbps.weight$all
```

```{r}
quantile(signif(abs(bal_none$value[bal_cbps.weight$unrefined == FALSE]), 3))
quantile(signif(abs(bal_cbps.weight$value[bal_cbps.weight$unrefined == FALSE]), 3))
```




### Lag = 5
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/int_panelV1.rds")

# Set up the folder path and saved object names
L = 5 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/int_panelV2/config1/"
folder_path_balance <- "balances/with_svc/int_panelV2/config1/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
summary(matches_list$none)
summary(matches_list$ps.match)
summary(matches_list$CBPS.match)
summary(matches_list$ps.weight)
summary(matches_list$CBPS.weight)

plot(matches_list$ps.match)
#print(matches_list$none)
```


Default Matching Weights Plots
```{r}
# # panel_df <- readRDS("panel_data/int_panelV1.rds")
# extracted_matches <- readRDS(paste0(folder_path_match, "extracted_", save_name, ".rds"))
# # plot(extracted_matches$CBPS.weight, panel.data = panel_df, include.missing = FALSE)
# names(extracted_matches$CBPS.weight)
# extracted_matches$CBPS.weight$`1.1995`
# extracted_matches$CBPS.match$`1.1995`
# weights(extracted_matches$CBPS.weight)
# weights(extracted_matches$ps.match)
```


Default Scatter Plots
```{r}
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.15), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```


My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))

ylims <- c(-0.2,0.06)
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims, 
                              ylim_ntee = ylims,
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims, 
                                  ylim_ntee = ylims, 
                                  ylim_unrefined = ylims,
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims,
                                  ylim_ntee = ylims,
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims,
                                  ylim_ntee = ylims,
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims,
                                  ylim_ntee = ylims,
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

p_none$base
# p_ps.match$base_unrefined
p_ps.match$base
p_cbps.match$base
p_ps.weight$base
p_cbps.weight$base

# p_none$ntee
# # p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
# p_cbps.weight$ntee
```




## Estimates
```{r, message = F}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(fastDummies)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

outcome_vars <- grep("^REV_", names(df), value = TRUE)

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "treat.FEMA_nDis", "TOT_REV"), outcome_vars)))

# log transform
# saveRDS(df |> filter(if_any(all_of(c("TOT_REV", "TOT_ASSET", outcome_vars)), ~ .x < 0)), "misc_analysis/no_intV1_negative_stratified.rds")
df <- df |> mutate(across(all_of(c("TOT_REV", outcome_vars)), ~ if_else(.x < 0, 0, .x)))
df <- df |> mutate(across(all_of(c("TOT_REV", outcome_vars)), log1p)) #log1p adds 1 first to avoid 0s

df$treat.FEMA_nDis[((df$CountyID == 1859) & (df$TAX_YEAR == 2001))] <- NA
df <- df |> rename(treat.FEMA.adj = treat.FEMA_nDis)
# https://stackoverflow.com/questions/74815244/panelmatch-r-errors-please-convert-time-id-to-consecutive-integers-and-pleas

panel_df <- PanelData(panel.data = as.data.frame(df),
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA.adj",
                      outcome = "TOT_REV")
# 
# panel_df2 <- as.data.frame(readRDS("panel_data/int_panelV2.rds"))
# 
# all.equal(as.data.frame(panel_df) |> dplyr::select("CountyID", "TAX_YEAR", "TOT_REV"),
#           panel_df2 |> dplyr::select("CountyID", "TAX_YEAR", "TOT_REV"))

rm(df)
```
### Without Matching
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/int_panelV2/config1/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

panel_df <- readRDS("panel_data/int_panelV2.rds")

# Do we get significant results if we don't do any matching?
res <- PanelEstimate(sets = matches_list$none,
                     panel.data = panel_df,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6)
plot(res)
summary(res)
```


### With Matching
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/int_panelV2/config1/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

panel_df <- readRDS("panel_data/int_panelV2.rds")

# Winner: CBPS Weighting, no SVC Coefficients
res <- PanelEstimate(sets = matches_list$CBPS.weight,
                     panel.data = panel_df,
                     se.method = "bootstrap",
                     parallel = FALSE,
                     num.cores = 6)
plot(res)
summary(res)
```

```{r}
# Winner: CBPS Weighting, no SVC Coefficients
res.uncond <- PanelEstimate(sets = matches_list$CBPS.weight,
                     panel.data = panel_df,
                     se.method = "unconditional")
plot(res.uncond)
summary(res.uncond)
```

## Choosing the number of SVC
```{r}
library(data.table)
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(writexl)
library(haven)
library(geosphere)
source("matches_helper.R")
library(fields)
library(MASS)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# These are some socioeconomic covariates we may evaluate balance for but need to convert from string to num
covs_to_check <- c("white_perc", "bachelors_perc", "poverty_perc", "unemployment", "med_family_income_adj", "med_household_income_adj", "total_population")
df <- df |> mutate(across(all_of(covs_to_check), as.numeric)) # make sure these are numeric

# These are the variables we will assess balance for
se_vars <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population")
other_vars <- c("TOT_REV", "TOT_ASSET")
ntee_vars <- c("NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA")

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA_nDis", "lat", "lng"), se_vars, other_vars, ntee_vars)))

df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
                                       .default = TOT_REV),
                   TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
                                       .default = TOT_ASSET))
df <- df |> mutate(across(c("TOT_REV", "TOT_ASSET"), log1p)) #log1p adds 1 first to avoid 0s

library(data.table)
df <- arrange(df, TAX_YEAR)
df.lag <- shift(df[,7:ncol(df)], n = 1:3, give.names = TRUE)
df <- bind_cols(df, df.lag)
rm(df.lag)

df <- df |> filter(TAX_YEAR >= 1994)

# Create new treatment variable that takes a distance-based cutoff adjustment
adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 25)
res <- add_exposed(panel_data = df, adj_list = adj_lists$union, treatment = "treat.FEMA_nDis")

df <- res$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))
sum(is.na(df$treat.FEMA.adj))

# Spatially Varying Coeffs model setup
df_sub <- df |> dplyr::select(CountyID, lat, lng) |> unique()

x <- cbind(df_sub$lat, df_sub$lng) # inputs: latitude and longitude
d <- rdist(x)
K <- Matern(d, smoothness=.5) # for smoothness paramter, should we do some sort of optimization?
E <- eigen(K)

options(repr.plot.width = 6, repr.plot.height = 4)
for (num_svc in seq(10,100,10)){
      x_s <- cbind(x[,1] * E$vectors[,1:num_svc], x[,2] * E$vectors[,1:num_svc]) # would have to play around with this number
      
      # add SVC back into panel data
      xs_vars <- paste0("x_s", seq_len(ncol(x_s)))
      colnames(x_s) <- xs_vars
      df_sub.temp <- cbind(df_sub, x_s)
      
      df.temp <- merge(df, df_sub.temp |> dplyr::select(-lat, -lng), by = "CountyID", all = TRUE)
      
      upper_formula <- paste(
        "treat.FEMA.adj ~ TAX_YEAR + bachelors_perc + med_household_income_adj + white_perc + total_population +",
        paste(ntee_vars, collapse = " + "), "+",
        paste(xs_vars, collapse = " + "))
      
      # regression
      ls <- glm(as.formula(upper_formula), data = df.temp, family = binomial(link = "logit"))
      summary(ls)

      yweight <- predict(ls, type="response")
      
      png(filename = paste0("svc_plots/int/svc", num_svc*2, ".png"))
      plot(sort(yweight), main = paste("# SVC =", num_svc*2), cex = 0.10)
      dev.off() # decided on 60... maybe 80 is the slightest bit smoother but if we do 60 its same for both
}
```

## For Testing



# Computing Estimates and SE myself
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/no_int_panelV2/config1/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))
msets_list <- readRDS(paste0(folder_path_match, "extracted_", save_name, ".rds"))

# Winner: CBPS Matching with SVC Coefficients
m.out <- matches_list$CBPS.match
msets <- msets_list$CBPS.match

panel_df <- readRDS("panel_data/no_int_panelV2.rds")
panel_df <- as.data.frame(panel_df) |> select(CountyID, TAX_YEAR, TOT_REV)

lead <- 1
num.treated <- length(names(msets))
treated.units <- names(msets)
```

```{r}
# https://chatgpt.com/share/69b9d05d-da30-8003-9e25-56b7953f1e0e
df <- tibble(treated.units) |> 
      separate(treated.units, into = c("CountyID", "t"), sep = "\\.") |>
      mutate(key = paste(CountyID, t, sep = "."),
             CountyID = as.integer(CountyID),
             t = as.integer(t),
             t_plus_F = t + lead)

df <- merge(df, as.data.frame(panel_df), by.x = c("CountyID", "t_plus_F"), by.y = c("CountyID", "TAX_YEAR"), all.x = TRUE)

# msets[["1.1995"]]
matched_controls <- weights(msets)

# class(matched_controls[["1.1995"]])
# names(matched_controls[["1.1995"]])
# matched_controls[["1.1995"]][matched_controls[["1.1995"]] > 0]
matched_controls <- lapply(matched_controls, function(x) x[x > 0])

# Create a lookup key in panel_df in the format CountyID.TAX_YEAR
panel_df$key <- paste(panel_df$CountyID, panel_df$TAX_YEAR, sep = ".")

# Named vector for fast lookup of TOT_REV by "CountyID.TAX_YEAR"
tot_rev_lookup <- setNames(panel_df$TOT_REV, panel_df$key)

# treated_key <- "1.1995"
impute_counterfactual <- function(treated_key){
      # weights for this treated unit
      w <- matched_controls[[treated_key]]
      
      # extract year from treated_key
      yr <- sub(".*\\.", "", treated_key)
      
      # need to grab the total revenue in the year t+lead
      yr <- as.character(as.integer(yr) + lead)
      
      # control lookup keys: controlID.TAX_YEAR
      control_keys <- paste(names(w), yr, sep = ".")
      
      # matched controls' outcomes in the same year
      y_controls <- tot_rev_lookup[control_keys]
      
      # weighted sum
      sum(w * y_controls, na.rm = FALSE)
}

# Compute counterfactual for each treated unit
df$counterfactual <- sapply(df$key, impute_counterfactual)

# Compute estimate
delFL <- mean(df$TOT_REV) - mean(df$counterfactual)

sum(is.na(df$TOT_REV))
sum(is.na(df$counterfactual))
```


# Misc Analyses
How do adjacency-border veruss adjacency-union compare?
```{r}
library(data.table)
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)
library(geosphere)
library(fastDummies)
source("matches_helper.R")

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# These are some socioeconomic covariates we may evaluate balance for but need to convert from string to num
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population")
df <- df |> mutate(across(all_of(covs_to_check), as.numeric)) # make sure these are numeric

# Add REGION dummy columns
df <- df |> dummy_cols(select_columns = c("REGION"), ignore_na = FALSE, remove_selected_columns = TRUE, remove_first_dummy = FALSE)

# These are the variables we will adjust for
adj_set <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "lat", "lng")

# Other variables
ntee_vars <- grep("^NTEEV2_", names(df), value = TRUE)
region_vars <- grep("^REGION_", names(df), value = TRUE)

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA_nDis"), adj_set, ntee_vars, region_vars)))

# log transform
#saveRDS(df |> filter(TOT_REV < 0 | TOT_ASSET < 0), "misc_analysis/no_intV1_negative.rds")
df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
                                       .default = TOT_REV),
                   TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
                                       .default = TOT_ASSET))
df <- df |> mutate(across(c("TOT_REV", "TOT_ASSET"), log1p)) #log1p adds 1 first to avoid 0s

# Create new treatment variable that takes a distance-based cutoff adjustment
# adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 25)
adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 35)

res_border <- add_exposed(panel_data = df, adj_list = adj_lists$border, treatment = "treat.FEMA_nDis")
res_dist <- add_exposed(panel_data = df, adj_list = adj_lists$dist, treatment = "treat.FEMA_nDis")
res_union <- add_exposed(panel_data = df, adj_list = adj_lists$union, treatment = "treat.FEMA_nDis")

df_border <- res_border$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))
# 1859 is New York County
df_border$treat.FEMA.adj[((df_border$CountyID == 1859) & (df_border$TAX_YEAR == 2001))] <- NA
sum(is.na(df_border$treat.FEMA.adj))

df_dist <- res_dist$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))
# 1859 is New York County
df_dist$treat.FEMA.adj[((df_dist$CountyID == 1859) & (df_dist$TAX_YEAR == 2001))] <- NA
sum(is.na(df_dist$treat.FEMA.adj))

df_union <- res_union$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))
# 1859 is New York County
df_union$treat.FEMA.adj[((df_union$CountyID == 1859) & (df_union$TAX_YEAR == 2001))] <- NA
sum(is.na(df_union$treat.FEMA.adj))

# # distance = 25
# sum(df$treat.FEMA_nDis, na.rm = TRUE)
# sum(df_border$treat.FEMA.adj, na.rm = TRUE) # 68182
# sum(df_dist$treat.FEMA.adj, na.rm = TRUE) # 47372
# sum(df_union$treat.FEMA.adj, na.rm = TRUE) # 68609 -- extra 427 treated compared with just border

# distance = 35
sum(df$treat.FEMA_nDis, na.rm = TRUE)
sum(df_border$treat.FEMA.adj, na.rm = TRUE) # 68182
sum(df_dist$treat.FEMA.adj, na.rm = TRUE) # 62003
sum(df_union$treat.FEMA.adj, na.rm = TRUE) # 70855 -- extra 2673 treated compared with just border
```

I would like to label each county by their "top" subsector according to number of nonprofits in that subsector and then count how many counties per subsector there are
```{r}
# https://chatgpt.com/share/69b99471-5294-8003-b284-406850e1205d
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)

df <- readRDS("county_long.rds")

# Create new column called top_ntee
# For a (county,year) pair, the value of top_ntee should be a string containing last 2-3 chars of the "winning" column
ntee_cols <- grep("^NTEEV2_", names(df), value = TRUE)
df <- df |> select(TAX_YEAR, geoid_2010, all_of(ntee_cols))

df <- df |> mutate(
      top_ntee = ntee_cols[max.col(across(all_of(ntee_cols)), ties.method = "first")],
      top_ntee = sub("^NTEEV2_", "", top_ntee)
)

table(df$top_ntee)

top_counts <- df |>
  count(TAX_YEAR, top_ntee) |>
  complete(TAX_YEAR, top_ntee = sub("^NTEEV2_", "", ntee_cols), fill = list(n = 0))

top_counts |> filter(top_ntee == "ART")
```

Now I would like to label "top subsector" by top assets that year
```{r}
df <- readRDS("mega.rds") |>
      filter(county.census.geoid != "00000")#

df.tst <- df |> select(TAX_YEAR, TOT_REV, TOT_EXP, TOT_ASSET, county.census.geoid, NTEEV2) |>
      rename(REV = TOT_REV, EXP = TOT_EXP, ASSET = TOT_ASSET) |>
      mutate(NTEEV2 = replace_na(NTEEV2, "NA")) |>
      group_by(county.census.geoid, TAX_YEAR, NTEEV2) |> 
      summarize(across(where(is.numeric), ~sum(is.na(.x))/n(), .names = "{.col}_prop_miss"),
                across(any_of(c("REV", "EXP", "ASSET")), ~sum(.x, na.rm = TRUE), .names = "TOT_{.col}"),
                .groups = "drop") |>
      ungroup() 

df.tst <- df.tst |> select(-TOT_EXP, -EXP_prop_miss, -ASSET_prop_miss, -REV_prop_miss) |>
      pivot_wider(names_from = NTEEV2, values_from = c("TOT_REV", "TOT_ASSET"), values_fill = 0)

rev_cols <- grep("^TOT_REV_", names(df.tst), value = TRUE)
ass_cols <- grep("^TOT_ASSET_", names(df.tst), value = TRUE)

df.tst <- df.tst |> select(county.census.geoid, TAX_YEAR, all_of(ass_cols))

df.tst <- df.tst |> mutate(
      top_ntee = ass_cols[max.col(across(all_of(ass_cols)), ties.method = "first")],
      top_ntee = sub("^TOT_ASSET_", "", top_ntee)
)

table(df.tst$top_ntee)

top_counts <- df.tst |>
  count(TAX_YEAR, top_ntee) |>
  complete(TAX_YEAR, top_ntee = sub("^TOT_ASSET_", "", ass_cols), fill = list(n = 0))

top_counts |> filter(n > 30) |> select(top_ntee) |> unique()

```

# MATCHIT: Total Effect of a County Experiencing 1 or More Disasters on the County's Nonprofit Revenue
Let the unit of analysis $i$ represent a US county.
Let $A_i^X$ represent the treatment of county $i$ in year $X$. If county $i$ had nonzero disaster damage in the year $X$, then $A_i^X=1$. Otherwise, $A_i^X=0$.
Let $Y_{tot,i}^X$ be the total nonprofit revenue of county $i$ reported in the year $X$. Let $Y_{med,i}^X$ be the median nonprofit revenue of county $i$ reported in the year $X$. 

First, let's create a county-level dataset where each row corresponds to a county and we have a column `treatment` representing $A_i^{2000}$. We will also include the following:
- geoid of the county (FIPS code)
- state the county is in 
- latitude and longitude of the centroid of the county
- the median temperature for the state the county belongs to for the years 1997, 1998, 1999
- socioeconomic factors for the year 1990
- the total number of disasters in that county in the period 1997-1999
- US Census Region or Division
## Data Pre-processing
```{r, message = F}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)

YEAR <- 2000

# Basic county information (lat/long of centroid, county FIPS, county name, state)
counties <- as.data.table(read_csv("counties_data/uscounties.csv", show_col_types = FALSE)) # https://simplemaps.com/data/us-counties

# socioeconomic factors
source("https://raw.githubusercontent.com/UI-Research/nccs-geo/main/get_census_data.R")
census_df  <- get_census_data( geo="county", years=c(1990))   #3,142 counties

df <- merge(census_df, counties, by.x = "geoid_2010", by.y = "county_fips", all.x = TRUE)

# removing unnecessary columns
df <- df |> select(-population, -county_ascii)

# disaster data
disasters <- readRDS("disasters.rds")
disasters <- disasters |> 
      filter(TAX_YEAR >= YEAR-3 & TAX_YEAR <= YEAR) |> 
      group_by(county.census.geoid, TAX_YEAR) |>
      summarize(
            n_disasters = n(),
            across(where(is.numeric), sum, .names = "{.col}_TOT")
      )

df <- merge(df, disasters |> 
                  filter(TAX_YEAR == YEAR) |> 
                  mutate(LOG_TotPropDmg2015 = log(`PropertyDmg(ADJ 2015)_TOT` + 1)) |> 
                  select(county.census.geoid, n_disasters, LOG_TotPropDmg2015), 
            by.x = 'geoid_2010', by.y = 'county.census.geoid', all.x = TRUE)

df <- df |> mutate(n_disasters = replace_na(n_disasters, 0),
                   LOG_TotPropDmg2015 = replace_na(LOG_TotPropDmg2015, 0))

df <- merge(df, disasters |> filter(TAX_YEAR < YEAR) |> group_by(county.census.geoid) |> summarize(n_disasters.PRE = sum(n_disasters)),
            by.x = 'geoid_2010', by.y = 'county.census.geoid', all.x = TRUE )

# If missing, that means there was no corresponding record in disasters df, which means county did not have any disasters in the period 1997-1999
df <- df |> mutate(n_disasters.PRE = replace_na(n_disasters.PRE, 0))

# location (categorical)
source("../SCRIPTS/clean_helper.R")
df <- add_regions_and_divisions(df |> rename(CENSUS_STATE_ABBR = state_id))
```

There are some missing values we need to handle.
```{r}
missing_dat <- na_counts_df(df)

missing_counties <- df |> filter(is.na(county)) |> select(geoid_2010)
missing_counties <- missing_counties$geoid_2010
```
From [County Geography Changes 1990 to Present](https://www.cdc.gov/nchs/nvss/bridged_race/county_geography-_changes1990-present.pdf#:~:text=Valdez%2DCordova%20Census%20Area%2C%20Alaska%20(FIPS,Census%20Area%20(FIPS%20Code%3D%2002066).)
- "02261" : Valdez-Cordova Census Area (Alaska). In 2019, was split into two areas: 02063 and 02066
- "02270" : Wade Hampton Census Area (Alaska). In 2015, was renamed and given new FIPS: 02158
- "46113" : Shannon County (South Dakota). In 2015, was renamed and given new FIPS: 46102

```{r}
cols_to_fill <- c("county", "county_full", "CENSUS_STATE_ABBR", "state_name", "lat", "lng", "REGION", "DIVISION")

df[df$geoid_2010 == "02261", cols_to_fill] <- list("Valdez-Cordova",
                                                   "Valdez-Cordova Census Area",
                                                   "AK",
                                                   "Alaska",
                                                   61.34984,
                                                   -145.023141,
                                                   "WEST",
                                                   "PACIFIC")

df[df$geoid_2010 == "02270", cols_to_fill] <- list("Kusilvak",
                                                   "Kusilvak Census Area",
                                                   "AK",
                                                   "Alaska",
                                                   62.1547,
                                                   -163.3812,
                                                   "WEST",
                                                   "PACIFIC")

df[df$geoid_2010 == "46113", cols_to_fill] <- list("Oglala Lakota",
                                                   "Oglala Lakota County",
                                                   "SD",
                                                   "South Dakota",
                                                   43.3355,
                                                   -102.5516,
                                                   "MIDWEST",
                                                   "CENTRAL-NW")
```

From [Bureau of Labor Statistics - New 2024 Connecticut Counties](https://www.bls.gov/cew/classifications/areas/new-2024-connecticut-counties.htm#:~:text=Table_title:%20New%20County%20Codes%20and%20Titles%20Table_content:,Title:%20Northeastern%20Connecticut%20Planning%20Region%2C%20Connecticut%20%7C):
"In 2022, the State of Connecticut requested that The U.S. Census Bureau adopt its nine planning regions as county-equivalents. These nine planning regions replace Connecticut's eight counties used historically... There is no direct mapping between the old and new counties."
- "09001" : Fairfield County
- "09003" : Hartford County
- "09005" : Litchfield County
- "09007" : Middlesex County
- "09009" : New Haven County
- "09011" : New London County
- "09013" : Tolland County
- "09015" : Windham County

```{r}
df[df$geoid_2010 == "09001", cols_to_fill] <- list("Fairfield", "Fairfield County",
                                                   "CT", "Connecticut",
                                                   41.227413, -73.367061,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09003", cols_to_fill] <- list("Hartford", "Hartford County",
                                                   "CT", "Connecticut",
                                                   41.806053, -72.732916,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09005", cols_to_fill] <- list("Litchfield", "Litchfield County",
                                                   "CT", "Connecticut",
                                                   41.791883, -73.235404,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09007", cols_to_fill] <- list("Middlesex", "Middlesex County",
                                                   "CT", "Connecticut",
                                                   41.433003, -72.52278,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09009", cols_to_fill] <- list("New Haven", "New Haven County",
                                                   "CT", "Connecticut",
                                                   41.349717, -72.900204,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09011", cols_to_fill] <- list("New London", "New London County",
                                                   "CT", "Connecticut",
                                                   41.472652, -72.108634,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09013", cols_to_fill] <- list("Tolland", "Tolland County",
                                                   "CT", "Connecticut",
                                                   41.858081, -72.340978,
                                                   "NORTHEAST", "NEW-ENGLAND")

df[df$geoid_2010 == "09015", cols_to_fill] <- list("Windham", "Windham County",
                                                   "CT", "Connecticut",
                                                   41.824999, -71.990702,
                                                   "NORTHEAST", "NEW-ENGLAND")
```

Now, add climate data and handle missing values.
```{r}
# The following counties had a data download failure: 02261, 02270, 11001, 46113, 51678
# I will fix these by hand below
# Starting with precipitation
precipitation <- readRDS("climate_data/all_county_precip_data.rds")

url_pars <- list(scope = "county",
                 parameter = "pcp",
                 timescale = 1,
                 month = 0,
                 begYear = 1990,
                 endYear = 2020,
                 format = "csv")

all_counties <- data.frame(state = character(), geoid = character(), year = integer(), month = integer(), value = numeric())

# 02261 : Valdez-Cordova Census Area (Alaska). In 2019, was split into two areas: 02063 (Chugach Census Area) and 02066 (Copper River Census Area). 
# Will combine the data via a weighted average: https://data.census.gov/profile gives area (in square miles) of US counties
# Chugach (9,575 qmi) + Copper River (24,647 sqmi) = Valdez-Cordova (34,222 sqmi)

url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-063", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

chugach <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dates <- chugach[[1]] # Note these dates, years, and months vectors will be the same for all records!
dat.year <- as.integer(substr(dates, 1, 4))
dat.month <- as.integer(substr(dates, 5, 6))
chugach.values <- as.double(chugach[[2]])

url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-066", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

copper <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
copper.values <- as.double(copper[[2]])

weighted_avg.values <- (0.28 * chugach.values) + (0.72 * copper.values)

all_counties <- rbind(all_counties, 
                      data.frame(state = "Alaska", geoid = "02261", year = dat.year, month = dat.month, val = weighted_avg.values))

# 02270 : Wade Hampton Census Area (Alaska). In 2015, was renamed and given new FIPS: 02158 (Kusilvak Census Area)
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-158", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "Alaska", geoid = "02270", year = dat.year, month = dat.month, val = dat.values))

# 11001 : Washington D.C. is listed as a county in Maryland (MD-511) in the NOAA API
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "MD-511", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "District of Columbia", geoid = "11001", year = dat.year, month = dat.month, val = dat.values))

# 46113 : Shannon County (South Dakota). In 2015, was renamed and given new FIPS: 46102 (Oglala Lakota County)
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "SD-102", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.year <- as.integer(substr(dat[[1]], 1, 4))
dat.month <- as.integer(substr(dat[[1]], 5, 6))
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "South Dakota", geoid = "46113", year = dat.year, month = dat.month, val = dat.values))

# 51678 : Lexington City (VA) - an independent city meaning not part of a county; either drop or impute as average/median values of Virginia. From Wikipedia: "It is the county seat of Rockbridge County, although the two are separate jurisdictions, and is combined with it for statistical purposes by the Bureau of Economic Analysis." - will impute with Rockbridge County values
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "VA-163", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.year <- as.integer(substr(dat[[1]], 1, 4))
dat.month <- as.integer(substr(dat[[1]], 5, 6))
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "Virginia", geoid = "51678", year = dat.year, month = dat.month, val = dat.values))

precipitation <- rbind(precipitation, all_counties)
saveRDS(precipitation, "climate_data/precipitation.rds")
```

```{r}
# Moving on to average temperature
rm(list=ls())
average_temp <- readRDS("climate_data/all_county_avg_temp_data.rds")

url_pars <- list(scope = "county",
                 parameter = "tavg",
                 timescale = 1,
                 month = 0,
                 begYear = 1990,
                 endYear = 2020,
                 format = "csv")

all_counties <- data.frame(state = character(), geoid = character(), year = integer(), month = integer(), value = numeric())

# 02261 : Valdez-Cordova Census Area (Alaska). In 2019, was split into two areas: 02063 (Chugach Census Area) and 02066 (Copper River Census Area). 
# Will combine the data via a weighted average: https://data.census.gov/profile gives area (in square miles) of US counties
# Chugach (9,575 qmi) + Copper River (24,647 sqmi) = Valdez-Cordova (34,222 sqmi)

url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-063", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

chugach <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dates <- chugach[[1]] # Note these dates, years, and months vectors will be the same for all records!
dat.year <- as.integer(substr(dates, 1, 4))
dat.month <- as.integer(substr(dates, 5, 6))
chugach.values <- as.double(chugach[[2]])

url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-066", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

copper <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
copper.values <- as.double(copper[[2]])

weighted_avg.values <- (0.28 * chugach.values) + (0.72 * copper.values)

all_counties <- rbind(all_counties, 
                      data.frame(state = "Alaska", geoid = "02261", year = dat.year, month = dat.month, val = weighted_avg.values))

# 02270 : Wade Hampton Census Area (Alaska). In 2015, was renamed and given new FIPS: 02158 (Kusilvak Census Area)
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "AK-158", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "Alaska", geoid = "02270", year = dat.year, month = dat.month, val = dat.values))

# 11001 : Washington D.C. is listed as a county in Maryland (MD-511) in the NOAA API
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "MD-511", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "District of Columbia", geoid = "11001", year = dat.year, month = dat.month, val = dat.values))

# 46113 : Shannon County (South Dakota). In 2015, was renamed and given new FIPS: 46102 (Oglala Lakota County)
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "SD-102", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.year <- as.integer(substr(dat[[1]], 1, 4))
dat.month <- as.integer(substr(dat[[1]], 5, 6))
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "South Dakota", geoid = "46113", year = dat.year, month = dat.month, val = dat.values))

# 51678 : Lexington City (VA) - an independent city meaning not part of a county; either drop or impute as average/median values of Virginia. From Wikipedia: "It is the county seat of Rockbridge County, although the two are separate jurisdictions, and is combined with it for statistical purposes by the Bureau of Economic Analysis." - will impute with Rockbridge County values
url <- sprintf("https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/%s/time-series/%s/%s/%d/%d/%d-%d/data.%s", 
               url_pars$scope, "VA-163", url_pars$parameter, url_pars$timescale, 
               url_pars$month, url_pars$begYear, url_pars$endYear, url_pars$format)

dat <- read.csv(url) |> slice(3:n()) # first two rows are unecessary
dat.year <- as.integer(substr(dat[[1]], 1, 4))
dat.month <- as.integer(substr(dat[[1]], 5, 6))
dat.values <- as.double(dat[[2]])
all_counties <- rbind(all_counties, 
                      data.frame(state = "Virginia", geoid = "51678", year = dat.year, month = dat.month, val = dat.values))

average_temp <- rbind(average_temp, all_counties)
saveRDS(average_temp, "climate_data/average_temperature.rds")
```

```{r}
# climate data
precipitation <- readRDS("climate_data/precipitation.rds") |> filter(year >= YEAR-3 & year < YEAR) 

df <- merge(df, precipitation |> group_by(geoid) |> summarize(med_precip = median(val, na.rm = TRUE), mean_precip = mean(val, na.rm = TRUE)),
      by.x = "geoid_2010", by.y = "geoid", all.x = TRUE)

avg_temp <- readRDS("climate_data/average_temperature.rds") |> filter(year >= YEAR-3 & year < YEAR)

df <- merge(df, avg_temp |> group_by(geoid) |> summarize(med_avg_temp = median(val, na.rm = TRUE), mean_avg_temp = mean(val, na.rm = TRUE)),
      by.x = "geoid_2010", by.y = "geoid", all.x = TRUE)

missing_dat <- na_counts_df(df)
```

Let's add in a few treatment options.
```{r}
df <- df |> mutate(treat.n_disasters = as.numeric(n_disasters > 0),
                   treat.median_log_damage = as.numeric(LOG_TotPropDmg2015 > 11), # about 60K of property damage
                   treat.cat = case_when(
                         LOG_TotPropDmg2015 <= 0 ~ "none",
                         LOG_TotPropDmg2015 <= 11 ~ "small", # up to ~60K
                         LOG_TotPropDmg2015 <= 13 ~ "med", # up to ~442K
                         LOG_TotPropDmg2015 <= 22 ~ "large", # up to 3.5 trillion
                   ))

df |> select(geoid_2010, LOG_TotPropDmg2015, n_disasters, treat.n_disasters, treat.median_log_damage, treat.cat)
```


## Pre-matching Balance
Now we can assess covariate balance prior to matching.
```{r}
library(MatchIt)

m.initial <- matchit(treat.median_log_damage ~ lat*lng + med_precip + med_avg_temp,
                     data = df,
                     method = NULL,
                     distance = "glm")

summary(m.initial)
```

# Scratchwork


Testing if simply manually changing the outcome attributes for PanelData and PanelMatch objects will allow PanelEstimate to do its thing
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/no_int_panelV2/config1/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

panel_df <- readRDS("panel_data/no_int_panelV2.rds")
# attributes(panel_df)
attr(panel_df, "outcome") <- "TOT_ASSET"

m.out <- matches_list$CBPS.match
# attributes(m.out)
attr(m.out, "outcome") <- "TOT_ASSET"
attr(m.out, "outcome.var") <- "TOT_ASSET"

res <- PanelEstimate(sets = m.out,
                     panel.data = panel_df,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6)
plot(res, ylim=c(-90000000, 40000000))
summary(res)
```


County FIPS disrepencies in SHELDUS data
```{r}
# 51: Virginia, 48: Texas, 46: SD, 30: Montana, 12: Florida, 02: Alaska
# 51780: South Boston City (After 1990, added to Halifax County: 51083 in 1995)
# 51560: Clifton Forge  (Added to Alleghany: 51005 in 2001)
# 51515: Bedford City (Added to Bedford County: 51019 in 2013)
# 51151: (*Princess Anne) (Was absorbed by Virginia Beach city in 1963 - SHELDUS)
# 51123: (*Nansemond) (Was added to City of Suffolk: 51800 in 1972)
# 48301: Loving  (tiny tiny little county that probably does not have any nonprofits big enough to appear in CORE)
# 30113: Yellowstone National Park; (Annexed to Gallatin (30-031) and Park (30-067) counties effective November 7, 1997)
# 12025: Dade County (Renamed as Miami-Dade County (12-086) effective July 22, 1997)
# 02280: Wrangell-Petersburg Census Area (split into 02195 and 02275 effective June 1, 2008.)
# 02260: (*Valdez-Chitina-Whittier) (Ceased to exist by 1980 - SHEDLUS)
# 02232: Skagway-Hoonah-Angoon Census Area (Created from the former Skagway-Yakutat-Angoon Census Area: 02231, September 22, 1992)
# 02231: Skagway-Yakutat-Angoon Census Area (Split to create the Skagway-Hoonah-Angoon Census Area (02232) and Yakutat City and Borough (02282) effective September 22, 1992) 
# 02201: Prince of Wales-Outer Ketchikan Census Area (Part (Outer Ketchikan area) annexed by Ketchikan Gateway Borough (02130) and part (Meyers Chuck area) included in new Wrangell City and Borough (02275), remainder renamed Prince of Wales-Hyder Census Area (02198) effective May 19, 2008.)
# 02160: (*Kuskokwim) - stopped existing by 1980 (SHELDUS)
# 02120: (*Kenai-Cook Inlet) - stopped existing by 1980 (SHELDUS)
# 02065: (*Cordova-McCarthy) - stopped existing by 1980 

disasters$geoid_2010[disasters$county.geoid == "02158"] <- "02270" # 02270 got new FIPS 02158 in 2015
disasters$geoid_2010[disasters$county.geoid == "46102"] <- "46113" # 46113 got new FIPS 46102 in 2015
disasters$geoid_2010[disasters$county.geoid == "51780"] <- "51083" # 51780 added to 51083 in 1995
disasters$geoid_2010[disasters$county.geoid == "51560"] <- "51005" # 51560 added to 51005 in 2001
disasters$geoid_2010[disasters$county.geoid == "51515"] <- "51019" # 51515 added to 51019 in 2013
disasters$geoid_2010[disasters$county.geoid == "12025"] <- "12086" # 12025 got new FIPS 12086 in 1997
```


stepwise regression
```{r}
# lower_formula  <- paste(
#   "treat.FEMA_nDis ~ TAX_YEAR + bachelors_perc + med_household_income_adj + white_perc + total_population +",
#   paste(ntee_vars, collapse = " + "))

# stepwise regression (inefficient for the large dataset I have right now)
# res <- step(object = ls,
#             scope = list("upper" = upper_formula, "lower" = lower_formula),
#             direction = "backward")
```

create_adj_list and add_exposed
```{r}
create_adj_list <- function(panel_data, dist_cutoff){
      coords <- as.data.frame(panel_data |> select(geoid_2010, lng, lat) |> unique())
      county_IDs <- coords$geoid_2010
      
      # Border-based adj list
      census_df <- read.delim("counties_data/county_adjacency2010.txt",
                       header = FALSE,
                       sep = "\t",
                       col.names = c("county_name", "county_geoid",
                                     "neighbor_name", "neighbor_geoid"),
                       colClasses = c("character", "character", "character", "character"),
                       stringsAsFactors = FALSE)
      
      census_df <- census_df |>
            mutate(across(everything(), \(x) na_if(x, y = "")))
      
      census_df <- census_df |>
            fill(county_name, county_geoid, .direction = "down")
      
      census_df <- census_df[census_df$county_geoid != census_df$neighbor_geoid, ]
      
      census_df <- census_df |> filter((county_geoid %in% county_IDs) & (neighbor_geoid %in% county_IDs))
      
      census_adj_list <- split(census_df$neighbor_geoid, census_df$county_geoid)
      
      missing <- setdiff(county_IDs, names(census_adj_list))
      for (m in missing) {
            census_adj_list[[m]] <- character(0)
      }
      
      # Distance based adj list
      D <- distm(coords |> select(-geoid_2010), fun = distHaversine) / 1609.344 #(meters to miles conversion)
      dimnames(D) <- list(county_IDs, county_IDs)
      
      dist_cutoff <- 25
      idx <- which(D <= dist_cutoff & D > 0, arr.ind = TRUE) # edge list (directed: (i,j) and (j,i) both appear)
      
      edges <- data.frame(county1 = county_IDs[idx[, 1]], 
                          county2 = county_IDs[idx[, 2]], 
                          stringsAsFactors = FALSE)
      
      dist_adj_list <- split(edges$county2, edges$county1)
      
      # If there are missing counties, that means that they had no neighboring counties within the cutoff distance -- so can list an empty vector for them
      missing <- setdiff(unique(df$geoid_2010), names(dist_adj_list))
      for (m in missing) {
            dist_adj_list[[m]] <- character(0)
      }
      
      # for every county, new adj_list is union of the two
      adj_union <- setNames(lapply(names(dist_adj_list), function(n) union(dist_adj_list[[n]], census_adj_list[[n]])),
                            names(dist_adj_list))
      
      return(list("union" = adj_union, "border" = census_adj_list, "dist" = dist_adj_list))
}

add_exposed <- function(panel_data, adj_list){
      panel_data$treat.FEMA.adj <- panel_data$treat.FEMA # initialize adjusted treatment with original
      years <- unique(panel_data$TAX_YEAR)
      
      for (yr in years) {
            idx <- which(panel_data$TAX_YEAR == yr) # Subset indices for this year
            treated <- panel_data$geoid_2010[idx][panel_data$treat.FEMA[idx] == 1]
            
            if (length(treated) == 0) next # If no treated counties, skip
            
            neighbors <- unique(unlist(adj_list[treated])) # Get neighbors of treated counties
            
            untreated_idx <- idx[panel_data$treat.FEMA[idx] == 0] # Identify untreated counties in this year
            
            exposed <- panel_data$geoid_2010[untreated_idx] %in% neighbors # Among untreated, find those exposed
            
            panel_data$treat.FEMA.adj[untreated_idx][exposed] <- NA # Set exposed controls to NA
      }
      
      # per year I want to see the counts of treated, control and missing for treat.FEMA.adj
      per_year_summary <- panel_data |> group_by(TAX_YEAR) |>
            summarise(
                  n_1 = sum(treat.FEMA.adj == 1, na.rm = TRUE),
                  n_0 = sum(treat.FEMA.adj == 0, na.rm = TRUE),
                  n_NA = sum(is.na(treat.FEMA.adj)),
                  n = n())
      
      return(list("df" = panel_data, "summary" = per_year_summary))
}

library(tidyr)
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)
library(geosphere)

df <- readRDS("county_long.rds") |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 50)
resUNION <- add_exposed(panel_data = df, adj_list = adj_lists$union)
resBORDER <- add_exposed(panel_data = df, adj_list = adj_lists$border)
resDIST <- add_exposed(panel_data = df, adj_list = adj_lists$dist)

resUNION$summary
resBORDER$summary
resDIST$summary
```

creating a formula with paste and collapse and ntee variables
```{r}
ntee <- colnames(df)[10:22]
paste0("I(lag(", ntee, ", 1:1))", collapse = " + ")
```

add_exposedOLD - when we used the NBER data
```{r}
add_exposedOLD <- function(dist_data, dist_cutoff, panel_data){
      # Filter by distance cutoff
      edges <- dist_data[dist_data$mi_to_county <= dist_cutoff, c("county1", "county2")]
      
      # Build adjacency list
      adj_list <- split(edges$county2, edges$county1) 
      
      # If there are missing counties, that means that they had no neighboring counties within the cutoff distance -- so can list an empty vector for them
      missing <- setdiff(unique(panel_data$geoid_2010), names(adj_list))
      for (m in missing) {
            adj_list[[m]] <- character(0)
      }
      
      panel_data$treat.FEMA.adj <- panel_data$treat.FEMA # initialize adjusted treatment with original
      years <- unique(panel_data$TAX_YEAR)
      
      for (yr in years) {
            idx <- which(panel_data$TAX_YEAR == yr) # Subset indices for this year
            treated <- panel_data$geoid_2010[idx][panel_data$treat.FEMA[idx] == 1]
            
            if (length(treated) == 0) next # If no treated counties, skip
            
            neighbors <- unique(unlist(adj_list[treated])) # Get neighbors of treated counties
            
            untreated_idx <- idx[panel_data$treat.FEMA[idx] == 0] # Identify untreated counties in this year
            
            exposed <- panel_data$geoid_2010[untreated_idx] %in% neighbors # Among untreated, find those exposed
            
            panel_data$treat.FEMA.adj[untreated_idx][exposed] <- NA # Set exposed controls to NA
      }
      
      # per year I want to see the counts of treated, control and missing for treat.FEMA.adj
      per_year_summary <- panel_data |> group_by(TAX_YEAR) |>
            summarise(
                  n_1 = sum(treat.FEMA.adj == 1, na.rm = TRUE),
                  n_0 = sum(treat.FEMA.adj == 0, na.rm = TRUE),
                  n_NA = sum(is.na(treat.FEMA.adj)),
                  n = n())
      
      return(list("df" = panel_data, "summary" = per_year_summary))
}
```

```{r}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)
library(fields)

df <- readRDS("county_long.rds") |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

nber_data <- read_dta("counties_data/sf12010countydistancemiles.dta")

res25 <- add_exposedOLD(dist_data = nber_data, dist_cutoff = 25, panel_data = df)
res50 <- add_exposedOLD(dist_data = nber_data, dist_cutoff = 50, panel_data = df)
res100 <- add_exposedOLD(dist_data = nber_data, dist_cutoff = 100, panel_data = df)

res25$summary
res50$summary
res100$summary
```

creating the nonprodit attributes covaraties (NTEE distribtuion)
```{r}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(fastDummies)

df <- readRDS("mega.rds") # 1773 orgs have "00000" as geoid; 86,752 are missing NTEE code (~6.74%) -- 1,285,846 total orgs

# Split up NTEEV2 into 14 dummy variables (12 categories, 1 unknown category, 1 missing category)
ntee <- df |> 
      select(EIN2, NTEEV2, TAX_YEAR, county.census.geoid) |>
      filter(county.census.geoid != "00000") |>
      dummy_cols(select_columns = c("NTEEV2"), ignore_na = FALSE, remove_selected_columns = TRUE)

# Add aggregate the totals of each NTEEV2 category by county and year
ntee <- ntee |> select(-EIN2) |>
      group_by(county.census.geoid, TAX_YEAR) |>
      summarise(across(everything(), ~sum(.x, na.rm = TRUE))) |>
      ungroup()

# Add in total number of organizations and proportion missing column
ntee <- ntee |> mutate(NUM_ORGS = rowSums(ntee |> select(-county.census.geoid, -TAX_YEAR)),
                       NTEE_NA_PROP = NTEEV2_NA/NUM_ORGS)

# Within each county,year pair, I want to keep track of the proportion of records with missing financial data
# I also want to compute the total revenue, expenses, and assets across that county in that year
# For now, we are dropping records that listed the invalid county code "00000" (1773 orgs)
df <- df |> select(TAX_YEAR, TOT_REV, TOT_EXP, TOT_ASSET, county.census.geoid) |>
      rename(REV = TOT_REV, EXP = TOT_EXP, ASSET = TOT_ASSET) |>
      group_by(county.census.geoid, TAX_YEAR) |> 
      summarize(across(where(is.numeric), ~sum(is.na(.x))/n(), .names = "{.col}_prop_miss"),
                across(any_of(c("REV", "EXP", "ASSET")), ~sum(.x, na.rm = TRUE), .names = "TOT_{.col}"),
                .groups = "drop") |>
      ungroup() |>
      filter(county.census.geoid != "00000")

df <- merge(df, ntee, by = c("county.census.geoid", "TAX_YEAR")) 

# IDs for two counties were changed after 2010. Some datasets use latest ID, others use 2010 ID
df <- df |> mutate(geoid_2010 = county.census.geoid)
df$geoid_2010[df$county.census.geoid == "02158"] <- "02270"
df$geoid_2010[df$county.census.geoid == "46102"] <- "46113"
```

adding SVC
```{r}
# File: perform_refinement.R
# glm(reformulate(response = treatment, termlabels = colnames(pooled)[-c(1:3)]), family = binomial(link = "logit"), data = pooled)
# CBPS::CBPS(reformulate(response = treatment, termlabels = colnames(pooled)[-c(1:3)]), family = binomial(link = "logit"), data = pooled)

library(fields)
library(MASS)

set.seed(2727)
df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

df_sub <- df |> select(CountyID, lat, lng) |> unique()

x <- cbind(df_sub$lat, df_sub$lng) # inputs: latitude and longitude

# SVC
d <- rdist(x)
K <- Matern(d, smoothness=.5) # for smoothness paramter, should we do some sort of optimization?
E <- eigen(K)

x_s <- cbind(x[,1] * E$vectors[,1:45], x[,2] * E$vectors[,1:45]) # would have to play around with this number
# add SVC back into panel data
xs_vars <- paste0("x_s", seq_len(ncol(x_s)))
colnames(x_s) <- xs_vars
df_sub <- cbind(df_sub, x_s)

df <- merge(df, df_sub |> dplyr::select(-lat, -lng), by = "CountyID", all = TRUE)

# y <- as.matrix(df$treat.FEMA)

ntee_vars <- c("NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA")
formula_tst <- paste(
  "treat.FEMA ~ TAX_YEAR + bachelors_perc + med_household_income_adj + white_perc + pop_estimate +",
  paste(ntee_vars, collapse = " + "), "+",
  paste(xs_vars, collapse = " + "))

# regression
ls <- glm(as.formula(formula_tst), data = df, family = binomial(link = "logit"))
summary(ls)
```