---
title: "Final_Analysis"
author: "Mayleen"
date: "2026-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PanelMatch - No Interference
## Setup
For matching (just tot rev)
```{r, message = F}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(fastDummies)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# make sure these are numeric
df <- df |> mutate(across(all_of(c("bachelors_perc", "med_household_income_adj", "white_perc")), as.numeric)) 

# Add REGION dummy columns
df <- df |> dummy_cols(select_columns = c("REGION"), ignore_na = FALSE, remove_selected_columns = TRUE, remove_first_dummy = FALSE)

# These are the variables we will adjust for
adj_set <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET","TOT_EXP", "lat", "lng")

# Other variables
ntee_vars <- grep("^NTEEV2_", names(df), value = TRUE)
region_vars <- grep("^REGION_", names(df), value = TRUE)

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA"), adj_set, ntee_vars, region_vars)))

# 1859 is New York County
# df$treat.FEMA[((df$CountyID == 1859) & (df$TAX_YEAR == 2001))] <- NA

# Log transformations
# saveRDS(df |> filter(TOT_REV < 0 | TOT_ASSET < 0 | TOT_EXP < 0), "misc_analysis/no_intV2_negative.rds")
# df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
#                                        .default = TOT_REV),
#                    TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
#                                        .default = TOT_ASSET),
#                    TOT_EXP = case_when(TOT_EXP < 0 ~ 0,
#                                        .default = TOT_EXP))
# 
# df <- df |> mutate(TOT_REV = log1p(TOT_REV),
#                    TOT_ASSET = log1p(TOT_ASSET),
#                    TOT_EXP = log1p(TOT_EXP),
#                    med_household_income_adj = log1p(med_household_income_adj),
#                    total_population = log(total_population))

# Adding spatially varying coefficients
source("matches_helper.R")
df <- add_SVC(df, num_E_vec = 50)

# https://stackoverflow.com/questions/74815244/panelmatch-r-errors-please-convert-time-id-to-consecutive-integers-and-pleas
panel_df <- PanelData(panel.data = as.data.frame(df),
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA",
                      outcome = "TOT_REV")
# rm(df)

# saveRDS(panel_df, "panel_data/no_int_panelV1.rds") # No SVC
# saveRDS(panel_df, "panel_data/no_int_panelV2.rds") # Yes SVC
# saveRDS(panel_df, "panel_data/no_int_panelV3.rds") # Yes SVC BUT NO LOG TRANSFORMS

# write.csv(df |> dplyr::select(TAX_YEAR, FEMA_threshADJ) |> unique(), "updated_threshold.csv", row.names = FALSE)
```

```{r}
sum(is.na(panel_df$treat.FEMA)) # county_FEMA_indicator <- 3.78 #2019

treatVarPlot <- DisplayTreatment(panel.data = panel_df, 
                                 color.of.treated = "#DC267F",
                                 color.of.untreated = "#648FFF",
                                 title = "Treatment Distribution Across Units and Time",
                                 xlab = "Tax Year", ylab = "County", 
                                 legend.position = "right",
                                 legend.labels = c("untreated", "treated"),
                                 hide.x.tick.label = TRUE, hide.y.tick.label = TRUE, 
                                 dense.plot = TRUE)

treatVarPlot
#ggsave("treatment_variation_plot.png", plot = treatVarPlot, width = 7.5, height = 10, units = "in")
```

## Matching Methods - Config 3, With SVC
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/no_int_panelV2.rds")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/no_int_panelV2/config3/"
folder_path_balance <- "balances/with_svc/no_int_panelV2/config3/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
# write.csv(as.data.frame(summary(matches_list$none)), "match_summary.csv")
summary(matches_list$none)
summary(matches_list$ps.match)
# summary(matches_list$CBPS.match)
# summary(matches_list$ps.weight)
# summary(matches_list$CBPS.weight)
# 
plot(matches_list$none)
```


Default Matching Weights Plots
```{r}
# plot(extract(matches_list$CBPS.weight), panel.data = panel_df)
```


Default Scatter Plots
```{r}
# plot(balances_list$none, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE, 
#      main = "Standardized Mean Difference of Covariates (No Matching)")
# plot(balances_list$mahalanobis, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
#      ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
# plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
#      ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
# plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
#      ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```

My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "TOT_EXP", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.15,0.15), "unrefined" = c(-0.15, 0.05), "all" = c(-0.15, 0.15))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

p_none$all
# p_ps.match$all_unrefined
p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
p_cbps.weight$all
# 
p_none$base
# p_ps.match$base_unrefined
p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
p_cbps.weight$base
# 
p_none$ntee
# p_ps.match$ntee_unrefined
p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
p_cbps.weight$ntee
# 
p_none$loc
# p_ps.match$loc_unrefined
p_ps.match$loc
# p_cbps.match$loc
# p_ps.weight$loc
p_cbps.weight$loc

```

```{r}
q_none <- quantile(signif(abs(bal_none$value[bal_none$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_ps.mat <- quantile(signif(abs(bal_ps.match$value[bal_ps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_cbps.mat <- quantile(signif(abs(bal_cbps.match$value[bal_cbps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_ps.w <- quantile(signif(abs(bal_ps.weight$value[bal_ps.weight$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_cbps.w <- quantile(signif(abs(bal_cbps.weight$value[bal_cbps.weight$unrefined == FALSE]), 3), probs = seq(0,1,0.1))

df <- cbind(as.data.frame(q_none), as.data.frame(q_ps.mat), as.data.frame(q_cbps.mat), as.data.frame(q_ps.w), as.data.frame(q_cbps.w))

df$winner <- colnames(df)[max.col(-df, ties.method = "first")]

write.csv(df, file = paste0(folder_path_balance, save_name, "deciles.csv"))

# write.csv(bal_none |> filter(unrefined == FALSE) |> select(-unrefined) |> 
#                 pivot_wider(names_from = covariate, values_from = value) |>
#                 mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
#           file = "balance.csv")
# 
# write.csv(bal_ps.match |> filter(unrefined == FALSE) |> select(-unrefined) |> 
#                 pivot_wider(names_from = covariate, values_from = value) |>
#                 mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
#           file = "balance.csv")
# 
# write.csv(bal_cbps.weight |> filter(unrefined == FALSE) |> select(-unrefined) |> 
#                 pivot_wider(names_from = covariate, values_from = value) |>
#                 mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
#           file = "balance.csv")
```

## Matching Methods - Config 4, With SVC
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/no_int_panelV3.rds")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/no_int_panelV3/config4/"
folder_path_balance <- "balances/with_svc/no_int_panelV3/config4/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
# write.csv(as.data.frame(summary(matches_list$none)), "match_summary.csv")
summary(matches_list$none)
summary(matches_list$ps.match)
# summary(matches_list$CBPS.match)
# summary(matches_list$ps.weight)
# summary(matches_list$CBPS.weight)
# 
plot(matches_list$none)
```


Default Matching Weights Plots
```{r}
# plot(extract(matches_list$CBPS.weight), panel.data = panel_df)
```


Default Scatter Plots
```{r}
# plot(balances_list$none, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE, 
#      main = "Standardized Mean Difference of Covariates (No Matching)")
# plot(balances_list$mahalanobis, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
#      ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
# plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
#      ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
# plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
#      ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```

My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "TOT_EXP", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.15,0.15), "unrefined" = c(-0.15, 0.05), "all" = c(-0.15, 0.15))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

p_none$all
# p_ps.match$all_unrefined
p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
# p_cbps.weight$all
# 
# p_none$base
# p_ps.match$base_unrefined
p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
# p_cbps.weight$base
# 
# p_none$ntee
# p_ps.match$ntee_unrefined
p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
# p_cbps.weight$ntee
# 
# p_none$loc
# p_ps.match$loc_unrefined
p_ps.match$loc
# p_cbps.match$loc
# p_ps.weight$loc
# p_cbps.weight$loc

```

```{r}
q_none <- quantile(signif(abs(bal_none$value[bal_none$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_ps.mat <- quantile(signif(abs(bal_ps.match$value[bal_ps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_cbps.mat <- quantile(signif(abs(bal_cbps.match$value[bal_cbps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_ps.w <- quantile(signif(abs(bal_ps.weight$value[bal_ps.weight$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_cbps.w <- quantile(signif(abs(bal_cbps.weight$value[bal_cbps.weight$unrefined == FALSE]), 3), probs = seq(0,1,0.1))

df <- cbind(as.data.frame(q_none), as.data.frame(q_ps.mat), as.data.frame(q_cbps.mat), as.data.frame(q_ps.w), as.data.frame(q_cbps.w))

df$winner <- colnames(df)[max.col(-df, ties.method = "first")]

write.csv(df, file = paste0(folder_path_balance, save_name, "deciles.csv"))

write.csv(bal_none |> filter(unrefined == FALSE) |> select(-unrefined) |>
                pivot_wider(names_from = covariate, values_from = value) |>
                mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
          file = "balance.csv")

write.csv(bal_ps.match |> filter(unrefined == FALSE) |> select(-unrefined) |>
                pivot_wider(names_from = covariate, values_from = value) |>
                mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
          file = "balance.csv")
# 
# write.csv(bal_cbps.weight |> filter(unrefined == FALSE) |> select(-unrefined) |> 
#                 pivot_wider(names_from = covariate, values_from = value) |>
#                 mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
#           file = "balance.csv")
```


## Estimates
### Config 3 versus Config 4 - No Stratification

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(tidyr)

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
save_name <- paste0("Lag", L)

folder_pathC3 <- "matches/with_svc/no_int_panelV2/config3/"
folder_pathC4 <- "matches/with_svc/no_int_panelV3/config4/"

# Retrieve the PanelMatch objects
matches_listC3 <- readRDS(paste0(folder_pathC3, save_name, ".rds"))
matches_listC4 <- readRDS(paste0(folder_pathC4, save_name, ".rds"))

# Retrieve the PanelData object
panel_dfC3 <- readRDS("panel_data/no_int_panelV2.rds")
panel_dfC4 <- readRDS("panel_data/no_int_panelV3.rds")

estimates_save_nameC3 <- "estimates/with_svc_no_intV2_con3_"
estimates_save_nameC4 <- "estimates/with_svc_no_intV3_con4_"

# No Matching
m.out.noneC3 <- matches_listC3$none
m.out.noneC4 <- matches_listC4$none
# attr(m.out.none, "outcome") <- "EXP_LN_REV"
# attr(m.out.none, "outcome.var") <- "EXP_LN_REV"

match_methodC3 <- "ps.match"
match_methodC4 <- "ps.match"

m.out.C3 <- matches_listC3[[match_methodC3]]
m.out.C4 <- matches_listC4[[match_methodC4]]

# outcome_vars <- c("TOT_REV", "REV_ART", "REV_EDU", "REV_HMS", "REV_PSB", "REV_ENV", "REV_HEL", "REV_HOS", "REV_UNI", "REV_MMB", "REV_ENV") # REV_REL, REV_IFA, REV_UNU, REV_NA

# outcome_vars <- c("REV_SIZE.1", "REV_SIZE.2", "REV_SIZE.3", "REV_SIZE.4", "REV_SIZE.5", "REV_SIZE.6") # "REV_SIZE.NA"
outcome_vars <- c("TOT_REV")
```

```{r}
# Estimates under no matching
res_no.match.C3 <- PanelEstimate(sets = m.out.noneC3,
                     panel.data = panel_dfC3,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

res_no.match.C4 <- PanelEstimate(sets = m.out.noneC4,
                     panel.data = panel_dfC4,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

# Estimates under CBPS Weighting
res.C3 <- PanelEstimate(sets = m.out.C3,
                     panel.data = panel_dfC3,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

res.C4 <- PanelEstimate(sets = m.out.C4,
                     panel.data = panel_dfC4,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)
```

```{r}
plot(res_no.match.C3, main = "Estimated Effects (LOG REV) - No Matching")
summary(res_no.match.C3)
hist(res_no.match.C3$bootstrapped.estimates) 
# as.data.frame(summary(res_no.match.C3)) |> mutate(across(where(is.numeric), ~ exp(.x)-1))

plot(res_no.match.C4, main = "Estimated Effects (REV) - No Matching")
summary(res_no.match.C4)
hist(res_no.match.C4$bootstrapped.estimates)

plot(res.C3, main = "Estimated Effects (LOG REV) - PS Matching")
summary(res.C3)
hist(res.C3$bootstrapped.estimates)

plot(res.C4, main = "Estimated Effects (REV) - PS Matching")
summary(res.C4)
hist(res.C4$bootstrapped.estimates)
```



### Config 3 versus Config 4 - With Stratification by SIZE
```{r, message = F}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(fastDummies)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# outcome_vars <- grep("^REV_", names(df), value = TRUE)
outcome_vars <- c("REV_SIZE.1", "REV_SIZE.2", "REV_SIZE.3", "REV_SIZE.4", "REV_SIZE.5", "REV_SIZE.6")

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR"), outcome_vars))) 

panel_dfC4 <- readRDS("panel_data/no_int_panelV3.rds")
df <- merge(panel_dfC4, df, by = c("CountyID", "TAX_YEAR"))

df <- as.data.frame(df) |> select(all_of(c(c("CountyID", "TAX_YEAR", "TOT_REV", "treat.FEMA"), outcome_vars)))
panel_dfC4 <- PanelData(panel.data = df,
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA",
                      outcome = "TOT_REV")

# Log transformations
panel_dfC3 <- panel_dfC4 |> mutate(across(all_of(outcome_vars), ~ if_else(.x < 0, 0, .x)))
panel_dfC3 <- panel_dfC3 |> mutate(across(all_of(outcome_vars), log1p))
rm(df)
```

```{r}
# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
save_name <- paste0("Lag", L)

folder_pathC3 <- "matches/with_svc/no_int_panelV2/config3/"
folder_pathC4 <- "matches/with_svc/no_int_panelV3/config4/"

# Retrieve the PanelMatch objects
matches_listC3 <- readRDS(paste0(folder_pathC3, save_name, ".rds"))
matches_listC4 <- readRDS(paste0(folder_pathC4, save_name, ".rds"))

estimates_save_nameC3 <- "estimates/with_svc_no_intV2_con3_"
estimates_save_nameC4 <- "estimates/with_svc_no_intV3_con4_"

match_methodC3 <- "ps.match"
match_methodC4 <- "ps.match"

m.out.C3 <- matches_listC3[[match_methodC3]]
m.out.C4 <- matches_listC4[[match_methodC4]]

res.C3 <- list()
for (x in outcome_vars){
      print(x)
      attr(panel_dfC3, "outcome") <- x
      attr(m.out.C3, "outcome") <- x
      attr(m.out.C3, "outcome.var") <- x
      
      res.C3[[x]] <- PanelEstimate(sets = m.out.C3,
                     panel.data = panel_dfC3,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)
}

for (x in outcome_vars){
      file_name <- paste0("estimates/c3", x, ".png")
      print(file_name)
      png(file_name)
      plot(res.C3[[x]], main = paste0("Estimated Effects (LOG ", x, ") - ", match_methodC3))
      dev.off()
}

file_name <- paste0(estimates_save_nameC3, save_name, ".xlsx")
print(file_name)
estimate_summaries.C3 <- lapply(res.C3, summary)
estimate_summaries.C3 <- lapply(estimate_summaries.C3, as.data.frame)
write_xlsx(estimate_summaries.C3, file_name)

res.C4 <- list()
for (x in outcome_vars){
      print(x)
      attr(panel_dfC4, "outcome") <- x
      attr(m.out.C4, "outcome") <- x
      attr(m.out.C4, "outcome.var") <- x
      
      res.C4[[x]] <- PanelEstimate(sets = m.out.C4,
                     panel.data = panel_dfC4,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)
}

for (x in outcome_vars){
      file_name <- paste0("estimates/c4", x, ".png")
      print(file_name)
      png(file_name)
      plot(res.C4[[x]], main = paste0("Estimated Effects (", x, ") - ", match_methodC4))
      dev.off()
}

file_name <- paste0(estimates_save_nameC4, save_name, ".xlsx")
print(file_name)
estimate_summaries.C4 <- lapply(res.C4, summary)
estimate_summaries.C4 <- lapply(estimate_summaries.C4, as.data.frame)
write_xlsx(estimate_summaries.C4, file_name)
```

## Choosing the number of SVC
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(fields)
library(MASS)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# These are some socioeconomic covariates we may evaluate balance for but need to convert from string to num
covs_to_check <- c("white_perc", "bachelors_perc", "poverty_perc", "unemployment", "med_family_income_adj", "med_household_income_adj", "total_population")
df <- df |> mutate(across(all_of(covs_to_check), as.numeric)) # make sure these are numeric

# These are the variables we will assess balance for
se_vars <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population")
other_vars <- c("TOT_REV", "TOT_ASSET")
ntee_vars <- c("NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA")

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA_nDis", "lat", "lng"), se_vars, other_vars, ntee_vars)))

df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
                                       .default = TOT_REV),
                   TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
                                       .default = TOT_ASSET))
df <- df |> mutate(across(c("TOT_REV", "TOT_ASSET"), log1p)) #log1p adds 1 first to avoid 0s

library(data.table)
df <- arrange(df, TAX_YEAR)
df.lag <- shift(df[,7:ncol(df)], n = 1:3, give.names = TRUE)
df <- bind_cols(df, df.lag)
rm(df.lag)

df <- df |> filter(TAX_YEAR >= 1994)

df_sub <- df |> dplyr::select(CountyID, lat, lng) |> unique()

x <- cbind(df_sub$lat, df_sub$lng) # inputs: latitude and longitude

# SVC
d <- rdist(x)
K <- Matern(d, smoothness=.5) # for smoothness paramter, should we do some sort of optimization?
E <- eigen(K)

options(repr.plot.width = 6, repr.plot.height = 4)
lls = list()
ls_list = list()
for (num_svc in seq(10,100,10)){
      curr_name <- as.character(num_svc*2)
      x_s <- cbind(x[,1] * E$vectors[,1:num_svc], x[,2] * E$vectors[,1:num_svc]) # would have to play around with this number
      
      # add SVC back into panel data
      xs_vars <- paste0("x_s", seq_len(ncol(x_s)))
      colnames(x_s) <- xs_vars
      df_sub.temp <- cbind(df_sub, x_s)
      
      df.temp <- merge(df, df_sub.temp |> dplyr::select(-lat, -lng), by = "CountyID", all = TRUE)

      upper_formula <- paste(
        "treat.FEMA_nDis ~ TAX_YEAR + bachelors_perc + med_household_income_adj + white_perc + total_population +",
        paste0(do.call(paste0, expand.grid(paste0(se_vars, "_lag_"), 1:3)), collapse = " + "), "+",
        paste0(do.call(paste0, expand.grid(paste0(ntee_vars, "_lag_"), 1:3)), collapse = " + "), "+",
        paste0(do.call(paste0, expand.grid(paste0(other_vars, "_lag_"), 1:3)), collapse = " + "), "+",
        paste(xs_vars, collapse = " + "))
      
      # regression
      ls_list[[curr_name]] <- glm(as.formula(upper_formula), data = df.temp, family = binomial(link = "logit"))
      #summary(ls)

      yweight <- predict(ls_list[[curr_name]], type="response")
      png(paste0("svc_plots/no_interference/svc", num_svc*2, "_Mar19.png"))
      plot(sort(yweight), main = paste("# SVC =", num_svc*2), cex = 0.10)
      dev.off()
} # winner was num_svc = 30 on March 10, but on March 19 200 is the winner by a landslide

library(lmtest)
my_test <- lrtest(ls_list[["100"]], ls_list[["200"]])
my_test

AIC(ls_list[["100"]], ls_list[["200"]])
BIC(ls_list[["100"]], ls_list[["200"]])

y_pred <- predict(ls_list[["200"]], df.temp, type = "response")
quantile(y_pred, na.rm = TRUE, probs = seq(0,1,0.1)) # all probs between 0.02 and 0.95
# sum(near(y_pred, 1), na.rm = TRUE)
# sort(y_pred, decreasing = FALSE) 
```

### Determining units that violate positivity assumption
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(data.table)

panel_df <- readRDS("panel_data/no_int_panelV2.rds")

se_vars <- colnames(panel_df)[which(colnames(panel_df) == "bachelors_perc"):which(colnames(panel_df) == "total_population")] 
org_vars <- colnames(panel_df)[which(colnames(panel_df) == "TOT_REV"):which(colnames(panel_df) == "NTEEV2_NA")] 
svc_vars <- colnames(panel_df)[which(colnames(panel_df) == "x_s1"):ncol(panel_df)]

panel_df  <- arrange(panel_df, TAX_YEAR)
df.lag <- shift(panel_df[,5:which(colnames(panel_df) == "NTEEV2_NA")], n = 1:3, give.names = TRUE)
panel_df <- bind_cols(panel_df, df.lag)
rm(df.lag)

vars_lagged <- colnames(panel_df)[which(colnames(panel_df) == "bachelors_perc_lag_1"):which(colnames(panel_df) == "NTEEV2_NA_lag_3")] 

panel_df <- panel_df |> filter(TAX_YEAR >= 1994 & TAX_YEAR <= 2020)

glm_formula <- paste("treat.FEMA ~ TAX_YEAR +", 
                     paste0(c(se_vars, org_vars), collapse = " + "), "+",
                     paste0(vars_lagged, collapse = " + "), "+",
                     paste(svc_vars, collapse = " + "))
      
# regression
ls <- glm(as.formula(glm_formula), data = panel_df, family = binomial(link = "logit"))
summary(ls)

yweight <- predict(ls, type="response")

# png(filename = paste0("svc_plots/int/svc", num_svc*2, ".png"))
plot(sort(yweight), main = paste("# SVC =", 100), cex = 0.10)
# dev.off()
min(yweight) # 0.000079
max(yweight) # 0.912
quantile(yweight) # 25-75% between 0.212 and 0.399 -- median is 0.294

prob0 <- names(yweight[yweight < 0.001]) # its two counties 2247 and 2248
as.data.frame(panel_df[prob0,] |> arrange(CountyID) |> select(CountyID, TAX_YEAR))
prob0_IDs <- panel_df[prob0,] |> select(CountyID) |> unique()

library(PanelMatch)
panel_sub <- panel_df |> filter(CountyID %in% prob0_IDs$CountyID) |> select(CountyID, TAX_YEAR, treat.FEMA, TOT_REV, geoid_2010) |> arrange(CountyID) |> 
      group_by(CountyID) |>
      mutate(ID = as.integer(cur_group_id())) |>
      ungroup()

panel_sub <- PanelData(panel.data = as.data.frame(panel_sub),
                      unit.id = "ID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA",
                      outcome = "TOT_REV")
DisplayTreatment(panel.data = panel_sub, 
                                 color.of.treated = "#DC267F",
                                 color.of.untreated = "#648FFF",
                                 title = "Treatment Distribution Across Units and Time",
                                 xlab = "Tax Year", ylab = "County", 
                                 legend.position = "right",
                                 legend.labels = c("untreated", "treated"),
                                 hide.x.tick.label = FALSE, hide.y.tick.label = FALSE)
```

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(data.table)

panel_df <- readRDS("panel_data/no_int_panelV3.rds")

se_vars <- colnames(panel_df)[which(colnames(panel_df) == "bachelors_perc"):which(colnames(panel_df) == "total_population")] 
org_vars <- colnames(panel_df)[which(colnames(panel_df) == "TOT_REV"):which(colnames(panel_df) == "NTEEV2_NA")] 
svc_vars <- colnames(panel_df)[which(colnames(panel_df) == "x_s1"):ncol(panel_df)]

panel_df  <- arrange(panel_df, TAX_YEAR)
df.lag <- shift(panel_df[,5:which(colnames(panel_df) == "NTEEV2_NA")], n = 1:3, give.names = TRUE)
panel_df <- bind_cols(panel_df, df.lag)
rm(df.lag)

vars_lagged <- colnames(panel_df)[which(colnames(panel_df) == "bachelors_perc_lag_1"):which(colnames(panel_df) == "NTEEV2_NA_lag_3")] 

panel_df <- panel_df |> filter(TAX_YEAR >= 1994 & TAX_YEAR <= 2020)

glm_formula <- paste("treat.FEMA ~ TAX_YEAR +", 
                     paste0(c(se_vars, org_vars), collapse = " + "), "+",
                     paste0(vars_lagged, collapse = " + "), "+",
                     paste(svc_vars, collapse = " + "))
      
# regression
ls <- glm(as.formula(glm_formula), data = panel_df, family = binomial(link = "logit"))
summary(ls)

yweight <- predict(ls, type="response")

# png(filename = paste0("svc_plots/int/svc", num_svc*2, ".png"))
plot(sort(yweight), main = paste("# SVC =", 100), cex = 0.10)
# dev.off()
min(yweight) # 0
max(yweight) # 0.912
quantile(yweight) # 25-75% between 0.212 and 0.399 -- median is 0.294

prob0 <- names(yweight[yweight < 0.001]) # 2247, 2248, 2544, 2545, 2546
as.data.frame(panel_df[prob0,] |> arrange(CountyID) |> select(CountyID, TAX_YEAR))
prob0_IDs <- panel_df[prob0,] |> select(CountyID) |> unique()

library(PanelMatch)
panel_sub <- panel_df |> filter(CountyID %in% prob0_IDs$CountyID) |> select(CountyID, TAX_YEAR, treat.FEMA, TOT_REV, geoid_2010) |> arrange(CountyID) |> 
      group_by(CountyID) |>
      mutate(ID = as.integer(cur_group_id())) |>
      ungroup()
panel_sub <- PanelData(panel.data = as.data.frame(panel_sub),
                      unit.id = "ID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA",
                      outcome = "TOT_REV")
DisplayTreatment(panel.data = panel_sub, 
                                 color.of.treated = "#DC267F",
                                 color.of.untreated = "#648FFF",
                                 title = "Treatment Distribution Across Units and Time",
                                 xlab = "Tax Year", ylab = "County", 
                                 legend.position = "right",
                                 legend.labels = c("untreated", "treated"),
                                 hide.x.tick.label = TRUE, hide.y.tick.label = FALSE)
```

```{r}
library(PanelMatch)

L = 3
folder_path_match <- "matches/with_svc/no_int_panelV3/config4/" # "matches/with_svc/no_int_panelV2/config3/"
save_name <- paste0("Lag", L)
match_method <- "none"

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))
m.out <- matches_list[[match_method]]
summary(m.out)

matched_sets <- extract(m.out)

treated <- names(matched_sets)

my_list <- c("2247.2011", "2247.2012", "2247.2013", "2247.2014", "2247.2015", "2247.2019", "2247.2020",
             "2248.2011", "2248.2012", "2248.2013", "2248.2014", "2248.2015", "2248.2019", "2248.2020",
             "2544.2019", "2545.2019", "2546.2019")

intersect(treated, my_list) # empty!!! well duh... would only show up in control units

matched_controls <- weights(matched_sets)
matched_controls <- lapply(matched_controls, function(x) x[x > 0])
untreated <- lapply(matched_controls, function(x) names(x))
untreated <- unlist(untreated, use.names = FALSE)
untreated <- unique(untreated)

intersect(untreated, my_list) # empty... makes sense that treated units probably all have nonzero propensities

my_list <- c("2247.2012", "2247.2013", "2247.2014", "2247.2015", "2247.2016", "2247.2020", "2247.2021",
             "2248.2012", "2248.2013", "2248.2014", "2248.2015", "2248.2016", "2248.2020", "2248.2021",
             "2544.2020", "2545.2020", "2546.2020")

intersect(treated, my_list) # empty!!! well duh... would only show up in control units
```

# PanelMatch - Interference
## Setup
```{r}
library(data.table)
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)
library(geosphere)
library(fastDummies)
source("matches_helper.R")

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# make sure these are numeric
df <- df |> mutate(across(all_of(c("bachelors_perc", "med_household_income_adj", "white_perc")), as.numeric)) 

# Add REGION dummy columns
df <- df |> dummy_cols(select_columns = c("REGION"), ignore_na = FALSE, remove_selected_columns = TRUE, remove_first_dummy = FALSE)

# These are the variables we will adjust for
adj_set <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET","TOT_EXP", "lat", "lng")

# Other variables
ntee_vars <- grep("^NTEEV2_", names(df), value = TRUE)
region_vars <- grep("^REGION_", names(df), value = TRUE)

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA"), adj_set, ntee_vars, region_vars)))

# Log transformations
# saveRDS(df |> filter(TOT_REV < 0 | TOT_ASSET < 0 | TOT_EXP < 0), "misc_analysis/intV2_negative.rds")
df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
                                       .default = TOT_REV),
                   TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
                                       .default = TOT_ASSET),
                   TOT_EXP = case_when(TOT_EXP < 0 ~ 0,
                                       .default = TOT_EXP))

df <- df |> mutate(TOT_REV = log1p(TOT_REV),
                   TOT_ASSET = log1p(TOT_ASSET),
                   TOT_EXP = log1p(TOT_EXP),
                   med_household_income_adj = log1p(med_household_income_adj),
                   total_population = log(total_population))

# Create new treatment variable that takes a distance-based cutoff adjustment
adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 25)

res <- add_exposed(panel_data = df, adj_list = adj_lists$union, treatment = "treat.FEMA")

df <- res$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))

per_year_summary <- df |> group_by(TAX_YEAR) |>
      summarise(
            n_1 = sum(treat.FEMA.adj == 1, na.rm = TRUE),
            n_0 = sum(treat.FEMA.adj == 0, na.rm = TRUE),
            n = n())

paste("Treated")
quantile(per_year_summary$n_1, probs = seq(0,1,0.1))

paste("Control")
quantile(per_year_summary$n_0, probs = seq(0,1,0.1))

paste("Total")
quantile(per_year_summary$n, probs = seq(0,1,0.1))

# 1859 is New York County
# df$treat.FEMA.adj[((df$CountyID == 1859) & (df$TAX_YEAR == 2001))] <- NA # note they were untreated before so don't have to correct above
# df[((df$CountyID == 1859) & (df$TAX_YEAR == 2001)),]

# Adding spatially varying coefficients
# df <- add_SVC(df, num_E_vec = 50)
df$treat.FEMA <- NULL

# https://stackoverflow.com/questions/74815244/panelmatch-r-errors-please-convert-time-id-to-consecutive-integers-and-pleas
panel_df <- PanelData(panel.data = as.data.frame(df),
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA.adj",
                      outcome = "TOT_REV")

# rm(df)
# saveRDS(panel_df, "panel_data/int_panelV1.rds") # No SVC, exposed treatment = 1
# saveRDS(panel_df, "panel_data/int_panelV2.rds") # With SVC, exposed treatment = 1
# saveRDS(panel_df, "panel_data/int_panelV3.rds") # V2 but no log transformations

DisplayTreatment(panel.data = panel_df, 
                  color.of.treated = "#DC267F", #"#FFB000"
                  color.of.untreated = "#648FFF", #"#785EF0"
                  title = "Exposure Distribution Across Units and Time",
                  xlab = "Tax Year", ylab = "County", 
                  legend.position = "right",
                  legend.labels = c("unexposed", "exposed"),
                  hide.x.tick.label = TRUE, hide.y.tick.label = TRUE, 
                  dense.plot = TRUE)
```

## Matching Methods - Config 3, With SVC
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/int_panelV2.rds")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/int_panelV2/config3/"
folder_path_balance <- "balances/with_svc/int_panelV2/config3/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
# write.csv(as.data.frame(summary(matches_list$none)), "match_summary.csv")
summary(matches_list$none)
summary(matches_list$ps.match)
# summary(matches_list$CBPS.match)
# summary(matches_list$ps.weight)
# summary(matches_list$CBPS.weight)
# 
plot(matches_list$none)
```


Default Matching Weights Plots
```{r}
# plot(extract(matches_list$CBPS.weight), panel.data = panel_df)
```


Default Scatter Plots
```{r}
# plot(balances_list$none, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE, 
#      main = "Standardized Mean Difference of Covariates (No Matching)")
# plot(balances_list$mahalanobis, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
#      ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```

My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "TOT_EXP", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.17,0.17), "unrefined" = c(-0.15, 0.05), "all" = c(-0.17, 0.17))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

p_none$all
# p_ps.match$all_unrefined
# p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
p_cbps.weight$all
# 
p_none$base
# p_ps.match$base_unrefined
# p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
p_cbps.weight$base
# 
p_none$ntee
# p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
p_cbps.weight$ntee
# 
p_none$loc
# p_ps.match$loc_unrefined
# p_ps.match$loc
# p_cbps.match$loc
# p_ps.weight$loc
p_cbps.weight$loc

library(cowplot)
tmp <- p_none$all + theme(legend.position = "bottom")
g <- ggplotGrob(tmp)
leg <- g$grobs[[which(g$layout$name == "guide-box-bottom")]]

prow <- plot_grid(p_none$all + theme(legend.position="none"), p_cbps.weight$all + theme(legend.position="none"), labels = "AUTO", hjust = -1, nrow = 1)
plot_grid(prow, leg, ncol = 1, rel_heights = c(1, .1))
```

```{r}
q_none <- quantile(signif(abs(bal_none$value[bal_none$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_ps.mat <- quantile(signif(abs(bal_ps.match$value[bal_ps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_cbps.mat <- quantile(signif(abs(bal_cbps.match$value[bal_cbps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_ps.w <- quantile(signif(abs(bal_ps.weight$value[bal_ps.weight$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_cbps.w <- quantile(signif(abs(bal_cbps.weight$value[bal_cbps.weight$unrefined == FALSE]), 3), probs = seq(0,1,0.1))

df <- cbind(as.data.frame(q_none), as.data.frame(q_ps.mat), as.data.frame(q_cbps.mat), as.data.frame(q_ps.w), as.data.frame(q_cbps.w))

df$winner <- colnames(df)[max.col(-df, ties.method = "first")]

write.csv(df, file = paste0(folder_path_balance, save_name, "deciles.csv"))

write.csv(bal_none |> filter(unrefined == FALSE) |> select(-unrefined) |>
                pivot_wider(names_from = covariate, values_from = value) |>
                mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
          file = "balance.csv")

write.csv(bal_cbps.weight |> filter(unrefined == FALSE) |> select(-unrefined) |>
                pivot_wider(names_from = covariate, values_from = value) |>
                mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
          file = "balance.csv")
```

## Matching Methods - Config 4, With SVC
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/int_panelV2.rds")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/int_panelV3/config4/"
folder_path_balance <- "balances/with_svc/int_panelV3/config4/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
# write.csv(as.data.frame(summary(matches_list$none)), "match_summary.csv")
summary(matches_list$none)
summary(matches_list$ps.match)
# summary(matches_list$CBPS.match)
# summary(matches_list$ps.weight)
# summary(matches_list$CBPS.weight)
# 
plot(matches_list$none)
```


Default Matching Weights Plots
```{r}
# plot(extract(matches_list$CBPS.weight), panel.data = panel_df)
```


Default Scatter Plots
```{r}
# plot(balances_list$none, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE, 
#      main = "Standardized Mean Difference of Covariates (No Matching)")
# plot(balances_list$mahalanobis, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
#      ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```

My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "TOT_EXP", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.17,0.17), "unrefined" = c(-0.15, 0.05), "all" = c(-0.17, 0.17))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

p_none$all
# p_ps.match$all_unrefined
# p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
p_cbps.weight$all
# 
# p_none$base
# p_ps.match$base_unrefined
# p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
p_cbps.weight$base
# 
# p_none$ntee
# p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
p_cbps.weight$ntee
# 
# p_none$loc
# p_ps.match$loc_unrefined
# p_ps.match$loc
# p_cbps.match$loc
# p_ps.weight$loc
p_cbps.weight$loc

```

```{r}
q_none <- quantile(signif(abs(bal_none$value[bal_none$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_ps.mat <- quantile(signif(abs(bal_ps.match$value[bal_ps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_cbps.mat <- quantile(signif(abs(bal_cbps.match$value[bal_cbps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_ps.w <- quantile(signif(abs(bal_ps.weight$value[bal_ps.weight$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_cbps.w <- quantile(signif(abs(bal_cbps.weight$value[bal_cbps.weight$unrefined == FALSE]), 3), probs = seq(0,1,0.1))

df <- cbind(as.data.frame(q_none), as.data.frame(q_ps.mat), as.data.frame(q_cbps.mat), as.data.frame(q_ps.w), as.data.frame(q_cbps.w))

df$winner <- colnames(df)[max.col(-df, ties.method = "first")]

write.csv(df, file = paste0(folder_path_balance, save_name, "deciles.csv"))

write.csv(bal_none |> filter(unrefined == FALSE) |> select(-unrefined) |>
                pivot_wider(names_from = covariate, values_from = value) |>
                mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
          file = "balance.csv")

write.csv(bal_cbps.weight |> filter(unrefined == FALSE) |> select(-unrefined) |>
                pivot_wider(names_from = covariate, values_from = value) |>
                mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
          file = "balance.csv")
```

## Matching Methods - Config 3, NO SVC
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# panel_df <- readRDS("panel_data/int_panelV1.rds")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/no_svc/int_panelV1/config3/"
folder_path_balance <- "balances/no_svc/int_panelV1/config3/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))

# Retrieve the PanelBalance objects
balances_list <- readRDS(paste0(folder_path_balance, save_name, ".rds"))

# Make the balances for each lag easy to work with as a dataframe
bal_df <- lapply(balances_list, summary)
bal_df <- lapply(bal_df, as.data.frame)
bal_df <- lapply(bal_df, tibble::rownames_to_column, var = "time")

# Print information about matching and balances
# write.csv(as.data.frame(summary(matches_list$none)), "match_summary.csv")
summary(matches_list$none)
summary(matches_list$ps.match)
# summary(matches_list$CBPS.match)
# summary(matches_list$ps.weight)
# summary(matches_list$CBPS.weight)
# 
plot(matches_list$none)
```


Default Matching Weights Plots
```{r}
# plot(extract(matches_list$CBPS.weight), panel.data = panel_df)
```


Default Scatter Plots
```{r}
# plot(balances_list$none, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE, 
#      main = "Standardized Mean Difference of Covariates (No Matching)")
# plot(balances_list$mahalanobis, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
#      ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$ps.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Matching)")
plot(balances_list$CBPS.match, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Matching)")
plot(balances_list$ps.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (PS Weighting)")
plot(balances_list$CBPS.weight, type = "scatter", include.unrefined.panel = TRUE, legend = TRUE,
     ylim = c(-0.01,.1), main = "Standardized Mean Difference of Covariates (CBPS Weighting)")
```


Default Covariate Balance Plots
```{r}
# plot(balances_list$none, type = "panel", include.unrefined.panel = FALSE, legend = TRUE, main = "No Matching") 
# plot(balances_list$ps.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Matching", main.unrefined = "PS Matching (unrefined)") 
# plot(balances_list$CBPS.match, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Matching", main.unrefined = "CBPS Matching (unrefined)") 
# plot(balances_list$ps.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "PS Weighting", main.unrefined = "PS Weighting (unrefined)") 
# plot(balances_list$CBPS.weight, type = "panel", include.unrefined.panel = TRUE, legend = FALSE, main = "CBPS Weighting", main.unrefined = "CBPS Weighting (unrefined)") 
```

My Covariate Balance Plots
```{r}
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "TOT_EXP", "NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA", "REGION_MIDWEST", "REGION_NORTHEAST", "REGION_SOUTH", "REGION_WEST")

bal_none <- tidy_bal_df(bal_data = bal_df$none, covariates = paste0("m.out.",covs_to_check))
bal_ps.match <- tidy_bal_df(bal_data = bal_df$ps.match, covariates = paste0("m.out.",covs_to_check))
bal_cbps.match <- tidy_bal_df(bal_data = bal_df$CBPS.match, covariates = paste0("m.out.",covs_to_check))
bal_ps.weight <- tidy_bal_df(bal_data = bal_df$ps.weight, covariates = paste0("m.out.",covs_to_check))
bal_cbps.weight <- tidy_bal_df(bal_data = bal_df$CBPS.weight, covariates = paste0("m.out.",covs_to_check))
```

```{r}
ylims <- list("base" = c(-0.15, 0.05), "ntee" = c(-0.15, 0.05), "loc" = c(-0.17,0.17), "unrefined" = c(-0.15, 0.05), "all" = c(-0.17, 0.17))
p_none <- no_int_no_SVC_plots(bal_none,
                              match_method = "No Matching",
                              ylim_base = ylims[["base"]], 
                              ylim_ntee = ylims[["ntee"]],
                              ylim_loc = ylims[["loc"]],
                              ylim_all = ylims[["all"]],
                              include.unrefined.panel = FALSE)
p_ps.match <- no_int_no_SVC_plots(bal_ps.match, 
                                  match_method = "PS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  ylim_unrefined = ylims[["unrefined"]],
                                  include.unrefined.panel = TRUE)
p_cbps.match <- no_int_no_SVC_plots(bal_cbps.match,
                                  match_method = "CBPS Matching",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_ps.weight <- no_int_no_SVC_plots(bal_ps.weight,
                                  match_method = "PS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)
p_cbps.weight <- no_int_no_SVC_plots(bal_cbps.weight,
                                  match_method = "CBPS Weighting",
                                  ylim_base = ylims[["base"]], 
                                  ylim_ntee = ylims[["ntee"]],
                                  ylim_loc = ylims[["loc"]],
                                  ylim_all = ylims[["all"]],
                                  include.unrefined.panel = FALSE)

options(repr.plot.width = 6, repr.plot.height = 4)

p_none$all
# p_ps.match$all_unrefined
# p_ps.match$all
# p_cbps.match$all
# p_ps.weight$all
p_cbps.weight$all
# 
p_none$base
# p_ps.match$base_unrefined
# p_ps.match$base
# p_cbps.match$base
# p_ps.weight$base
p_cbps.weight$base
# 
p_none$ntee
# p_ps.match$ntee_unrefined
# p_ps.match$ntee
# p_cbps.match$ntee
# p_ps.weight$ntee
p_cbps.weight$ntee
# 
p_none$loc
# p_ps.match$loc_unrefined
# p_ps.match$loc
# p_cbps.match$loc
# p_ps.weight$loc
p_cbps.weight$loc

```

```{r}
q_none <- quantile(signif(abs(bal_none$value[bal_none$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_ps.mat <- quantile(signif(abs(bal_ps.match$value[bal_ps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_cbps.mat <- quantile(signif(abs(bal_cbps.match$value[bal_cbps.match$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_ps.w <- quantile(signif(abs(bal_ps.weight$value[bal_ps.weight$unrefined == FALSE]), 3), probs = seq(0,1,0.1))
q_cbps.w <- quantile(signif(abs(bal_cbps.weight$value[bal_cbps.weight$unrefined == FALSE]), 3), probs = seq(0,1,0.1))

df <- cbind(as.data.frame(q_none), as.data.frame(q_ps.mat), as.data.frame(q_cbps.mat), as.data.frame(q_ps.w), as.data.frame(q_cbps.w))

df$winner <- colnames(df)[max.col(-df, ties.method = "first")]

write.csv(df, file = paste0(folder_path_balance, save_name, "deciles.csv"))

write.csv(bal_none |> filter(unrefined == FALSE) |> select(-unrefined) |>
                pivot_wider(names_from = covariate, values_from = value) |>
                mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
          file = "balance.csv")

write.csv(bal_cbps.weight |> filter(unrefined == FALSE) |> select(-unrefined) |>
                pivot_wider(names_from = covariate, values_from = value) |>
                mutate(across(where(is.numeric), ~ round(.x, digits = 5))),
          file = "balance.csv")
```




## Estimates
### Config 3 versus Config 4 - Revenue, No Stratification

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(tidyr)

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
save_name <- paste0("Lag", L)

folder_pathC3 <- "matches/with_svc/int_panelV2/config3/"
folder_pathC4 <- "matches/with_svc/int_panelV3/config4/"

# Retrieve the PanelMatch objects
matches_listC3 <- readRDS(paste0(folder_pathC3, save_name, ".rds"))
matches_listC4 <- readRDS(paste0(folder_pathC4, save_name, ".rds"))

# Retrieve the PanelData object
panel_dfC3 <- readRDS("panel_data/int_panelV2.rds")
panel_dfC4 <- readRDS("panel_data/int_panelV3.rds")

estimates_save_nameC3 <- "estimates/with_svc_intV2_con3_"
estimates_save_nameC4 <- "estimates/with_svc_intV3_con4_"

# No Matching
m.out.noneC3 <- matches_listC3$none
m.out.noneC4 <- matches_listC4$none
# attr(m.out.none, "outcome") <- "EXP_LN_REV"
# attr(m.out.none, "outcome.var") <- "EXP_LN_REV"

match_methodC3 <- "CBPS.weight"
match_methodC4 <- "CBPS.weight"

m.out.C3 <- matches_listC3[[match_methodC3]]
m.out.C4 <- matches_listC4[[match_methodC4]]

# outcome_vars <- c("TOT_REV", "REV_ART", "REV_EDU", "REV_HMS", "REV_PSB", "REV_ENV", "REV_HEL", "REV_HOS", "REV_UNI", "REV_MMB", "REV_ENV") # REV_REL, REV_IFA, REV_UNU, REV_NA

# outcome_vars <- c("REV_SIZE.1", "REV_SIZE.2", "REV_SIZE.3", "REV_SIZE.4", "REV_SIZE.5", "REV_SIZE.6") # "REV_SIZE.NA"
outcome_vars <- c("TOT_REV")

panel_dfC4 <- panel_dfC4 |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 1, .default = TOT_REV))
```

```{r}
# Estimates under no matching
res_no.match.C3 <- PanelEstimate(sets = m.out.noneC3,
                     panel.data = panel_dfC3,
                     se.method = "conditional", # "bootstrap" "unconditional"
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

res_no.match.C4 <- PanelEstimate(sets = m.out.noneC4,
                     panel.data = panel_dfC4,
                     se.method = "unconditional",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

# Estimates under CBPS Weighting
res.C3 <- PanelEstimate(sets = m.out.C3,
                     panel.data = panel_dfC3,
                     se.method = "conditional",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

res.C4 <- PanelEstimate(sets = m.out.C4,
                     panel.data = panel_dfC4,
                     se.method = "unconditional",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)
```

```{r}
plot(res_no.match.C3, main = "Estimated Effects (LOG REV) - No Matching")
summary(res_no.match.C3)
# hist(res_no.match.C3$bootstrapped.estimates) 
# as.data.frame(summary(res_no.match.C3)) |> mutate(across(where(is.numeric), ~ exp(.x)-1))

plot(res_no.match.C4, main = "Estimated Effects (REV) - No Matching")
summary(res_no.match.C4)
# hist(res_no.match.C4$bootstrapped.estimates)

plot(res.C3, main = "Estimated Effects (LOG REV) - CBPS Weighting")
summary(res.C3)
write.csv(as.data.frame(summary(res.C3)) |> mutate(across(where(is.numeric), ~ round(exp(.x), 5))), "res.csv")
# hist(res.C3$bootstrapped.estimates)

plot(res.C4, main = "Estimated Effects (REV) - CBPS Weighting")
summary(res.C4)
# hist(res.C4$bootstrapped.estimates)
```



### Config 3 versus Config 4 - Assets, No Stratification

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(tidyr)

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
save_name <- paste0("Lag", L)

folder_pathC3 <- "matches/with_svc/int_panelV2/config3/"
folder_pathC4 <- "matches/with_svc/int_panelV3/config4/"

# Retrieve the PanelMatch objects
matches_listC3 <- readRDS(paste0(folder_pathC3, save_name, ".rds"))
matches_listC4 <- readRDS(paste0(folder_pathC4, save_name, ".rds"))

# Retrieve the PanelData object
panel_dfC3 <- readRDS("panel_data/int_panelV2.rds")
panel_dfC4 <- readRDS("panel_data/int_panelV3.rds")

estimates_save_nameC3 <- "estimates/with_svc_intV2_con3_"
estimates_save_nameC4 <- "estimates/with_svc_intV3_con4_"

# No Matching
m.out.noneC3 <- matches_listC3$none
m.out.noneC4 <- matches_listC4$none

match_methodC3 <- "CBPS.weight"
match_methodC4 <- "CBPS.weight"

m.out.C3 <- matches_listC3[[match_methodC3]]
m.out.C4 <- matches_listC4[[match_methodC4]]

attr(m.out.noneC3, "outcome") <- "TOT_ASSET"
attr(m.out.noneC3, "outcome.var") <- "TOT_ASSET"
attr(m.out.noneC4, "outcome") <- "TOT_ASSET"
attr(m.out.noneC4, "outcome.var") <- "TOT_ASSET"
attr(m.out.C3, "outcome") <- "TOT_ASSET"
attr(m.out.C3, "outcome.var") <- "TOT_ASSET"
attr(m.out.C4, "outcome") <- "TOT_ASSET"
attr(m.out.C4, "outcome.var") <- "TOT_ASSET"
```

```{r}
# Estimates under no matching
res_no.match.C3 <- PanelEstimate(sets = m.out.noneC3,
                     panel.data = panel_dfC3,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

res_no.match.C4 <- PanelEstimate(sets = m.out.noneC4,
                     panel.data = panel_dfC4,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

# Estimates under CBPS Weighting
res.C3 <- PanelEstimate(sets = m.out.C3,
                     panel.data = panel_dfC3,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

res.C4 <- PanelEstimate(sets = m.out.C4,
                     panel.data = panel_dfC4,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)
```

```{r}
plot(res_no.match.C3, main = "Estimated Effects (LOG ASSETS) - No Matching")
summary(res_no.match.C3)
hist(res_no.match.C3$bootstrapped.estimates) 
# as.data.frame(summary(res_no.match.C3)) |> mutate(across(where(is.numeric), ~ exp(.x)-1))

plot(res_no.match.C4, main = "Estimated Effects (ASSETS) - No Matching")
summary(res_no.match.C4)
hist(res_no.match.C4$bootstrapped.estimates)

plot(res.C3, main = "Estimated Effects (LOG ASSETS) - CBPS Weighting")
summary(res.C3)
hist(res.C3$bootstrapped.estimates)

plot(res.C4, main = "Estimated Effects (ASSETS) - CBPS Weighting")
summary(res.C4)
hist(res.C4$bootstrapped.estimates)
```


### Config 3 versus Config 4 - With Stratification by SIZE

```{r, message = F}
library(data.table)
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(fastDummies)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# outcome_vars <- grep("^REV_", names(df), value = TRUE)
outcome_vars <- c("REV_SIZE.1", "REV_SIZE.2", "REV_SIZE.3", "REV_SIZE.4", "REV_SIZE.5", "REV_SIZE.6")

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR"), outcome_vars))) 

panel_dfC4 <- readRDS("panel_data/int_panelV3.rds")
df <- merge(panel_dfC4, df, by = c("CountyID", "TAX_YEAR"))

df <- as.data.frame(df) |> select(all_of(c(c("CountyID", "TAX_YEAR", "TOT_REV", "treat.FEMA.adj"), outcome_vars)))
panel_dfC4 <- PanelData(panel.data = df,
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA.adj",
                      outcome = "TOT_REV")

# Log transformations
panel_dfC3 <- panel_dfC4 |> mutate(across(all_of(outcome_vars), ~ if_else(.x < 0, 0, .x)))
panel_dfC3 <- panel_dfC3 |> mutate(across(all_of(outcome_vars), log1p))
rm(df)
```

```{r}
# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
save_name <- paste0("Lag", L)

folder_pathC3 <- "matches/with_svc/int_panelV2/config3/"
folder_pathC4 <- "matches/with_svc/int_panelV3/config4/"

# Retrieve the PanelMatch objects
matches_listC3 <- readRDS(paste0(folder_pathC3, save_name, ".rds"))
matches_listC4 <- readRDS(paste0(folder_pathC4, save_name, ".rds"))

estimates_save_nameC3 <- "estimates/with_svc_intV2_con3_"
estimates_save_nameC4 <- "estimates/with_svc_intV3_con4_"

match_methodC3 <- "CBPS.weight"
match_methodC4 <- "CBPS.weight"

m.out.C3 <- matches_listC3[[match_methodC3]]
m.out.C4 <- matches_listC4[[match_methodC4]]

res.C3 <- list()
for (x in outcome_vars){
      print(x)
      attr(panel_dfC3, "outcome") <- x
      attr(m.out.C3, "outcome") <- x
      attr(m.out.C3, "outcome.var") <- x
      
      res.C3[[x]] <- PanelEstimate(sets = m.out.C3,
                     panel.data = panel_dfC3,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)
}

for (x in outcome_vars){
      file_name <- paste0("estimates/interference/c3", x, ".png")
      print(file_name)
      png(file_name)
      plot(res.C3[[x]], main = paste0("Estimated Effects (LOG ", x, ") - ", match_methodC3))
      dev.off()
}

file_name <- paste0(estimates_save_nameC3, save_name, ".xlsx")
print(file_name)
estimate_summaries.C3 <- lapply(res.C3, summary)
estimate_summaries.C3 <- lapply(estimate_summaries.C3, as.data.frame)
write_xlsx(estimate_summaries.C3, file_name)

res.C4 <- list()
for (x in outcome_vars){
      print(x)
      attr(panel_dfC4, "outcome") <- x
      attr(m.out.C4, "outcome") <- x
      attr(m.out.C4, "outcome.var") <- x
      
      res.C4[[x]] <- PanelEstimate(sets = m.out.C4,
                     panel.data = panel_dfC4,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)
}

for (x in outcome_vars){
      file_name <- paste0("estimates/interference/c4", x, ".png")
      print(file_name)
      png(file_name)
      plot(res.C4[[x]], main = paste0("Estimated Effects (", x, ") - ", match_methodC4))
      dev.off()
}

file_name <- paste0(estimates_save_nameC4, save_name, ".xlsx")
print(file_name)
estimate_summaries.C4 <- lapply(res.C4, summary)
estimate_summaries.C4 <- lapply(estimate_summaries.C4, as.data.frame)
write_xlsx(estimate_summaries.C4, file_name)
```

### Config 3 - Number of nonprofits
```{r}
ntee_vars <- colnames(panel_dfC3)[which(colnames(panel_dfC3) == "NTEEV2_ART"):which(colnames(panel_dfC3) == "NTEEV2_NA")] 
num_non <- panel_dfC3 |> select(all_of(ntee_vars)) |> rowSums(na.rm = TRUE)
panel_dfC3$NUM_NONPROFITS <- log1p(num_non)
```


```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(tidyr)

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
save_name <- paste0("Lag", L)

folder_pathC3 <- "matches/with_svc/int_panelV2/config3/"
folder_pathC4 <- "matches/with_svc/int_panelV3/config4/"

# Retrieve the PanelMatch objects
matches_listC3 <- readRDS(paste0(folder_pathC3, save_name, ".rds"))
matches_listC4 <- readRDS(paste0(folder_pathC4, save_name, ".rds"))

# Retrieve the PanelData object
panel_dfC3 <- readRDS("panel_data/int_panelV2.rds")
panel_dfC4 <- readRDS("panel_data/int_panelV3.rds")

estimates_save_nameC3 <- "estimates/with_svc_intV2_con3_"
estimates_save_nameC4 <- "estimates/with_svc_intV3_con4_"

# No Matching
m.out.noneC3 <- matches_listC3$none
m.out.noneC4 <- matches_listC4$none

match_methodC3 <- "CBPS.weight"
match_methodC4 <- "CBPS.weight"

m.out.C3 <- matches_listC3[[match_methodC3]]
m.out.C4 <- matches_listC4[[match_methodC4]]

attr(m.out.noneC3, "outcome") <- "NUM_NONPROFITS"
attr(m.out.noneC3, "outcome.var") <- "NUM_NONPROFITS"
attr(m.out.noneC4, "outcome") <- "NUM_NONPROFITS"
attr(m.out.noneC4, "outcome.var") <- "NUM_NONPROFITS"
attr(m.out.C3, "outcome") <- "NUM_NONPROFITS"
attr(m.out.C3, "outcome.var") <- "NUM_NONPROFITS"
attr(m.out.C4, "outcome") <- "NUM_NONPROFITS"
attr(m.out.C4, "outcome.var") <- "NUM_NONPROFITS"
```

```{r}
# Estimates under no matching
res_no.match.C3 <- PanelEstimate(sets = m.out.noneC3,
                     panel.data = panel_dfC3,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

res_no.match.C4 <- PanelEstimate(sets = m.out.noneC4,
                     panel.data = panel_dfC4,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

# Estimates under CBPS Weighting
res.C3 <- PanelEstimate(sets = m.out.C3,
                     panel.data = panel_dfC3,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

res.C4 <- PanelEstimate(sets = m.out.C4,
                     panel.data = panel_dfC4,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)
```

```{r}
plot(res_no.match.C3, main = "Estimated Effects (LOG ASSETS) - No Matching")
summary(res_no.match.C3)
hist(res_no.match.C3$bootstrapped.estimates) 
# as.data.frame(summary(res_no.match.C3)) |> mutate(across(where(is.numeric), ~ exp(.x)-1))

plot(res_no.match.C4, main = "Estimated Effects (ASSETS) - No Matching")
summary(res_no.match.C4)
hist(res_no.match.C4$bootstrapped.estimates)

plot(res.C3, main = "Estimated Effects (Number of Nonprofits) - CBPS Weighting")
summary(res.C3)
hist(res.C3$bootstrapped.estimates)

plot(res.C4, main = "Estimated Effects (ASSETS) - CBPS Weighting")
summary(res.C4)
hist(res.C4$bootstrapped.estimates)
```
### Config 3 versus Config 4 - Expenses, No Stratification

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(tidyr)

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
save_name <- paste0("Lag", L)

folder_pathC3 <- "matches/with_svc/int_panelV2/config3/"
folder_pathC4 <- "matches/with_svc/int_panelV3/config4/"

# Retrieve the PanelMatch objects
matches_listC3 <- readRDS(paste0(folder_pathC3, save_name, ".rds"))
matches_listC4 <- readRDS(paste0(folder_pathC4, save_name, ".rds"))

# Retrieve the PanelData object
panel_dfC3 <- readRDS("panel_data/int_panelV2.rds")
panel_dfC4 <- readRDS("panel_data/int_panelV3.rds")

estimates_save_nameC3 <- "estimates/with_svc_intV2_con3_"
estimates_save_nameC4 <- "estimates/with_svc_intV3_con4_"

# No Matching
m.out.noneC3 <- matches_listC3$none
m.out.noneC4 <- matches_listC4$none

match_methodC3 <- "CBPS.weight"
match_methodC4 <- "CBPS.weight"

m.out.C3 <- matches_listC3[[match_methodC3]]
m.out.C4 <- matches_listC4[[match_methodC4]]

attr(m.out.noneC3, "outcome") <- "TOT_ASSET"
attr(m.out.noneC3, "outcome.var") <- "TOT_ASSET"
attr(m.out.noneC4, "outcome") <- "TOT_ASSET"
attr(m.out.noneC4, "outcome.var") <- "TOT_ASSET"
attr(m.out.C3, "outcome") <- "TOT_EXP"
attr(m.out.C3, "outcome.var") <- "TOT_EXP"
attr(m.out.C4, "outcome") <- "TOT_EXP"
attr(m.out.C4, "outcome.var") <- "TOT_EXP"
```

```{r}
# Estimates under no matching
res_no.match.C3 <- PanelEstimate(sets = m.out.noneC3,
                     panel.data = panel_dfC3,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

res_no.match.C4 <- PanelEstimate(sets = m.out.noneC4,
                     panel.data = panel_dfC4,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

# Estimates under CBPS Weighting
res.C3 <- PanelEstimate(sets = m.out.C3,
                     panel.data = panel_dfC3,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)

res.C4 <- PanelEstimate(sets = m.out.C4,
                     panel.data = panel_dfC4,
                     se.method = "bootstrap",
                     parallel = TRUE,
                     num.cores = 6,
                     number.iterations = 10000)
```

```{r}
plot(res_no.match.C3, main = "Estimated Effects (LOG ASSETS) - No Matching")
summary(res_no.match.C3)
hist(res_no.match.C3$bootstrapped.estimates) 
# as.data.frame(summary(res_no.match.C3)) |> mutate(across(where(is.numeric), ~ exp(.x)-1))

plot(res_no.match.C4, main = "Estimated Effects (ASSETS) - No Matching")
summary(res_no.match.C4)
hist(res_no.match.C4$bootstrapped.estimates)

plot(res.C3, main = "Estimated Effects (LOG EXPENSES) - CBPS Weighting")
summary(res.C3)
hist(res.C3$bootstrapped.estimates)

plot(res.C4, main = "Estimated Effects (Expenses) - CBPS Weighting")
summary(res.C4)
hist(res.C4$bootstrapped.estimates)
```


## Choosing the number of SVC
```{r}
library(data.table)
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(writexl)
library(haven)
library(geosphere)
source("matches_helper.R")
library(fields)
library(MASS)

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# These are some socioeconomic covariates we may evaluate balance for but need to convert from string to num
covs_to_check <- c("white_perc", "bachelors_perc", "poverty_perc", "unemployment", "med_family_income_adj", "med_household_income_adj", "total_population")
df <- df |> mutate(across(all_of(covs_to_check), as.numeric)) # make sure these are numeric

# These are the variables we will assess balance for
se_vars <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population")
other_vars <- c("TOT_REV", "TOT_ASSET")
ntee_vars <- c("NTEEV2_ART", "NTEEV2_EDU", "NTEEV2_ENV", "NTEEV2_HEL", "NTEEV2_HMS", "NTEEV2_HOS", "NTEEV2_IFA", "NTEEV2_MMB", "NTEEV2_PSB", "NTEEV2_REL", "NTEEV2_UNI", "NTEEV2_UNU", "NTEEV2_NA")

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA_nDis", "lat", "lng"), se_vars, other_vars, ntee_vars)))

df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
                                       .default = TOT_REV),
                   TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
                                       .default = TOT_ASSET))
df <- df |> mutate(across(c("TOT_REV", "TOT_ASSET"), log1p)) #log1p adds 1 first to avoid 0s

library(data.table)
df <- arrange(df, TAX_YEAR)
df.lag <- shift(df[,7:ncol(df)], n = 1:3, give.names = TRUE)
df <- bind_cols(df, df.lag)
rm(df.lag)

df <- df |> filter(TAX_YEAR >= 1994)

# Create new treatment variable that takes a distance-based cutoff adjustment
adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 25)
res <- add_exposed(panel_data = df, adj_list = adj_lists$union, treatment = "treat.FEMA_nDis")

df <- res$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))
sum(is.na(df$treat.FEMA.adj))

# Spatially Varying Coeffs model setup
df_sub <- df |> dplyr::select(CountyID, lat, lng) |> unique()

x <- cbind(df_sub$lat, df_sub$lng) # inputs: latitude and longitude
d <- rdist(x)
K <- Matern(d, smoothness=.5) # for smoothness paramter, should we do some sort of optimization?
E <- eigen(K)

options(repr.plot.width = 6, repr.plot.height = 4)
for (num_svc in seq(10,100,10)){
      x_s <- cbind(x[,1] * E$vectors[,1:num_svc], x[,2] * E$vectors[,1:num_svc]) # would have to play around with this number
      
      # add SVC back into panel data
      xs_vars <- paste0("x_s", seq_len(ncol(x_s)))
      colnames(x_s) <- xs_vars
      df_sub.temp <- cbind(df_sub, x_s)
      
      df.temp <- merge(df, df_sub.temp |> dplyr::select(-lat, -lng), by = "CountyID", all = TRUE)
      
      upper_formula <- paste(
        "treat.FEMA.adj ~ TAX_YEAR + bachelors_perc + med_household_income_adj + white_perc + total_population +",
        paste(ntee_vars, collapse = " + "), "+",
        paste(xs_vars, collapse = " + "))
      
      # regression
      ls <- glm(as.formula(upper_formula), data = df.temp, family = binomial(link = "logit"))
      summary(ls)

      yweight <- predict(ls, type="response")
      
      png(filename = paste0("svc_plots/int/svc", num_svc*2, ".png"))
      plot(sort(yweight), main = paste("# SVC =", num_svc*2), cex = 0.10)
      dev.off() # decided on 60... maybe 80 is the slightest bit smoother but if we do 60 its same for both
}
```

### Determining units that violate positivity assumption
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(data.table)

panel_df <- readRDS("panel_data/int_panelV2.rds")

se_vars <- colnames(panel_df)[which(colnames(panel_df) == "bachelors_perc"):which(colnames(panel_df) == "total_population")] 
org_vars <- colnames(panel_df)[which(colnames(panel_df) == "TOT_REV"):which(colnames(panel_df) == "NTEEV2_NA")] 
svc_vars <- colnames(panel_df)[which(colnames(panel_df) == "x_s1"):ncol(panel_df)]

panel_df  <- arrange(panel_df, TAX_YEAR)
df.lag <- shift(panel_df[,5:which(colnames(panel_df) == "NTEEV2_NA")], n = 1:3, give.names = TRUE)
panel_df <- bind_cols(panel_df, df.lag)
rm(df.lag)

vars_lagged <- colnames(panel_df)[which(colnames(panel_df) == "bachelors_perc_lag_1"):which(colnames(panel_df) == "NTEEV2_NA_lag_3")] 

panel_df <- panel_df |> filter(TAX_YEAR >= 1994 & TAX_YEAR <= 2020)

glm_formula <- paste("treat.FEMA.adj ~ TAX_YEAR +", 
                     paste0(c(se_vars, org_vars), collapse = " + "), "+",
                     paste0(vars_lagged, collapse = " + "), "+",
                     paste(svc_vars, collapse = " + "))
      
# regression
ls <- glm(as.formula(glm_formula), data = panel_df, family = binomial(link = "logit"))
summary(ls)

yweight <- predict(ls, type="response")

# png(filename = paste0("svc_plots/int/svc", num_svc*2, ".png"))
plot(sort(yweight), main = paste("# SVC =", 100), cex = 0.10)
# dev.off()
min(yweight) # 0.025
max(yweight) # 0.993
quantile(yweight) # 25-75% between 0.612 and 0.813 -- median is 0.711
```

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(data.table)

panel_df <- readRDS("panel_data/int_panelV3.rds")

se_vars <- colnames(panel_df)[which(colnames(panel_df) == "bachelors_perc"):which(colnames(panel_df) == "total_population")] 
org_vars <- colnames(panel_df)[which(colnames(panel_df) == "TOT_REV"):which(colnames(panel_df) == "NTEEV2_NA")] 
svc_vars <- colnames(panel_df)[which(colnames(panel_df) == "x_s1"):ncol(panel_df)]

panel_df  <- arrange(panel_df, TAX_YEAR)
df.lag <- shift(panel_df[,5:which(colnames(panel_df) == "NTEEV2_NA")], n = 1:3, give.names = TRUE)
panel_df <- bind_cols(panel_df, df.lag)
rm(df.lag)

vars_lagged <- colnames(panel_df)[which(colnames(panel_df) == "bachelors_perc_lag_1"):which(colnames(panel_df) == "NTEEV2_NA_lag_3")] 

panel_df <- panel_df |> filter(TAX_YEAR >= 1994 & TAX_YEAR <= 2020)

glm_formula <- paste("treat.FEMA.adj ~ TAX_YEAR +", 
                     paste0(c(se_vars, org_vars), collapse = " + "), "+",
                     paste0(vars_lagged, collapse = " + "), "+",
                     paste(svc_vars, collapse = " + "))
      
# regression
ls <- glm(as.formula(glm_formula), data = panel_df, family = binomial(link = "logit"))
summary(ls)

yweight <- predict(ls, type="response")

# png(filename = paste0("svc_plots/int/svc", num_svc*2, ".png"))
plot(sort(yweight), main = paste("# SVC =", 100), cex = 0.10)
# dev.off()
min(yweight) # 0.007
max(yweight) # 1
quantile(yweight) # 25-75% between 0.613 and 0.829 -- median is 0.723

prob1 <- names(yweight[yweight==max(yweight)])

panel_df[81068:81069,] # Brazoria County, TX (2543) and Brazos County, TX (2544) both in the year 2019

prob0 <- names(yweight[yweight==min(yweight)])
panel_df[77243,] # New York County (basically Manhattan) in year 2018, (1859)

library(PanelMatch)
panel_sub <- panel_df |> filter(CountyID %in% c(2543, 2544, 1859)) |> select(CountyID, TAX_YEAR, treat.FEMA.adj, TOT_REV) |> arrange(CountyID) |> 
      group_by(CountyID) |>
      mutate(ID = as.integer(cur_group_id())) |>
      ungroup()
panel_sub <- PanelData(panel.data = as.data.frame(panel_sub),
                      unit.id = "ID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA.adj",
                      outcome = "TOT_REV")
DisplayTreatment(panel.data = panel_sub, 
                                 color.of.treated = "#DC267F",
                                 color.of.untreated = "#648FFF",
                                 title = "Treatment Distribution Across Units and Time",
                                 xlab = "Tax Year", ylab = "County", 
                                 legend.position = "right",
                                 legend.labels = c("untreated", "treated"),
                                 hide.x.tick.label = TRUE, hide.y.tick.label = FALSE)
```

```{r}
library(PanelMatch)

L = 3
folder_path_match <- "matches/with_svc/int_panelV3/config4/" # "matches/with_svc/int_panelV2/config3/"
save_name <- paste0("Lag", L)
match_method <- "none"

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))
m.out <- matches_list[[match_method]]
summary(m.out)

matched_sets <- extract(m.out)

treated <- names(matched_sets)

my_list <- c("2543.2019", "2544.2019", "1859.2018")

intersect(treated, my_list) # empty!!! well duh... would only show up in control units

matched_controls <- weights(matched_sets)
matched_controls <- lapply(matched_controls, function(x) x[x > 0])
untreated <- lapply(matched_controls, function(x) names(x))
untreated <- unlist(untreated, use.names = FALSE)
untreated <- unique(untreated)

intersect(untreated, my_list) # empty... makes sense that treated units probably all have nonzero propensities

my_list <- c("2543.2020", "2544.2020", "1859.2019")

intersect(treated, my_list) # empty!!! well duh... would only show up in control units
```

# Computing Estimates and SE myself
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(tidyr)
library(stringr)
library(khroma)
source("matches_helper.R")

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
folder_path_match <- "matches/with_svc/no_int_panelV2/config1/"
save_name <- paste0("Lag", L)

# Retrieve the PanelMatch objects
matches_list <- readRDS(paste0(folder_path_match, save_name, ".rds"))
msets_list <- readRDS(paste0(folder_path_match, "extracted_", save_name, ".rds"))

# Winner: CBPS Matching with SVC Coefficients
m.out <- matches_list$CBPS.match
msets <- msets_list$CBPS.match

panel_df <- readRDS("panel_data/no_int_panelV2.rds")
panel_df <- as.data.frame(panel_df) |> select(CountyID, TAX_YEAR, TOT_REV)

lead <- 1
num.treated <- length(names(msets))
treated.units <- names(msets)
```

```{r}
# https://chatgpt.com/share/69b9d05d-da30-8003-9e25-56b7953f1e0e
df <- tibble(treated.units) |> 
      separate(treated.units, into = c("CountyID", "t"), sep = "\\.") |>
      mutate(key = paste(CountyID, t, sep = "."),
             CountyID = as.integer(CountyID),
             t = as.integer(t),
             t_plus_F = t + lead)

df <- merge(df, as.data.frame(panel_df), by.x = c("CountyID", "t_plus_F"), by.y = c("CountyID", "TAX_YEAR"), all.x = TRUE)

# msets[["1.1995"]]
matched_controls <- weights(msets)

# class(matched_controls[["1.1995"]])
# names(matched_controls[["1.1995"]])
# matched_controls[["1.1995"]][matched_controls[["1.1995"]] > 0]
matched_controls <- lapply(matched_controls, function(x) x[x > 0])

# Create a lookup key in panel_df in the format CountyID.TAX_YEAR
panel_df$key <- paste(panel_df$CountyID, panel_df$TAX_YEAR, sep = ".")

# Named vector for fast lookup of TOT_REV by "CountyID.TAX_YEAR"
tot_rev_lookup <- setNames(panel_df$TOT_REV, panel_df$key)

# treated_key <- "1.1995"
impute_counterfactual <- function(treated_key){
      # weights for this treated unit
      w <- matched_controls[[treated_key]]
      
      # extract year from treated_key
      yr <- sub(".*\\.", "", treated_key)
      
      # need to grab the total revenue in the year t+lead
      yr <- as.character(as.integer(yr) + lead)
      
      # control lookup keys: controlID.TAX_YEAR
      control_keys <- paste(names(w), yr, sep = ".")
      
      # matched controls' outcomes in the same year
      y_controls <- tot_rev_lookup[control_keys]
      
      # weighted sum
      sum(w * y_controls, na.rm = FALSE)
}

# Compute counterfactual for each treated unit
df$counterfactual <- sapply(df$key, impute_counterfactual)

# Compute estimate
delFL <- mean(df$TOT_REV) - mean(df$counterfactual)

sum(is.na(df$TOT_REV))
sum(is.na(df$counterfactual))
```


# Misc Analyses
Defining exposure with different distance cutoffs
```{r}
library(data.table)
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)
library(geosphere)
library(fastDummies)
source("matches_helper.R")

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# make sure these are numeric
df <- df |> mutate(across(all_of(c("bachelors_perc", "med_household_income_adj", "white_perc")), as.numeric)) 

# Add REGION dummy columns
df <- df |> dummy_cols(select_columns = c("REGION"), ignore_na = FALSE, remove_selected_columns = TRUE, remove_first_dummy = FALSE)

# These are the variables we will adjust for
adj_set <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET","TOT_EXP", "lat", "lng")

# Other variables
ntee_vars <- grep("^NTEEV2_", names(df), value = TRUE)
region_vars <- grep("^REGION_", names(df), value = TRUE)

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA"), adj_set, ntee_vars, region_vars)))

# Log transformations
df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
                                       .default = TOT_REV),
                   TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
                                       .default = TOT_ASSET),
                   TOT_EXP = case_when(TOT_EXP < 0 ~ 0,
                                       .default = TOT_EXP))

df <- df |> mutate(TOT_REV = log1p(TOT_REV),
                   TOT_ASSET = log1p(TOT_ASSET),
                   TOT_EXP = log1p(TOT_EXP),
                   med_household_income_adj = log1p(med_household_income_adj),
                   total_population = log(total_population))

# Create new treatment variable that takes a distance-based cutoff adjustment
adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 50)

res <- add_exposed(panel_data = df, adj_list = adj_lists$union, treatment = "treat.FEMA")
# res <- add_exposed(panel_data = df, adj_list = adj_lists$border, treatment = "treat.FEMA")

df <- res$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))

per_year_summary <- df |> group_by(TAX_YEAR) |>
      summarise(
            n_1 = sum(treat.FEMA.adj == 1, na.rm = TRUE),
            n_0 = sum(treat.FEMA.adj == 0, na.rm = TRUE),
            n = n())

paste("Treated")
quantile(per_year_summary$n_1, probs = seq(0,1,0.1))

paste("Control")
quantile(per_year_summary$n_0, probs = seq(0,1,0.1))

paste("Total")
quantile(per_year_summary$n, probs = seq(0,1,0.1))

# Adding spatially varying coefficients
df <- add_SVC(df, num_E_vec = 50)

# https://stackoverflow.com/questions/74815244/panelmatch-r-errors-please-convert-time-id-to-consecutive-integers-and-pleas
panel_df <- PanelData(panel.data = as.data.frame(df),
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA.adj",
                      outcome = "TOT_REV")

# rm(df)
saveRDS(panel_df, "panel_data/int_panel_d50.rds") # distance cutoff = 50
# saveRDS(panel_df, "panel_data/int_panel_d00.rds") # only adjacency list i.e. d = 0

DisplayTreatment(panel.data = panel_df, 
                  color.of.treated = "#DC267F", #"#FFB000"
                  color.of.untreated = "#648FFF", #"#785EF0"
                  title = "Exposure Distribution Across Units and Time (d = 0)",
                  xlab = "Tax Year", ylab = "County", 
                  legend.position = "right",
                  legend.labels = c("unexposed", "exposed"),
                  hide.x.tick.label = TRUE, hide.y.tick.label = TRUE, 
                  dense.plot = TRUE)
```

How many observations in the final sample had negative total revenue?
```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(tidyr)

# Set up the folder path and saved object names
L = 3 # lag, options: 1, 3, 5
save_name <- paste0("Lag", L)

folder_pathC3 <- "matches/with_svc/int_panelV2/config3/"
folder_pathC4 <- "matches/with_svc/int_panelV3/config4/"

# Retrieve the PanelMatch objects
matches_listC3 <- readRDS(paste0(folder_pathC3, save_name, ".rds"))
matches_listC4 <- readRDS(paste0(folder_pathC4, save_name, ".rds"))

# Retrieve the PanelData object
panel_dfC3 <- readRDS("panel_data/int_panelV2.rds")
panel_dfC4 <- readRDS("panel_data/int_panelV3.rds")

match_methodC3 <- "CBPS.weight"
match_methodC4 <- "CBPS.weight"

m.out.C3 <- matches_listC3[[match_methodC3]]
m.out.C4 <- matches_listC4[[match_methodC4]]
```

```{r}
# I need the treated units for both configurations
msets.C3 <- extract(m.out.C3)
msets.C4 <- extract(m.out.C4)

treated.C3 <- names(msets.C3)
treated.C4 <- names(msets.C4)

df.c3 <- tibble(treated.C3) |> 
      separate(treated.C3, into = c("CountyID", "t"), sep = "\\.") |>
      mutate(key = paste(CountyID, t, sep = "."),
             CountyID = as.integer(CountyID),
             t = as.integer(t),
             t_plus_1 = t + 1,
             t_plus_2 = t + 2,
             t_plus_3 = t + 3,
             t_plus_4 = t + 4,
             t_plus_5 = t + 5)

df.c4 <- tibble(treated.C4) |> 
      separate(treated.C4, into = c("CountyID", "t"), sep = "\\.") |>
      mutate(key = paste(CountyID, t, sep = "."),
             CountyID = as.integer(CountyID),
             t = as.integer(t),
             t_plus_1 = t + 1,
             t_plus_2 = t + 2,
             t_plus_3 = t + 3,
             t_plus_4 = t + 4,
             t_plus_5 = t + 5)

all.equal(df.c3, df.c4) # all the same, as they should be

matched_controls.C3 <- weights(msets.C3)
matched_controls.C3 <- lapply(matched_controls.C3, function(x) x[x > 0])
untreated.C3 <- lapply(matched_controls.C3, function(x) names(x))
result <- lapply(names(untreated.C3), function(nm) {
  year <- sub(".*\\.", "", nm)  # extract year after the dot
  paste0(untreated.C3[[nm]], ".", year)
})
names(result) <- names(untreated.C3)
untreated.C3 <- unlist(result, use.names = FALSE) #1,068,234
untreated.C3 <- unique(untreated.C3) # 7722 county-years

matched_controls.C4 <- weights(msets.C4)
matched_controls.C4 <- lapply(matched_controls.C4, function(x) x[x > 0])
untreated.C4 <- lapply(matched_controls.C4, function(x) names(x))
result <- lapply(names(untreated.C4), function(nm) {
  year <- sub(".*\\.", "", nm)  # extract year after the dot
  paste0(untreated.C4[[nm]], ".", year)
})
names(result) <- names(untreated.C4)
untreated.C4 <- unlist(result, use.names = FALSE) #1,068,234
untreated.C4 <- unique(untreated.C4) # 7722 county-years

length(intersect(untreated.C3, untreated.C4)) # 7722 -- word, same matched control set

df.control <- tibble(untreated.C3) |> 
      separate(untreated.C3, into = c("CountyID", "t"), sep = "\\.") |>
      mutate(key = paste(CountyID, t, sep = "."),
             CountyID = as.integer(CountyID),
             t = as.integer(t),
             t_plus_1 = t + 1,
             t_plus_2 = t + 2,
             t_plus_3 = t + 3,
             t_plus_4 = t + 4,
             t_plus_5 = t + 5)

final_data <- rbind(df.c3 |> mutate(status = "treated"), df.control |> mutate(status = "control"))
final_data <- final_data |> mutate(t_minus_1 = t - 1)
```

```{r}
# Now, the question is: how many of these units originally had negative rev, assets, or expenses
final_long <- final_data |> 
      select(-t) |>
      pivot_longer(all_of(c("t_minus_1", "t_plus_1", "t_plus_2", "t_plus_3", "t_plus_4", "t_plus_5")),
                   names_to = "time",
                   values_to = "TAX_YEAR")

final_long <- merge(final_long, panel_dfC4 |> select(CountyID, TAX_YEAR, TOT_REV, TOT_ASSET, TOT_EXP),
                    by = c("CountyID", "TAX_YEAR"), all.x = TRUE)

sum(final_long$TOT_REV < 0, na.rm = TRUE) # 18 
sum(final_long$TOT_REV[final_long$TOT_REV < 0], na.rm = TRUE) # -103,224,361 total

sum(final_long$TOT_REV < 0 & final_long$status == "treated", na.rm = TRUE) # 12
sum(final_long$TOT_REV[final_long$TOT_REV < 0 & final_long$status == "treated"], na.rm = TRUE) # -94,540,069

final_long |> filter(TOT_REV < 0) # of the 6 control units, 3 had negative in t-1 (which would inflate their pre-treatment outcome)

sum(final_long$TOT_REV[final_long$TOT_REV > 0], na.rm = TRUE) # this is in the trillions 117,140,800,000,000

sum(final_long$TOT_ASSET < 0, na.rm = TRUE) # 2 
sum(final_long$TOT_ASSET[final_long$TOT_ASSET < 0], na.rm = TRUE) # -56053

sum(final_long$TOT_EXP < 0, na.rm = TRUE) # 4 
sum(final_long$TOT_EXP[final_long$TOT_EXP < 0], na.rm = TRUE) # -786,841,336
final_long |> filter(TOT_EXP < 0) 

panel_dfC4 |> filter(TOT_REV < 0) # Only 44 county-years
panel_dfC4 |> filter(TOT_EXP < 0) # Only 2 county-years
panel_dfC4 |> filter(TOT_ASSET < 0) # Only 13 county-years

panel_dfC4 |> filter(TOT_REV < 0 | TOT_EXP < 0 | TOT_ASSET < 0) # 59
```

What counties are not in the final sample?
```{r}
missing_rev <- as.data.frame(panel_dfC3 |> filter(is.na(TOT_REV)) |> select(CountyID, TAX_YEAR, treat.FEMA.adj)) 

missing_rev |> group_by(CountyID) |>
      summarise(n = n()) |> arrange(-n) # 121 counties

by_year <- missing_rev |> group_by(TAX_YEAR) |>
      summarise(n = n())
plot(x = by_year$TAX_YEAR, y = by_year$n)
hist(missing_rev$TAX_YEAR)
hist(missing_rev$CountyID)

counties <- c(568, 1617, 1712, 1659, 2744)

panel_dfC3 |> select(CountyID, geoid_2010, REGION_MIDWEST, REGION_NORTHEAST, REGION_SOUTH, REGION_WEST) |> 
      filter(CountyID %in% counties) |> unique() |> as.data.frame()

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

other_counties <- setdiff(unique(missing_rev$CountyID), counties)
df <- df |> filter(CountyID %in% other_counties)

df <- df |> select(CountyID, geoid_2010, REGION) |> distinct()

panel_df <- PanelData(panel.data = as.data.frame(df),
                      unit.id = "CountyID",
                      time.id = "TAX_YEAR",
                      treatment = "treat.FEMA",
                      outcome = "TOT_REV")
length(unique(panel_df$CountyID))

table(df$REGION)
```

How do adjacency-border veruss adjacency-union compare?
```{r}
library(data.table)
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
library(PanelMatch)
library(ggplot2)
library(writexl)
library(haven)
library(geosphere)
library(fastDummies)
source("matches_helper.R")

df <- readRDS("county_long.rds")
df <- df |> 
      group_by(geoid_2010) |>
      mutate(CountyID = as.integer(cur_group_id())) |>
      ungroup()

# These are some socioeconomic covariates we may evaluate balance for but need to convert from string to num
covs_to_check <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population")
df <- df |> mutate(across(all_of(covs_to_check), as.numeric)) # make sure these are numeric

# Add REGION dummy columns
df <- df |> dummy_cols(select_columns = c("REGION"), ignore_na = FALSE, remove_selected_columns = TRUE, remove_first_dummy = FALSE)

# These are the variables we will adjust for
adj_set <- c("bachelors_perc", "med_household_income_adj", "white_perc",  "total_population", "TOT_REV", "TOT_ASSET", "lat", "lng")

# Other variables
ntee_vars <- grep("^NTEEV2_", names(df), value = TRUE)
region_vars <- grep("^REGION_", names(df), value = TRUE)

df <- df |> dplyr::select(all_of(c(c("CountyID", "TAX_YEAR", "geoid_2010", "treat.FEMA_nDis"), adj_set, ntee_vars, region_vars)))

# log transform
#saveRDS(df |> filter(TOT_REV < 0 | TOT_ASSET < 0), "misc_analysis/no_intV1_negative.rds")
df <- df |> mutate(TOT_REV = case_when(TOT_REV < 0 ~ 0,
                                       .default = TOT_REV),
                   TOT_ASSET = case_when(TOT_ASSET < 0 ~ 0,
                                       .default = TOT_ASSET))
df <- df |> mutate(across(c("TOT_REV", "TOT_ASSET"), log1p)) #log1p adds 1 first to avoid 0s

# Create new treatment variable that takes a distance-based cutoff adjustment
# adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 25)
adj_lists <- create_adj_list(panel_data = df, dist_cutoff = 35)

res_border <- add_exposed(panel_data = df, adj_list = adj_lists$border, treatment = "treat.FEMA_nDis")
res_dist <- add_exposed(panel_data = df, adj_list = adj_lists$dist, treatment = "treat.FEMA_nDis")
res_union <- add_exposed(panel_data = df, adj_list = adj_lists$union, treatment = "treat.FEMA_nDis")

df_border <- res_border$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))
# 1859 is New York County
df_border$treat.FEMA.adj[((df_border$CountyID == 1859) & (df_border$TAX_YEAR == 2001))] <- NA
sum(is.na(df_border$treat.FEMA.adj))

df_dist <- res_dist$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))
# 1859 is New York County
df_dist$treat.FEMA.adj[((df_dist$CountyID == 1859) & (df_dist$TAX_YEAR == 2001))] <- NA
sum(is.na(df_dist$treat.FEMA.adj))

df_union <- res_union$df |> mutate(treat.FEMA.adj = replace_na(treat.FEMA.adj, 1))
# 1859 is New York County
df_union$treat.FEMA.adj[((df_union$CountyID == 1859) & (df_union$TAX_YEAR == 2001))] <- NA
sum(is.na(df_union$treat.FEMA.adj))

# # distance = 25
# sum(df$treat.FEMA_nDis, na.rm = TRUE)
# sum(df_border$treat.FEMA.adj, na.rm = TRUE) # 68182
# sum(df_dist$treat.FEMA.adj, na.rm = TRUE) # 47372
# sum(df_union$treat.FEMA.adj, na.rm = TRUE) # 68609 -- extra 427 treated compared with just border

# distance = 35
sum(df$treat.FEMA_nDis, na.rm = TRUE)
sum(df_border$treat.FEMA.adj, na.rm = TRUE) # 68182
sum(df_dist$treat.FEMA.adj, na.rm = TRUE) # 62003
sum(df_union$treat.FEMA.adj, na.rm = TRUE) # 70855 -- extra 2673 treated compared with just border
```


